---
title: "NWGOM Population Genetics"
author: "Ryan Eckert"
output:
  html_document:
    theme: flatly
    toc: yes
    toc_depth: 3
    toc_float: yes
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = '../data/index.html',
      envir = globalenv()
    )
  })
---

<a href="https://github.com/RyanEckert/Nwgom_PopGen" class="github-corner" aria-label="View source on GitHub"><svg width="80" height="80" viewBox="0 0 250 250" style="fill:#2C3E50; color:#fff; position: absolute; top: 0; border: 0; right: 0;" aria-hidden="true"><path d="M0,0 L115,115 L130,115 L142,142 L250,250 L250,0 Z"></path><path d="M128.3,109.0 C113.8,99.7 119.0,89.6 119.0,89.6 C122.0,82.7 120.5,78.6 120.5,78.6 C119.2,72.0 123.4,76.3 123.4,76.3 C127.3,80.9 125.5,87.3 125.5,87.3 C122.9,97.6 130.6,101.9 134.4,103.2" fill="currentColor" style="transform-origin: 130px 106px;" class="octo-arm"></path><path d="M115.0,115.0 C114.9,115.1 118.7,116.5 119.8,115.4 L133.7,101.6 C136.9,99.2 139.9,98.4 142.2,98.6 C133.8,88.0 127.5,74.4 143.8,58.0 C148.5,53.4 154.0,51.2 159.7,51.0 C160.3,49.4 163.2,43.6 171.4,40.1 C171.4,40.1 176.1,42.5 178.8,56.2 C183.1,58.6 187.2,61.8 190.9,65.4 C194.5,69.0 197.7,73.2 200.1,77.6 C213.8,80.2 216.3,84.9 216.3,84.9 C212.7,93.1 206.9,96.0 205.4,96.6 C205.1,102.4 203.0,107.8 198.3,112.5 C181.9,128.9 168.3,122.5 157.7,114.1 C157.9,116.9 156.7,120.9 152.7,124.9 L141.0,136.5 C139.8,137.7 141.6,141.9 141.8,141.8 Z" fill="currentColor" class="octo-body"></path></svg></a>

```{=html}
<style>.github-corner:hover .octo-arm{animation:octocat-wave 560ms ease-in-out}@keyframes octocat-wave{0%,100%{transform:rotate(0)}20%,60%{transform:rotate(-25deg)}40%,80%{transform:rotate(10deg)}}@media (max-width:500px){.github-corner:hover .octo-arm{animation:none}.github-corner .octo-arm{animation:octocat-wave 560ms ease-in-out}}</style>
```
#### version: `r library(magrittr)` `r Sys.Date() %>% format(format="%d %B, %Y")`

#### [GitHub repository](https://github.com/RyanEckert/NWGOM_PopGen.git){target="_blank"}

# About this document

------------------------------------------------------------------------

#### All analyses preformed with R version `r getRversion()`.

# Basic setup of R environment

------------------------------------------------------------------------

```{r, setup, include = FALSE}
# setup rmarkdown environment first

knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
options(width = 88)

library(magrittr)

#setting working directory to the directory containing this .Rmd file
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

## Loading required packages

For the following analyses we will require the use of a number of different R packages. We can use the following code to quickly load in the packages and install any packages not previously installed in the R console.

```{r, load packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")

pacman::p_load_gh("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis", "ropensci/rnaturalearthhires", "KarstensLab/microshades")

pacman::p_load("flextable", "cowplot", "car", "ggrepel", "ggspatial", "paletteer", "patchwork", "rnaturalearth", "sf", "Hmisc", "MCMC.OTU", "pairwiseAdonis", "RColorBrewer", "Redmonder", "lubridate", "officer", "adegenet", "dendextend", "gdata", "ggdendro", "hierfstat", "kableExtra", "poppr", "reshape2", "StAMPP", "vcfR", "vegan", "boa", "magick", "sdmpredictors", "ggcorrplot", "tidyverse", "TeachingDemos", "LaplacesDemon", "adespatial", "ggnewscale", "ggbeeswarm", "multcomp", "rstatix", "R.utils", "graph4lg", "ggstance", "ggtreeExtra", "ggtree", "multcompView")

options("scipen" = 10)

load("nwgom.RData")

```

<br>

## Themes and colors

Making color palettes to use throughout all plots

```{r, colors}
gomPal = c("#D53E4F", "#F88D51", "#FEE08B", "#66C2A5", "#3288BD")

boundPal = c("gray30", paletteer_d("vapoRwave::vapoRwave")[10])

pink = "#FF6A8BFF"
purple = paletteer_d("vapoRwave::vapoRwave")[10]

# mcavKColPal = c(colorRampPalette(colors = c("#1B9E77","#FFFFFF"))(3)[1:2], "azure3")
# sintKColPal = c(colorRampPalette(colors = c("#D95F02","#FFFFFF"))(3)[1:2], "azure3")
# ofavKColPal = colorRampPalette(colors = c("#7570B3","#FFFFFF"))(4)[1:3]
# xestoKColPal = c(colorRampPalette(colors = c("#E7298A","#FFFFFF"))(7)[c(1:6)], "azure3")
mcavKColPal = c(colorRampPalette(colors = c("#1B9E77","#FFFFFF"))(7)[c(1,5,2,3,4,6)], "azure3")
sintKColPal = c(colorRampPalette(colors = c("#D95F02","#FFFFFF"))(7)[c(1,5,2,3,4,6)], "azure3")
ofavKColPal = colorRampPalette(colors = c("#7570B3","#FFFFFF"))(7)[c(1,5,3,2,4,6)]
xestoKColPal = c(colorRampPalette(colors = c("#E7298A","#FFFFFF"))(7)[c(1:6)], "azure3")

profPal = rev(c(microshades_palette("micro_green", 5), microshades_palette("micro_cvd_turquoise", 5),  microshades_palette("micro_cvd_orange", 3),microshades_palette("micro_cvd_purple", 1, lightest = F), microshades_palette("micro_purple", 5)))

colPalZoox = c("#6A4C93", "#1BE7FF", "#C6FCA2", "#FF6A8B")

```


Plot themes

```{r, themes}
dendTheme = theme(axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.line.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.y = element_text(size = 10, color = "black", angle = 90),
  axis.text.y = element_text(size = 8, color = "black"),
  axis.line.y = element_line(),
  axis.ticks.y = element_line(),
  panel.grid = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  legend.key.size = unit(0.75, "line"),
  legend.key = element_blank(),
  legend.title = element_text(size = 10),
  legend.text = element_text(size = 8),
  legend.direction = "horizontal",
  legend.box = "vertical",
  legend.position = c(0.13, 0.1))

pcaTheme = theme(axis.title.x = element_text(color = "black", size = 10),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.title.y = element_text(color = "black", size = 10),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        legend.position = "none",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.key.size = unit(5, "pt"),
        panel.border = element_rect(color = "black", linewidth = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
  
admixTheme = theme(
  panel.grid = element_blank(),
  panel.background = element_rect(fill = "gray85"),
  plot.background = element_blank(),
  panel.border = element_rect(fill = NA, color = "black", linewidth = 0.75, linetype = "solid"),
  panel.spacing.x = grid:::unit(0.05, "lines"),
  panel.spacing.y = grid:::unit(0.05, "lines"),
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.ticks.x = element_blank(),
  axis.ticks.y = element_blank(),
  axis.title = element_blank(),
  strip.background.x = element_blank(),
  strip.background.y = element_blank(),
  strip.text = element_text(size = 8),
  strip.text.y.left = element_text(size = 10, angle = 90),
  strip.text.x.bottom = element_text(vjust = 1, color = NA),
  legend.key = element_blank(),
  legend.position = "none",
  legend.title = element_text(size = 8))

violinTheme = theme(axis.title = element_text(color = "black", size = 12),
        axis.ticks = element_blank(),
        axis.text.x = element_text(color = "black", size = 10),
        legend.position = "none",
        legend.key.size = unit(0.3, 'cm'),
        panel.border = element_rect(color = "black", linewidth = .5),
        panel.background = element_blank(),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

zooxTheme = theme(plot.title = element_text(),
        panel.grid = element_blank(),
        # panel.background = element_blank(),
        panel.background = element_rect(fill = "gray85"),
        panel.border = element_rect(fill = NA, color = "black", linewidth = 0.75, linetype = "solid"),
        panel.spacing.x = grid:::unit(0.05, "lines"),
        panel.spacing.y = grid:::unit(0.82, "lines"),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title = element_blank(),
        strip.background.x = element_blank(),
        strip.background.y = element_blank(),
        strip.text = element_text(size = 12),
        strip.text.y.left = element_text(size = 12, angle = 90),
        strip.text.x.bottom = element_text(vjust = -.1, color = NA),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.key.size = unit(0.75, "line"),
        legend.key = element_blank(),
        legend.position = "right",
        legend.direction = "vertical",
        legend.box = "vertical")

```

<br>


# Sampling info 
***
## Map of study sites

<br>

```{r, map data}
nwgomSites = read.csv("../data/nwgomMetaData.csv", header = TRUE)
nwgomSites$depthZone = factor(nwgomSites$depthZone)
nwgomSites$depthZone = factor(nwgomSites$depthZone, levels = levels(nwgomSites$depthZone)[c(2,1)])

nwgomSites$site = factor(nwgomSites$site)
nwgomSites$bank = factor(nwgomSites$bank)
nwgomSites$bank = factor(nwgomSites$bank, levels = levels(nwgomSites$bank)[c(5, 2, 1, 3, 4)])
nwgomSites$date = mdy(nwgomSites$date) %>% format("%d %b %Y")

nwgomBanks = nwgomSites %>% group_by(bank) %>% summarise(latDD = mean(latDD), longDD = mean(longDD), n = n()) %>% droplevels()

nwgomsampleSites = nwgomSites %>% group_by(bank, site, depthZone) %>% summarise(latDD = min(latDD), longDD = min(longDD))

fgbnmsBounds = read_sf("../data/shp/FGBNMS_py.shp") %>% st_transform(crs = 4326)
fgbnmsBounds$Bank = c("NA","NA","Geyer","NA","Bright","WFGB","NA","McGrail","NA","NA","NA","McGrail","EFGB","NA","NA","NA","NA","NA","NA")

fgbnmsBounds$Bank = factor(fgbnmsBounds$Bank)
fgbnmsBounds$Bank = factor(fgbnmsBounds$Bank, levels = levels(fgbnmsBounds$Bank)[c(6, 2, 1, 3, 4,5)])

states = st_as_sf(ne_states(country = c("United States of America")), scale = "count",  crs = 4326) %>% filter(name_en %in% c("Florida", "Alabama","Mississippi", "Louisiana", "Texas", "Georgia", "South Carolina"))

countries = st_as_sf(ne_countries(country = c("Cuba", "Mexico", "Cayman Islands", "The Bahamas", "Belize", "Guatemala"), scale = "Large"), crs = 4326)
# bahamas = read_sf("../data/shp/bahamasShoreline.shp") %>% st_transform(crs = 4326)
# cuba = read_sf("../data/shp/cubaShoreline.shp") %>% st_transform(crs = 4326)
# florida = read_sf("../data/shp/floridaShoreline.shp") %>% st_transform(crs = 4326)

wfgbBathy = read_sf("../data/shp/wb_10m_Cleaned.shp") %>% st_transform(crs = 4326)
efgbBathy = read_sf("../data/shp/eb_10m_Cleaned.shp") %>% st_transform(crs = 4326)
brightBathy = read_sf("../data/shp/Bright_10m_Cleaned.shp") %>% st_transform(crs = 4326)
geyerBathy = read_sf("../data/shp/Geyer_10m_Cleaned.shp") %>% st_transform(crs = 4326)
mcgrailBathy = read_sf("../data/shp/McGrail_10m_Cleaned.shp") %>% st_transform(crs = 4326)

```


```{r, maps}
nwgomMap = ggplot() +
  geom_sf(data = fgbnmsBounds %>% filter(Bank != "NA"), aes(fill = Bank), linewidth = 0.25, color = "black") +
  geom_sf(data = fgbnmsBounds %>% filter(Bank == "NA"), fill = "black", linewidth = 0.25, color = "black", alpha = 0.1) +
  geom_sf(data = states, fill = "white", linewidth = 0.3) +
  scale_fill_manual(values = c(gomPal,"black"), name = "Bank",) +
  geom_point(data = nwgomSites, aes(x = longDD, y = latDD, shape = depthZone), color = NA, fill = NA) +
  scale_shape_manual(values = c(21, 23), name = "Depth") +
  coord_sf(xlim = c(-94.5, -91.5), ylim = c(26.75, 29.75)) +
  annotation_scale(location = "br") +
  guides(fill = guide_legend(override.aes = list(shape = 22, color = "black", size = 2.5, stroke = 0.25), order = 1), shape = guide_legend(override.aes = list(size = c(2.5, 2.25), stroke = 0.25, color = "black"), order = 2), color = "none") +
  theme_bw() +
  theme(panel.background = element_rect(fill = "aliceblue"),

        plot.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1, fill = NA),
        axis.title = element_blank(),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = c(0.9085, 0.882),
        legend.box.background = element_rect(linewidth = 0.45, fill = "white"),
        legend.title = element_text(color = "black", size = 8),
        legend.text = element_text(color = "black", size = 8),
        legend.spacing = unit(-5, "pt"),
        legend.key.size = unit(5, "pt"),
        legend.key = element_blank(),
        legend.background = element_blank()
        )

nwgomMap

largeMap = ggplot() +
  geom_sf(data = states, fill = "white", linewidth = 0.3) +
  geom_sf(data = countries, fill = "white", linewidth = 0.3) +
  geom_sf(data = fgbnmsBounds, fill = "black", linewidth = 0.1, color = "black", alpha = 0.1) +
  geom_rect(aes(xmin = -95, xmax = -91, ymin = 26, ymax = 30), fill = NA, color = "black", linewidth = 0.75) +
  coord_sf(xlim = c(-99, -80), ylim = c(18, 32)) +
  theme_bw() +
  theme(legend.title = element_text(size = 9, face = "bold"),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(color = "black", size = 1, fill = NA),
        legend.position = "none",
        plot.background = element_blank())

largeMap


inset = ggplot() +
  geom_sf(data = wfgbBathy, color = "gray75", size = 0.25) +
  geom_sf(data = efgbBathy, color = "gray75", size = 0.25) +
  geom_sf(data = brightBathy, color = "gray75", size = 0.25) +
  geom_sf(data = geyerBathy, color = "gray75", size = 0.25) +
  geom_sf(data = mcgrailBathy, color = "gray75", size = 0.25) +
  geom_sf(data = fgbnmsBounds %>% filter(Bank != "NA"), aes(fill = Bank), linewidth = 0.4, color = "black", alpha = 0.2) +
  geom_sf(data = fgbnmsBounds %>% filter(Bank == "NA"), fill = "black", linewidth = 0.4, color = "black", alpha = 0.2) +
  geom_point(data = nwgomsampleSites, aes(x = longDD, y = latDD, fill = bank, shape = depthZone, size = depthZone), alpha = 0.75) +
  scale_fill_manual(values = gomPal, name = "Bank") +
  scale_shape_manual(values = c(21, 23), name = "Depth") +
  scale_size_manual(values = c(2.5, 2.25)) +
  theme_bw() +
  theme(legend.title = element_text(size = 9, face = "bold"),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(color = "black", size = 1, fill = NA),
        legend.position = "none",
        plot.background = element_blank())


wfgb = inset +
  geom_label(aes(x = -93.8269, y = 27.882), label = "WFGB", fill = gomPal[1], color = "black", size = 3) +
  annotation_scale(location = "br") +
  coord_sf(xlim = c(-93.83, -93.808), ylim = c(27.867, 27.883)) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10)

efgb = inset +
  geom_label(aes(x = -93.6074, y = 27.92435), label = "EFGB", fill = gomPal[2], color = "black", size = 3) +
  annotation_scale(location = "br") +
  coord_sf(xlim = c(-93.61, -93.59), ylim = c(27.905, 27.925)) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10)

bright = inset +
  geom_label(aes(x = -93.3135, y = 27.8984), label = "Bright", fill = gomPal[3], color = "black", size = 3) +
  annotation_scale(location = "br") +
  coord_sf(xlim = c(-93.316, -93.296), ylim = c(27.879, 27.899)) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10)

geyer = inset +
  geom_label(aes(x = -93.07329, y = 27.8614), label = "Geyer", fill = gomPal[4], color = "black", size = 3) +
  annotation_scale(location = "br") +
  coord_sf(xlim = c(-93.076, -93.056), ylim = c(27.842, 27.862)) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10)

mcgrail = inset +
  geom_label(aes(x = -92.6018, y = 27.9691), label = "McGrail", fill = gomPal[5], color = "black", size = 3) +
  annotation_scale(location = "br") +
  coord_sf(xlim = c(-92.605, -92.585), ylim = c(27.955, 27.97)) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10)


map = (nwgomMap + inset_element(largeMap, top = 1.12, right = 0.33, bottom = 0.63, left = -0.0075, ignore_tag = TRUE) +
         inset_element(wfgb, top = 0.8, right = 0.2875, bottom = 0.5, left = -0.0075, ignore_tag = TRUE) +
         inset_element(efgb, top = 0.33, right = 0.2875, bottom = 0.04, left = -0.0075, ignore_tag = TRUE) +
         inset_element(bright, top = 0.33, right = 0.68, bottom = 0.04, left = 0.33, ignore_tag = TRUE) +
         inset_element(geyer, top = 0.33, right = 1.0, bottom = 0.04, left = 0.705, ignore_tag = TRUE) +
         inset_element(mcgrail, top = 0.8, right = 1.0, bottom = 0.5, left = 0.7, ignore_tag = TRUE)
  ) 
  
ggsave("../figures/figure1.png", plot = map, height = 7, width = 7, units = "in", dpi = 300)

```




# Detecting clones
***

## *Montastraea cavernosa*
***

```{r, mcav clones}
mcavCloneBams = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "M. cavernosa") %>% dplyr::filter(!sample %in% c("MGM010")) # list of bam files

mcavCloneMa = as.matrix(read.table("../data/mcav/mcavClones.ibsMat")) # reads in IBS matrix produced by ANGSD 

dimnames(mcavCloneMa) = list(mcavCloneBams[,1],mcavCloneBams[,1])
mcavClonesHc = hclust(as.dist(mcavCloneMa),"ave")

mcavClonePops = mcavCloneBams$bank
mcavCloneDepth = mcavCloneBams$depthZone

mcavCloneDend = mcavCloneMa %>% as.dist() %>% hclust(.,"ave") %>% as.dendrogram()
mcavCloneDData = mcavCloneDend %>% dendro_data()

# Making the branches hang shorter so we can easily see clonal groups
mcavCloneDData$segments$yend2 = mcavCloneDData$segments$yend
for(i in 1:nrow(mcavCloneDData$segments)) {
  if (mcavCloneDData$segments$yend2[i] == 0) {
    mcavCloneDData$segments$yend2[i] = (mcavCloneDData$segments$y[i] - 0.01)}}

mcavCloneDendPoints = mcavCloneDData$labels
mcavCloneDendPoints$pop = mcavClonePops[order.dendrogram(mcavCloneDend)]
mcavCloneDendPoints$depth=mcavCloneDepth[order.dendrogram(mcavCloneDend)]
rownames(mcavCloneDendPoints) = mcavCloneDendPoints$label

# Making points at the leaves to place symbols for populations
point = as.vector(NA)
for(i in 1:nrow(mcavCloneDData$segments)) {
  if (mcavCloneDData$segments$yend[i] == 0) {
    point[i] = mcavCloneDData$segments$y[i] - 0.01
  } else {
    point[i] = NA}}

mcavCloneDendPoints$y = point[!is.na(point)]

mcavTechReps = c("MGM002.1", "MGM002.2", "MGM002.3", "MGM008.1", "MGM008.2", "MGM008.3", "MGM013.1", "MGM013.2", "MGM013.3", "MGM024.1", "MGM024.2", "MGM038.1", "MGM038.2", "MGM038.3", "MGM072.1", "MGM072.2", "MGM072.3")

mcavCloneDendPoints$depth = factor(mcavCloneDendPoints$depth)

mcavCloneDendPoints$depth = factor(mcavCloneDendPoints$depth,levels(mcavCloneDendPoints$depth)[c(2,1)])

mcavCloneDendPoints$pop = factor(mcavCloneDendPoints$pop)

mcavCloneDendPoints$pop = factor(mcavCloneDendPoints$pop,levels(mcavCloneDendPoints$pop)[c(5, 2, 1, 4, 3)])

mcavCloneDendA = ggplot() +
  geom_segment(data = segment(mcavCloneDData), aes(x = x, y = y, xend = xend, yend = yend2), size = 0.5) +
  geom_point(data = mcavCloneDendPoints, aes(x = x, y = y, fill = pop, shape = depth), size = 4, stroke = 0.25) +
  scale_fill_manual(values =gomPal, name= "Bank")+
  scale_shape_manual(values = c(21, 23), name = "Depth Zone")+
  geom_hline(yintercept = 0.1575, color = pink, lty = 5, size = 0.75) + # creating a dashed line to indicate a clonal distance threshold
  geom_text(data = subset(mcavCloneDendPoints, subset = label %in% mcavTechReps), aes(x = x, y = (y - .015), label = label), angle = 90) + # spacing technical replicates further from leaf
  geom_text(data = subset(mcavCloneDendPoints, subset = !label %in% mcavTechReps), aes(x = x, y = (y - .010), label = label), angle = 90) +
  labs(y = "Genetic distance (1 - IBS)") +
  guides(fill = guide_legend(override.aes = list(shape = 22)))+
  coord_cartesian(xlim = c(0,182), ylim = c(0.075, 0.3),expand = FALSE) +
  theme_classic()

mcavCloneDend = mcavCloneDendA + theme(
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.line.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.y = element_text(size = 12, color = "black", angle = 90),
  axis.text.y = element_text(size = 10, color = "black"),
  axis.line.y = element_line(),
  axis.ticks.y = element_line(),
  panel.grid = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  plot.background = element_blank(),
  legend.key = element_blank(),
  legend.title = element_text(size = 12),
  legend.text = element_text(size = 10),
  legend.position = "bottom")

mcavCloneDend

ggsave("../figures/extras/mcavCloneDend.png", plot = mcavCloneDend, height = 9, width = 32, units = "in", dpi = 300)

```

<br>


## *Stephanocoenia intersepta*
***

```{r, sint clones}
sintCloneBams = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "S. intersepta") %>% dplyr::filter(!sample %in% c("SGM013", "SGM169", "SGM177"))

sintCloneMa = as.matrix(read.table("../data/sint/sintClones.ibsMat")) # reads in IBS matrix produced by ANGSD 

dimnames(sintCloneMa) = list(sintCloneBams[,1],sintCloneBams[,1])
sintClonesHc = hclust(as.dist(sintCloneMa),"ave")

sintClonePops = sintCloneBams$bank
sintCloneDepth = sintCloneBams$depthZone

sintCloneDend = sintCloneMa %>% as.dist() %>% hclust(.,"ave") %>% as.dendrogram()
sintCloneDData = sintCloneDend %>% dendro_data()

# Making the branches hang shorter so we can easily see clonal groups
sintCloneDData$segments$yend2 = sintCloneDData$segments$yend
for(i in 1:nrow(sintCloneDData$segments)) {
  if (sintCloneDData$segments$yend2[i] == 0) {
    sintCloneDData$segments$yend2[i] = (sintCloneDData$segments$y[i] - 0.01)}}

sintCloneDendPoints = sintCloneDData$labels
sintCloneDendPoints$pop = sintClonePops[order.dendrogram(sintCloneDend)]
sintCloneDendPoints$depth=sintCloneDepth[order.dendrogram(sintCloneDend)]
rownames(sintCloneDendPoints) = sintCloneDendPoints$label

# Making points at the leaves to place symbols for populations
point = as.vector(NA)
for(i in 1:nrow(sintCloneDData$segments)) {
  if (sintCloneDData$segments$yend[i] == 0) {
    point[i] = sintCloneDData$segments$y[i] - 0.01
  } else {
    point[i] = NA}}

sintCloneDendPoints$y = point[!is.na(point)]

sintTechReps = c("SGM027.1", "SGM027.2", "SGM027.3", "SGM046.1", "SGM046.2", "SGM046.3", "SGM152.1", "SGM152.2", "SGM152.3", "SGM186.1", "SGM186.2", "SGM186.3", "SGM197.1", "SGM197.2", "SGM197.3", "SGM206.1", "SGM206.2", "SGM206.3")

sintCloneDendPoints$depth = factor(sintCloneDendPoints$depth)

sintCloneDendPoints$depth = factor(sintCloneDendPoints$depth,levels(sintCloneDendPoints$depth)[c(2,1)])

sintCloneDendPoints$pop = factor(sintCloneDendPoints$pop)

sintCloneDendPoints$pop = factor(sintCloneDendPoints$pop,levels(sintCloneDendPoints$pop)[c(5, 2, 1, 4, 3)])

sintCloneDendA = ggplot() +
  geom_segment(data = segment(sintCloneDData), aes(x = x, y = y, xend = xend, yend = yend2), size = 0.5) +
  geom_point(data = sintCloneDendPoints, aes(x = x, y = y, fill = pop, shape = depth), size = 4, stroke = 0.25) +
  scale_fill_manual(values = gomPal, name= "Bank")+
  scale_shape_manual(values = c(24, 25), name = "Depth Zone")+
  geom_hline(yintercept = 0.154, color = pink, lty = 5, size = 0.75) + # creating a dashed line to indicate a clonal distance threshold
  geom_text(data = subset(sintCloneDendPoints, subset = label %in% sintTechReps), aes(x = x, y = (y - .015), label = label), angle = 90) + # spacing technical replicates further from leaf
  geom_text(data = subset(sintCloneDendPoints, subset = !label %in% sintTechReps), aes(x = x, y = (y - .010), label = label), angle = 90) +
  labs(y = "Genetic distance (1 - IBS)") +
  guides(fill = guide_legend(override.aes = list(shape = 22)))+
  coord_cartesian(xlim = c(0, 220), ylim = c(0.075, 0.3), expand = FALSE) +
  theme_classic()

sintCloneDend = sintCloneDendA + theme(
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.line.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.y = element_text(size = 12, color = "black", angle = 90),
  axis.text.y = element_text(size = 10, color = "black"),
  axis.line.y = element_line(),
  axis.ticks.y = element_line(),
  panel.grid = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  plot.background = element_blank(),
  legend.key = element_blank(),
  legend.title = element_text(size = 12),
  legend.text = element_text(size = 10),
  legend.position = "bottom")

sintCloneDend

ggsave("../figures/extras/sintCloneDend.png", plot = sintCloneDend, height = 9, width = 30, units = "in", dpi = 300)

```

<br>


## *Orbicella faveolata*
***

```{r, ofav clones}
ofavCloneBams = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "O. faveolata") %>% dplyr::filter(!sample %in% c("OGM003", "OGM043", "OGM075", "OGM076", "OGM121"))

ofavCloneMa = as.matrix(read.table("../data/ofav/ofavClones.ibsMat")) # reads in IBS matrix produced by ANGSD 

dimnames(ofavCloneMa) = list(ofavCloneBams[,1],ofavCloneBams[,1])
ofavClonesHc = hclust(as.dist(ofavCloneMa),"ave")

ofavClonePops = ofavCloneBams$bank
ofavCloneDepth = ofavCloneBams$depthZone

ofavCloneDend = ofavCloneMa %>% as.dist() %>% hclust(.,"ave") %>% as.dendrogram()
ofavCloneDData = ofavCloneDend %>% dendro_data()

# Making the branches hang shorter so we can easily see clonal groups
ofavCloneDData$segments$yend2 = ofavCloneDData$segments$yend
for(i in 1:nrow(ofavCloneDData$segments)) {
  if (ofavCloneDData$segments$yend2[i] == 0) {
    ofavCloneDData$segments$yend2[i] = (ofavCloneDData$segments$y[i] - 0.01)}}

ofavCloneDendPoints = ofavCloneDData$labels
ofavCloneDendPoints$pop = ofavClonePops[order.dendrogram(ofavCloneDend)]
ofavCloneDendPoints$depth=ofavCloneDepth[order.dendrogram(ofavCloneDend)]
rownames(ofavCloneDendPoints) = ofavCloneDendPoints$label

# Making points at the leaves to place symbols for populations
point = as.vector(NA)
for(i in 1:nrow(ofavCloneDData$segments)) {
  if (ofavCloneDData$segments$yend[i] == 0) {
    point[i] = ofavCloneDData$segments$y[i] - 0.01
  } else {
    point[i] = NA}}

ofavCloneDendPoints$y = point[!is.na(point)]

ofavTechReps = c("OGM024.1", "OGM024.2", "OGM024.3", "OGM066.1", "OGM066.2", "OGM066.3", "OGM108.1", "OGM108.2", "OGM108.3")

ofavCloneDendPoints$depth = factor(ofavCloneDendPoints$depth)

ofavCloneDendPoints$depth = factor(ofavCloneDendPoints$depth,levels(ofavCloneDendPoints$depth)[c(2,1)])

ofavCloneDendPoints$pop = factor(ofavCloneDendPoints$pop)

ofavCloneDendPoints$pop = factor(ofavCloneDendPoints$pop,levels(ofavCloneDendPoints$pop)[c(5, 2, 1, 4, 3)])

ofavCloneDendA = ggplot() +
  geom_segment(data = segment(ofavCloneDData), aes(x = x, y = y, xend = xend, yend = yend2), size = 0.5) +
  geom_point(data = ofavCloneDendPoints, aes(x = x, y = y, fill = pop, shape = depth), size = 4, stroke = 0.25) +
  scale_fill_manual(values = gomPal, name= "Bank")+
  scale_shape_manual(values = c(24, 25), name = "Depth Zone")+
  geom_hline(yintercept = 0.1425, color = pink, lty = 5, size = 0.75) + # creating a dashed line to indicate a clonal distance threshold
  geom_text(data = subset(ofavCloneDendPoints, subset = label %in% ofavTechReps), aes(x = x, y = (y - .015), label = label), angle = 90) + # spacing technical replicates further from leaf
  geom_text(data = subset(ofavCloneDendPoints, subset = !label %in% ofavTechReps), aes(x = x, y = (y - .010), label = label), angle = 90) +
  labs(y = "Genetic distance (1 - IBS)") +
  guides(fill = guide_legend(override.aes = list(shape = 22)))+
  coord_cartesian(xlim = c(0,137), ylim = c(0.075, 0.3), expand = FALSE) + 
  theme_classic()

ofavCloneDend = ofavCloneDendA + theme(
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.line.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.y = element_text(size = 12, color = "black", angle = 90),
  axis.text.y = element_text(size = 10, color = "black"),
  axis.line.y = element_line(),
  axis.ticks.y = element_line(),
  panel.grid = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  plot.background = element_blank(),
  legend.key = element_blank(),
  legend.title = element_text(size = 12),
  legend.text = element_text(size = 10),
  legend.position = "bottom")

ofavCloneDend

ggsave("../figures/extras/ofavCloneDend.png", plot = ofavCloneDend, height = 9, width = 30, units = "in", dpi = 300)

```

<br>


## *Xestospongia muta*
***

```{r, xesto clones}
xestoCloneBams = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "X. muta") %>% dplyr::filter(!sample %in% c("XGM013", "XGM067", "XGM129"))

xestoCloneMa = as.matrix(read.table("../data/xesto/xestoClones.ibsMat")) # reads in IBS matrix produced by ANGSD 

dimnames(xestoCloneMa) = list(xestoCloneBams[,1],xestoCloneBams[,1])
xestoClonesHc = hclust(as.dist(xestoCloneMa),"ave")

xestoClonePops = xestoCloneBams$bank
xestoCloneDepth = xestoCloneBams$depthZone

xestoCloneDend = xestoCloneMa %>% as.dist() %>% hclust(.,"ave") %>% as.dendrogram()
xestoCloneDData = xestoCloneDend %>% dendro_data()

# Making the branches hang shorter so we can easily see clonal groups
xestoCloneDData$segments$yend2 = xestoCloneDData$segments$yend
for(i in 1:nrow(xestoCloneDData$segments)) {
  if (xestoCloneDData$segments$yend2[i] == 0) {
    xestoCloneDData$segments$yend2[i] = (xestoCloneDData$segments$y[i] - 0.01)}}

xestoCloneDendPoints = xestoCloneDData$labels
xestoCloneDendPoints$pop = xestoClonePops[order.dendrogram(xestoCloneDend)]
xestoCloneDendPoints$depth=xestoCloneDepth[order.dendrogram(xestoCloneDend)]
rownames(xestoCloneDendPoints) = xestoCloneDendPoints$label

# Making points at the leaves to place symbols for populations
point = as.vector(NA)
for(i in 1:nrow(xestoCloneDData$segments)) {
  if (xestoCloneDData$segments$yend[i] == 0) {
    point[i] = xestoCloneDData$segments$y[i] - 0.01
  } else {
    point[i] = NA}}

xestoCloneDendPoints$y = point[!is.na(point)]

xestoTechReps = c("XGM034.1", "XGM034.2", "XGM034.3", "XGM072.1", "XGM072.2", "XGM072.3", "XGM158.1", "XGM158.2", "XGM158.3", "XGM179.1", "XGM179.2", "XGM179.3", "XGM193.1", "XGM193.2", "XGM193.3", "XGM203.1", "XGM203.2", "XGM203.3")

xestoCloneDendPoints$depth = factor(xestoCloneDendPoints$depth)

xestoCloneDendPoints$depth = factor(xestoCloneDendPoints$depth,levels(xestoCloneDendPoints$depth)[c(2,1)])

xestoCloneDendPoints$pop = factor(xestoCloneDendPoints$pop)

xestoCloneDendPoints$pop = factor(xestoCloneDendPoints$pop,levels(xestoCloneDendPoints$pop)[c(5, 2, 1, 4, 3)])

xestoCloneDendA = ggplot() +
  geom_segment(data = segment(xestoCloneDData), aes(x = x, y = y, xend = xend, yend = yend2), size = 0.5) +
  geom_point(data = xestoCloneDendPoints, aes(x = x, y = y, fill = pop, shape = depth), size = 4, stroke = 0.25) +
  scale_fill_manual(values = gomPal, name= "Bank")+
  scale_shape_manual(values = c(24, 25), name = "Depth Zone")+
  geom_hline(yintercept = 0.174, color = pink, lty = 5, linewidth = 0.75) + # creating a dashed line to indicate a clonal distance threshold
  geom_text(data = subset(xestoCloneDendPoints, subset = label %in% xestoTechReps), aes(x = x, y = (y - .015), label = label), angle = 90) + # spacing technical replicates further from leaf
  geom_text(data = subset(xestoCloneDendPoints, subset = !label %in% xestoTechReps), aes(x = x, y = (y - .010), label = label), angle = 90) +
  labs(y = "Genetic distance (1 - IBS)") +
  guides(fill = guide_legend(override.aes = list(shape = 22)))+
  coord_cartesian(xlim = c(0, 217), ylim = c(0.075, 0.3), expand = FALSE) +
  theme_classic()

xestoCloneDend = xestoCloneDendA + theme(
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.line.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.y = element_text(size = 12, color = "black", angle = 90),
  axis.text.y = element_text(size = 10, color = "black"),
  axis.line.y = element_line(),
  axis.ticks.y = element_line(),
  panel.grid = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  plot.background = element_blank(),
  legend.key = element_blank(),
  legend.title = element_text(size = 12),
  legend.text = element_text(size = 10),
  legend.position = "bottom")

# xestoCloneDend

ggsave("../figures/extras/xestoCloneDend.png", plot = xestoCloneDend, height = 8, width = 35, units = "in", dpi = 300)

```
<br><br>



# *Montastraea cavernosa*
***

## Dendrogram and structure
***

```{r, mcav dend no clones, warning=FALSE, message=FALSE}
mcavBams = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "M. cavernosa") %>% dplyr::filter(!sample %in% c("MGM072.1", "MGM072.3", "MGM013.2", "MGM013.3", "MGM008.1", "MGM008.2", "MGM010", "MGM024.1", "MGM004", "MGM002.2", "MGM002.3", "MGM038.2", "MGM038.3"))# list of bams files and their populations (tech reps removed)
mcavBams$sample <- gsub("\\.[1-3]*$", "", mcavBams$sample)


mcavBams$bank = factor(mcavBams$bank)
mcavBams$bank = factor(mcavBams$bank, levels = levels(mcavBams$bank)[c(5, 2, 1, 3, 4)])

mcavBams$depthZone = factor(mcavBams$depthZone)
mcavBams$depthZone = factor(mcavBams$depthZone, levels = levels(mcavBams$depthZone)[c(2, 1)])

mcavMa = as.matrix(read.table("../data/mcav/mcavNoClones.ibsMat")) # reads in IBS matrix produced by ANGSD

dimnames(mcavMa) = list(mcavBams[,3],mcavBams[,3])

## admixture K = 2
#-------------------------------------
mcavK2ad = read.table("../data/mcav/admix/mcavK2.output") %>% dplyr::select(V6, V7) 
mcavK2ad %>% summarise(sum(V6),sum(V7))
mcavAdmixK2 = mcavBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(mcavK2ad) %>% 
  dplyr::rename("Mc1" = "V7", "Mc2" = "V6") %>% dplyr::select(order(colnames(.))) %>% dplyr::relocate(sample)

mcavAdmixK2_melt = melt(mcavAdmixK2, id = c("sample", "bank", "depth", "depthm"))

## admixture K = 3
#-------------------------------------
mcavK3ad = read.table("../data/mcav/admix/mcavK3.output") %>% dplyr::select(V6, V7, V8) 
mcavK3ad %>% summarise(sum(V6),sum(V7), sum(V8))
mcavAdmixK3 = mcavBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(mcavK3ad) %>% 
  dplyr::rename("Mc1" = "V7", "Mc2" = "V6",, "Mc3" = "V8") %>%
  dplyr::select(order(colnames(.)))
mcavAdmixK3_melt = melt(mcavAdmixK3, id = c("sample", "bank", "depth", "depthm"))


## ngsAdmix K = 4
#-------------------------------------
mcavK4ad = read.table("../data/mcav/admix/mcavK4.output") %>% dplyr::select(V6, V7, V8, V9) 
mcavK4ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9))
mcavAdmixK4 = mcavBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(mcavK4ad) %>% 
  dplyr::rename("Mc1" = "V7", "Mc2" = "V6", "Mc3" = "V8", "Mc4" = "V9") %>%
  dplyr::select(order(colnames(.)))
mcavAdmixK4_melt = melt(mcavAdmixK4, id = c("sample", "bank", "depth", "depthm"))


## admixture K = 5
#-------------------------------------
mcavK5ad = read.table("../data/mcav/admix/mcavK5.output") %>% dplyr::select(V6, V7, V8, V9, V10)
mcavK5ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9), sum(V10))
mcavAdmixK5 = mcavBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(mcavK5ad) %>% 
  dplyr::rename("Mc1" = "V7", "Mc2" = "V6", "Mc3" = "V8", "Mc4" = "V9", "Mc5" = "V10") %>% dplyr::select(order(colnames(.)))
mcavAdmixK5_melt = melt(mcavAdmixK5, id = c("sample", "bank", "depth", "depthm"))


## admixture K = 6
#-------------------------------------
mcavK6ad = read.table("../data/mcav/admix/mcavK6.output") %>% dplyr::select(V6, V7, V8, V9, V10,V11)
mcavK6ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9), sum(V10), sum(V11))
mcavAdmixK6 = mcavBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(mcavK6ad) %>% 
  dplyr::rename("Mc1" = "V7", "Mc2" = "V6", "Mc3" = "V8", "Mc4" = "V9", "Mc5" = "V10", "Mc6" = "V11") %>% dplyr::select(order(colnames(.)))
mcavAdmixK6_melt = melt(mcavAdmixK6, id = c("sample", "bank", "depth", "depthm"))

## construct figure
#-------------------------------------
# ggtree(tr) + geom_nodelab(aes(label = node), hjust = -0.5) 

{
  mcavTr = hclust(dist(mcavMa),"ave") 

  mcavP1 = ggtree(mcavTr, layout = "rectangular", size = 0.35) 

  mcavP2 = facet_plot(mcavP1, panel = "location", data=mcavAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = bank), color = "grey25", size = 0.1) +
    scale_fill_manual("Site", values = gomPal, guide = guide_legend(order = 1)) +
    new_scale_fill()

  mcavP3 = facet_plot(mcavP2, panel = "depth zone", data=mcavAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = depth), color = "grey25", size = 0.1) +
    scale_fill_manual("Depth zone", values = c("#A7E1BCFF", "#414083FF"), guide = guide_legend(order = 2)) +
    new_scale_fill()
  
  mcavP4 = facet_plot(mcavP3, panel = "depth", data=mcavAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = depthm), color = "grey25", size = 0.1) +
    scale_fill_viridis_c("Depth (m)", option = "mako", trans = "reverse", limits = c(60, 10)) +
    new_scale_fill()

  mcavP5 = facet_plot(mcavP4, panel="K = 2", data=mcavAdmixK2_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) +
    scale_fill_manual("Lineage",values = mcavKColPal[1:6])

  mcavP6 = facet_plot( mcavP5, panel="K = 3", data=mcavAdmixK3_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE)  

  mcavP7 = facet_plot(mcavP6, panel="K = 4", data=mcavAdmixK4_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE)  

  mcavP8 = facet_plot(mcavP7, panel="K = 5", data=mcavAdmixK5_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) 
    
  mcavP9 = facet_plot(mcavP8, panel="K = 6", data=mcavAdmixK6_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) +
    theme_tree(strip.text = element_blank(),
               panel.spacing = unit(.1, "line")) 
}


mcavImg = image_read("../figures/icons/mcav.png") %>% image_ggplot()

mcavStructure = facet_widths(mcavP9, widths = c(0.5, 0.025, 0.025, 0.025, 0.2, 0.1, 0.1, 0.1, 0.1)) #+ inset_element(mcavImg, 0.0, 0.9, 0.1, 1.0, align_to = "full", ignore_tag = TRUE)

mcavStructure

```
<br>

## Population structure
***

```{r}
mcavData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "M. cavernosa") %>% dplyr::filter(!sample %in% c("MGM072.1", "MGM072.3", "MGM013.2", "MGM013.3", "MGM008.1", "MGM008.2", "MGM010", "MGM024.1", "MGM004", "MGM002.2", "MGM002.3", "MGM038.2", "MGM038.3"))

mcavData$depthZone = factor(mcavData$depthZone)
mcavData$depthZone = factor(mcavData$depthZone, levels(mcavData$depthZone)[c(2,1)])

mcavData$bank = factor(mcavData$bank)
mcavData$bank = factor(mcavData$bank,levels(mcavData$bank)[c(5, 2, 1, 4, 3)])

mcavPcadmix = read.table("../data/mcav/admix/mcavK2.output") %>%dplyr::select(V6, V7)
mcavPcadmix %>% summarise(sum(V6),sum(V7)) 

mcavPcadmix = mcavData %>% cbind(mcavPcadmix) %>% rename("cluster1" = "V7", "cluster2" = "V6") %>%dplyr::select(order(colnames(.))) 

mcavPcadmixClust = mcavPcadmix %>% mutate(cluster = ifelse(cluster1 < 0.75 & cluster2 < 0.75, "NA", ifelse(cluster1 >=0.75, 1, ifelse(cluster2 >= 0.75, 2, 0))))

mcavPcadmix = mcavPcadmix %>% mutate(mcavPcadmixClust)

mcavPcadmix$cluster = as.factor(mcavPcadmix$cluster)
levels(mcavPcadmix$cluster) = c("Mc_1", "Mc_2", "Admixed")

mcavSiteLineages = mcavPcadmix %>% dplyr::select(depthZone, cluster) %>% 
group_by(depthZone) %>% count(cluster) %>% mutate(Freq = n/sum(n)) %>% apply(2, function(x) gsub("\\s+", "", x)) %>% as.data.frame()

```

### PCA
***

```{r, mcav pcangsd}
mcavCov = read.table("../data/mcav/mcavPcangsd.cov") %>% as.matrix()

mcavPcAdmix = read.table("../data/mcav/admix/mcavK2.output") %>%dplyr::select(V6, V7)
mcavPcAdmix %>% summarise(sum(V6),sum(V7)) 

mcavPcAdmix = mcavPcAdmix %>% rename("cluster1" = "V7", "cluster2" = "V6") %>% dplyr::select(order(colnames(.)))
  

mcavPcaEig = eigen(mcavCov)
mcavPcaVar = mcavPcaEig$values/sum(mcavPcaEig$values)*100
head(mcavPcaVar)

mcavPcangsd = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "M. cavernosa") %>% dplyr::filter(!sample %in% c("MGM072.1", "MGM072.3", "MGM013.2", "MGM013.3", "MGM008.1", "MGM008.2", "MGM010", "MGM024.1", "MGM004", "MGM002.2", "MGM002.3", "MGM038.2", "MGM038.3")) %>% dplyr::select("sample" = sample, "bank" = bank, "depth" = depthZone, "depthm" = depthM)

mcavPcangsd$sitedepth = as.factor(paste(mcavPcangsd$bank, mcavPcangsd$depth, sep = " "))

mcavPcangsd$sitedepth = factor(mcavPcangsd$sitedepth, levels(mcavPcangsd$sitedepth)[c(7, 6, 3, 2, 1, 4, 5)])

mcavPcangsd$bank = factor(mcavPcangsd$bank)
mcavPcangsd$bank = factor(mcavPcangsd$bank, levels(mcavPcangsd$bank)[c(5, 2, 1, 3, 4)])

mcavPcangsd$depth = factor(mcavPcangsd$depth)
mcavPcangsd$depth = factor(mcavPcangsd$depth, levels(mcavPcangsd$depth)[c(2, 1)])

mcavPcangsd$PC1 = mcavPcaEig$vectors[,1]
mcavPcangsd$PC2 = mcavPcaEig$vectors[,2]
mcavPcangsd$PC3 = mcavPcaEig$vectors[,3]
mcavPcangsd$PC4 = mcavPcaEig$vectors[,4]

mcavPcangsdClust = mcavPcAdmix %>% mutate(cluster = ifelse(cluster1 < 0.75 & cluster2 < 0.75, "NA", ifelse(cluster1 >=0.75, 1, ifelse(cluster2 >= 0.75, 2, 0))))

# pcangsdClust$clusterX = as.vector(d_clust$classification)

mcavPcangsd = mcavPcangsd %>% mutate(mcavPcangsdClust)

mcavPcangsd$cluster = as.factor(mcavPcangsd$cluster)
levels(mcavPcangsd$cluster) = c("Mc_Deep", "Mc_1", "Admixed")

mcavBamsClusters = mcavPcangsd %>% dplyr::select(sample, cluster) %>% dplyr::arrange(sample) 
mcavBamssamples = read.delim("../data/mcav/bamsNoClones", header = FALSE)
mcavBamsClusters$sample = mcavBamssamples$V1

# bamsClusters = as.data.frame(bamsClusters)

write.table(x = mcavBamsClusters, file = "../data/mcav/mcavBamsClusters", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

mcavPcangsd = merge(mcavPcangsd, aggregate(cbind(mean.x = PC1, mean.y = PC2)~sitedepth, mcavPcangsd, mean), by="sitedepth")


```

```{r}
mcavPcangsd %>% dplyr::group_by(depth,cluster) %>% dplyr::summarize(n = n())

```

<br>

#### Plot PCA
***

```{r, mcav pca plots}
mcavPcaPlot12SA = ggplot() +
  geom_hline(yintercept = 0, color = "gray90", size = 0.25) +
  geom_vline(xintercept = 0, color = "gray90", size = 0.25) +
  geom_point(data = mcavPcangsd, aes(x = PC1, y = PC2, fill = bank, shape = depth, color = bank), stroke = 0, size = 3, alpha = 0.5, show.legend = FALSE) +
  geom_point(data = mcavPcangsd, aes(x = mean.x, y = mean.y, fill = bank, shape = depth), color = "black", size = 3.75, alpha = 1, stroke = 0.25) +
  scale_shape_manual(values = c(21, 23), name = "Depth Zone") +
  scale_fill_manual(values = gomPal, name = "Site") +
  scale_color_manual(values = gomPal, name = "Site") +
  labs(x = paste0("PC 1 (", format(round(mcavPcaVar[1], 1), nsmall = 1)," %)")) +
  annotate(geom = "text", x = -0.1275, y = 0.027, label = paste0("PC 2 (", format(round(mcavPcaVar[2], 1), nsmall = 1), " %)"), size = 3.65, angle = 90) +
  guides(shape = guide_legend(override.aes = list(size = 2, stroke = 0.25, alpha = NA), order = 2, ncol = 1), fill = guide_legend(override.aes = list(shape = 22, size = 2, fill = gomPal[c(1,2,3,4,5)], alpha = NA), order = 1, ncol = 1)) +
  coord_cartesian(xlim = c(-0.1, 0.225), clip = "off") +
  theme_bw()

mcavPcaPlot12S = mcavPcaPlot12SA +
  pcaTheme +
  theme(legend.position = c(0.87, 0.76),
        axis.title.y = element_blank())

mcavPcaPlot12S


mcavPcaPlot12LA = ggplot() +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_vline(xintercept = 0, color = "gray90", size = 0.5) +
  geom_point(data = mcavPcangsd, aes(x = PC1, y = PC2, fill = cluster, shape = depth), color = "black", size = 3, alpha = 1, show.legend = TRUE) +
  scale_shape_manual(values = c(21, 23), name = "Depth Zone") +
  scale_fill_manual(values = mcavKColPal[c(1, 2, 7)], name = "Lineage") +
  labs(x = paste0("PC 1 (", format(round(mcavPcaVar[1], 1), nsmall = 1)," %)"), y = paste0("PC 2 (", format(round(mcavPcaVar[2], 1), nsmall = 1), " %)")) +
  guides(shape = "none", fill = guide_legend(override.aes = list(shape = 22, size = 2, fill = mcavKColPal[c(1, 2, 7)], alpha = NA), order = 1, ncol = 1))+
  coord_cartesian(xlim = c(-0.1, 0.225), clip = "off") +
  theme_bw()

mcavPcaPlot12L = mcavPcaPlot12LA +
  pcaTheme +
  theme(legend.position = c(0.9,0.86))

mcavPcaPlot12L

```
<br>

#### Put all plots together
***

```{r, mcav pca 1 plot}
mcavPcaPlots = mcavPcaPlot12S + mcavPcaPlot12L +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 18),
        panel.background = element_rect(fill = "white"),
        legend.spacing = unit(-5, "pt"),
        legend.key = element_blank(),
        legend.background = element_blank())

mcavPcaPlots

```

<br>

### Admixture
***

Prepare admixture outputs for plotting

```{r, mcav prep admix}
mcavAdmix = mcavPcangsd %>%dplyr::select(-PC1, -PC2, -PC3, -PC4, -cluster, -depthm, -mean.x, -mean.y)
mcavAdmix$bank = factor(mcavAdmix$bank)

mcavAdmix = arrange(mcavAdmix, bank, depth, -cluster1, -cluster2)
mcavPopCounts = mcavAdmix %>% group_by(bank, depth) %>% summarise(n = n())

mcavPopCounts

mcavPopCountList = reshape2::melt(lapply(mcavPopCounts$n,function(x){c(1:x)}))
mcavAdmix$ord = mcavPopCountList$value

mcavAdmixMelt = melt(mcavAdmix, id.vars=c("sample", "bank", "depth", "sitedepth", "ord"), variable.name="Ancestry", value.name="Fraction")

mcavAdmixMelt$Ancestry = factor(mcavAdmixMelt$Ancestry)
mcavAdmixMelt$Ancestry = factor(mcavAdmixMelt$Ancestry, levels = rev(levels(mcavAdmixMelt$Ancestry)))

mcavPopAnno = data.frame(x1 = c(0.5, 0.5, 0.5, 0.5, 0.5), x2 = c(28.5, 28.5, 28.5, 28.5, 28.5),
                     y1 = -0.1, y2 = -0.1, sample = NA, Ancestry = NA, depth = "Mesophotic", 
                     ord  = NA, Fraction = NA,
                     bank = c("WFGB", "EFGB", "Bright", "Geyer", "McGrail"))
mcavPopAnno$bank = factor(mcavPopAnno$bank)
mcavPopAnno$bank = factor(mcavPopAnno$bank, levels = levels(mcavPopAnno$bank)[c(5, 2, 1, 3, 4)])

```

<br>

#### Make admixture plots

```{r, mcav admix plots}
mcavAdmixPlotA = ggplot(data = mcavAdmixMelt, aes(x = ord, y = Fraction, fill = Ancestry, order = sample)) +
  geom_segment(data = mcavPopAnno, aes(x = x1, xend = x2, y = -.1, yend = -.1, color = bank), size = 7) +
  geom_bar(stat = "identity", position = "fill", width = 1, colour = "grey25", size = 0.2) +
  facet_grid(factor(depth) ~ bank, switch = "both", scales = "free_x") +
  geom_text(data = (mcavAdmixMelt %>% filter(depth == "Mesophotic", bank %in% c("WFGB", "EFGB", "Bright", "Geyer", "McGrail"), sample %in% c(	
"MGM116", "MGM015", "MGM001", "MGM144", "MGM062"), Ancestry == "cluster1")), x = 15.5, y = -.1, aes(label = bank), size = 4, color = "black") +
  
  scale_fill_manual(values = mcavKColPal[1:2]) +
  scale_color_manual(values = gomPal) +
  scale_x_discrete(expand = c(0.005, 0.005)) +
  scale_y_continuous(expand = c(0.001, 0.001)) +
  coord_cartesian(ylim = c(0.0, 1.0), clip = "off") +
theme_bw()
  
mcavAdmixPlot = mcavAdmixPlotA + 
  theme_bw()+
  admixTheme

mcavAdmixPlot

```

<br><br>

## Lineage demographics
***

### Lineage depth distribuiton
***

```{r, mcav depth distribution}
ticks = data.frame(y = c(seq(5,75,5)), x = 0.15, xend = 0.08)

leveneTest(lm(depthm ~ cluster, data = mcavPcangsd))

mcavDepthAov = welch_anova_test(depthm ~ cluster, data = subset(mcavPcangsd, subset = mcavPcangsd$cluster!="Admixed"))

mcavDF = mcavDepthAov$statistic[[1]]

mcavDepthPH = games_howell_test(depthm ~ cluster, data = mcavPcangsd, conf.level = 0.95) %>% as.data.frame()

mcavDepthComps = paste(mcavDepthPH$group1,"-",mcavDepthPH$group2, sep ="")
# mcavComps = c("D1-D2", "D1-D3", "D1-D4", "D1-S1", "D1-S2", "D1-A",)
mcavDepthPs = mcavDepthPH$p.adj

mcDepthMultComp = mcavDepthPs < .05
names(mcDepthMultComp) = mcavDepthComps
mcDepthLet = multcompLetters(mcDepthMultComp)
mcDepthLetters = mcDepthLet$Letters %>% as.data.frame()

  
mcavDepthLetters = data.frame(x = factor(rownames(mcDepthLetters)), y = 5,  grp = mcDepthLetters$.)

mcavLineageViolinA = ggplot(data = mcavPcangsd, aes(x = cluster, y = depthm)) +
  annotate(geom = "rect", xmin = -Inf, xmax = Inf, ymin = 30, ymax = Inf,  fill = "black", alpha = 0.10, color = NA) +
  geom_violin(aes(fill = cluster),adjust = 1, linewidth = 0, color = "black", alpha = 0.75, width = 0.9, trim = F, scale = "area") +
    geom_beeswarm(aes(fill = cluster), shape = 21, size = 2, cex = 1.5, alpha = 0.5) +
  geom_violin(adjust = 1, linewidth = 0.4, color = "black", alpha = 1, width = 0.9, trim = F, fill = NA, scale = "area") +
  geom_boxplot(width = 0.1, color = "black", fill = "white", outlier.colour = NA, linewidth = 0.6, alpha = 0.5) +
  annotate(geom = "text", x = 2.8, y = 75, label = bquote(italic("F")~" = "~.(mcavDF)*"; "~italic("p")~" < 0.001"), size = 3) +
  geom_segment(data = ticks, aes(x = x, xend = xend, y = y), color = "black") +
  annotate(geom = "text", x = mcavDepthLetters$x, y = mcavDepthLetters$y, label = mcavDepthLetters$grp) +
  scale_fill_discrete(type = mcavKColPal[c(1, 2, 7)], name = "Lineage") +
  scale_color_discrete(type = mcavKColPal[c(1, 2, 7)], name = "Lineage") +
  xlab("Lineage") +
  ylab("Depth (m)") +
  scale_y_reverse(breaks = seq(5, 90, 5)) +
  coord_cartesian(xlim = c(.75, 3.25), clip = "off") +
  theme_bw()+
  theme(axis.ticks.y = element_blank(),
        panel.background = element_rect(linewidth = 0))

mcavLineageViolin = mcavLineageViolinA + violinTheme

mcavLineageViolin

```

<br>

### Heterozygosity

```{r, mcav het data}
mcavData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "M. cavernosa") %>% dplyr::filter(!sample %in% c("MGM072.1", "MGM072.3", "MGM013.2", "MGM013.3", "MGM008.1", "MGM008.2", "MGM010", "MGM024.1", "MGM004", "MGM002.2", "MGM002.3", "MGM038.2", "MGM038.3"))

mcavHetData = read.delim("../data/mcav/mcavHet.out", header = FALSE, sep = " ") %>% mutate(sample = mcavData$sample)  %>% rename("Het" = "V2") %>% left_join(mcavPcangsd) %>% select(sample, Het, cluster)
  
mcavHet = mcavData %>% left_join(mcavHetData)

```


```{r, mcav het AOV}
mcavHetAOV = welch_anova_test(mcavHet, Het ~ cluster)
mcavHetAOV$p

mcavHetPH = games_howell_test(mcavHet, Het ~ cluster)

mcavHetPH$p.adj

mcavComps = paste(mcavHetPH$group1,"-",mcavHetPH$group2, sep ="")
mcavPs = mcavHetPH$p.adj

mcMultComp = mcavPs < .05
names(mcMultComp) = mcavComps
mcLet = multcompLetters(mcMultComp)
mcLetters = mcLet$Letters %>% as.data.frame()
  
mcavHetLetters = data.frame(x = factor(rownames(mcLetters)), y = 0.0025,  grp = mcLetters$.)

mcavFST = read.delim("../data/mcav/mcavKFst.out")

```

```{r, mcav het plots}
mcavHetPlot = ggplot(data = mcavHet, aes(x = cluster, y = Het)) +
  geom_violin(aes(fill = cluster, group = cluster), adjust = 1, linewidth = 0.4, color = "black", alpha = 1, trim = FALSE, scale = "area", width = 1) +
  geom_beeswarm(aes(fill = cluster, group = cluster),shape = 21, cex = 0.5) +
  geom_boxplot(aes(fill = cluster, group = cluster), width = 0.2, color = "black", fill = "white", outlier.colour = NA, linewidth = 0.4, alpha = 0.5) +  xlab("Lineage") +
  annotate(geom = "text", x = 2.85, y = 0, label = bquote(italic("F")~" = "~.(mcavHetAOV$statistic)*"; "~italic("p")~" < 0.001"), size = 3) +
  geom_text(data = mcavHetLetters, aes(x = x, y = y, label = grp), size = 4) +
  geom_segment(data = mcavFST, aes(y = 0.0022, x = pop1, xend = pop2)) +
  annotate(geom = "text", x = 1.5, y = 0.0023, label = bquote(italic("F")["ST"]~" = "~.(round(mcavFST$weightedFst, 3)))) +
  scale_fill_discrete(type = mcavKColPal[c(1,2,7)], name = "Lineage") +
  xlab("Lineage") +
  ylab("Heterozygosity") +
  scale_y_continuous(breaks = seq(0, 0.0025, 0.0005)) +
  coord_cartesian(expand = TRUE, xlim = c(0.85, 3.15), ylim = c(0, 0.0025)) +
  theme_bw() + 
  violinTheme

mcavHetPlot

```

<br>

### Nucleotide diversity plot

```{r, mcav pi}
mcavNpgList = list(read_tsv("../data/mcav/Mc_Deep.thetas.idx.pestPG") %>% mutate(lineage = "Mc_Deep"),
               read_tsv("../data/mcav/Mc_1.thetas.idx.pestPG") %>% mutate(lineage = "Mc_1"))


mcavPiAll = purrr::reduce(mcavNpgList, rbind) %>% 
  dplyr::group_by(lineage) %>%
  filter(tP != 0) %>%
  mutate(tPps = tP/nSites) %>%
  dplyr::summarize(tPps = mean(tPps))

mcavPiAll$Ne = (mcavPiAll$tPps)/(4*2e-8)
mcavPiAll

mcavPiAll$lineage = factor(mcavPiAll$lineage)
mcavPiAll$lineage = factor(mcavPiAll$lineage, levels = levels(mcavPiAll$lineage)[c(2,1)])



```


```{r, mcav nuvDiv plot}
mcavNuclDivPlot = ggplot(mcavPiAll, aes(x = lineage, y = tPps, fill = lineage, group = lineage)) +
  geom_bar(position = position_dodge2(preserve = "single"), stat = "identity", color = "black") +
  scale_fill_manual(values = mcavKColPal) +
  geom_text(position = position_dodge2(preserve = "single", width = 0.9), aes(y = tPps-.0015, label= scales::comma(round(Ne,0)), hjust = 0, fontface = "bold"), angle = 90, color = "black", ) +
  labs(x = "Lineage", y = bquote("Nucleotide diversity ("*pi*")")) + 
  theme_bw() +
  violinTheme +
  theme(axis.text.x = ggtext::element_markdown())
  
mcavNuclDivPlot

```

<br>

## Symbiodiniaceae
***

```{r, mcav sym}
mcavData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "M. cavernosa") %>% dplyr::filter(!sample %in% c("MGM072.1", "MGM072.3", "MGM013.2", "MGM013.3", "MGM008.1", "MGM008.2", "MGM010", "MGM024.1", "MGM004", "MGM002.2", "MGM002.3", "MGM038.2", "MGM038.3")) %>% dplyr::select("sample" = sample, "site" = bank, "depth" = depthZone)

mcavZoox = read.delim("../data/mcav/mcavZooxReads", header = FALSE, check.names = FALSE)

head(mcavZoox)

# Reconstruct read mapping output into dataframe usable for analysis
mcavZoox$V2[is.na(mcavZoox$V2)] <- as.character(mcavZoox$V1[is.na(mcavZoox$V2)])
mcavZoox$V1 = gsub(pattern = "MGM*.", "chr", mcavZoox$V1)
mcavZoox$V2 = gsub(".trim.*", "", mcavZoox$V2)
mcavZoox = mcavZoox %>% filter(mcavZoox$V1 != "*")
mcavZooxLst = split(mcavZoox$V2, as.integer(gl(length(mcavZoox$V2), 20, length(mcavZoox$V2))))

mcavZooxMaps = NULL

for(i in mcavZooxLst){
  mcavZooxMaps = rbind(mcavZooxMaps, data.frame(t(i)))
}

# remove tech reps
mcavZooxMaps = mcavZooxMaps %>% dplyr::filter(!X1 %in% c("MGM072.1", "MGM072.3", "MGM013.2", "MGM013.3", "MGM008.1", "MGM008.2", "MGM010", "MGM024.1", "MGM004", "MGM002.2", "MGM002.3", "MGM038.2", "MGM038.3"))


colnames(mcavZooxMaps) = c("sample",mcavZoox$V1[c(2:20)])

# convert characters to numeric
str(mcavZooxMaps)

for(i in c(2:20)){
  mcavZooxMaps[,i] = as.numeric(mcavZooxMaps[,i])
  }

str(mcavZooxMaps)

# collapse fake chromosomes into representative genera
mcavZooxMaps$Symbiodinium = rowSums(mcavZooxMaps[2:6])
mcavZooxMaps$Breviolum = rowSums(mcavZooxMaps[7:10])
mcavZooxMaps$Cladocopium = rowSums(mcavZooxMaps[11:16])
mcavZooxMaps$Durusdinium = rowSums(mcavZooxMaps[17:20])

# keep genera totals and turn into proportions for barplot
mcavZooxMaps = mcavZooxMaps[,c(1, 21:24)]

mcavZooxProp = mcavZooxMaps
mcavZooxProp$sum = apply(mcavZooxProp[, c(2:length(mcavZooxProp[1,]))], 1, function(x) {
sum(x, na.rm = T)
})
mcavZooxProp = cbind(mcavZooxProp$sample, (mcavZooxProp[, c(2:(ncol(mcavZooxProp)-1))]
/ mcavZooxProp$sum))

colnames(mcavZooxProp)[1] = "sample"

head(mcavZooxProp)

# Check that all samples total to 1
apply(mcavZooxProp[, c(2:(ncol(mcavZooxProp)))], 1, function(x) {
sum(x, na.rm = T)
})

# add sample metadata to proportions
mcavSym = mcavData %>% left_join(mcavZooxProp)

mcavAdmixOrd = mcavAdmix %>% dplyr::select(sample, ord) 
mcavPcangsdITS = mcavPcangsd %>% dplyr::select(sample, cluster)

mcavSym = mcavSym %>% left_join(mcavAdmixOrd) %>% relocate(ord,.after = depth) %>% left_join(mcavPcangsdITS) %>% relocate(cluster, .after = depth)
```

<br>

```{r, mcav sym permanova}
mcavZooxPermData = mcavData %>% left_join(mcavZooxMaps) %>% left_join(mcavPcangsdITS) %>% relocate(cluster, .after = depth)

mcavZooxPermData$depth = factor(mcavZooxPermData$depth)
mcavZooxPermData$depth = factor(mcavZooxPermData$depth, levels = levels(mcavZooxPermData$depth)[c(2, 1)])

mcavZooxPermData$site = factor(mcavZooxPermData$site)
mcavZooxPermData$site = factor(mcavZooxPermData$site, levels = levels(mcavZooxPermData$site)[c(5, 2, 1, 3, 4)])

head(mcavZooxPermData)


set.seed(694)


mcavSymPerm = adonis2(decostand(mcavZooxPermData[ c(5:((ncol(mcavZooxPermData))))], "normalize") ~ depth * site, data = mcavZooxPermData, permutations = 9999, method = "bray", sqrt.dist = TRUE, by = "terms")

mcavSymPerm

mcavSymPermPW = pairwise.adonis2(decostand(mcavZooxPermData[ c(5:((ncol(mcavZooxPermData))))], "normalize") ~ site, data = mcavZooxPermData, nperm = 9999, method = "bray", sqrt.dist = TRUE, by = "terms")

mcavSymPermPW

```


M. cavernosa lineage  vs Symbiodiniaceae B-C distance

```{r, mcav procrustes its2 snps}
mcavZooxDist = mcavZooxPermData[ c(5:ncol(mcavZooxPermData))] %>% decostand("normalize") %>% vegdist(method = "bray")

mcavZooxPCoA = cmdscale(mcavZooxDist, eig = TRUE, x.ret = TRUE)

mcavAdmixDist = mcavAdmixK2[c(5:ncol(mcavAdmixK2))] %>% vegdist(method = "euclidean")

mcavAdmixPCoA = cmdscale(mcavAdmixDist, eig = TRUE, x.ret = TRUE)

set.seed(981) 
mcavProcrustes = protest(X = mcavAdmixPCoA, Y = mcavZooxPCoA , permutations = 9999)

mcavProcrustes


```

<br>

```{r, mcav sym 2}
mcavSym$depth = factor(mcavSym$depth)
mcavSym$depth = factor(mcavSym$depth, levels = levels(mcavSym$depth)[c(2, 1)])

mcavSym$site = factor(mcavSym$site)
mcavSym$site = factor(mcavSym$site, levels = levels(mcavSym$site)[c(5, 2, 1, 3, 4)])

#turn into melted dataframe with otustack() and remove "summ" rows
mcavGssSym = otuStack(mcavSym, count.columns = c(6:length(mcavSym[1, ])),
 condition.columns = c(1:5)) %>% filter(otu != "summ") %>% droplevels()

#check that levels are correct/ordered
levels(mcavGssSym$otu)
levels(mcavGssSym$depth)
levels(mcavGssSym$site)

```

<br>

#### Symbiodiniaceae genera barplot
***

```{r, mcav sym plot}
mcavZooxAnno = data.frame(x1 = c(0.5, 0.5, 0.5, 0.5, 0.5), x2 = c(28.5, 28.5, 28.5, 28.5, 28.5),
y1 = -0.22, y2 = -0.22, site = c("WFGB", "EFGB", "Bright", "Geyer", "McGrail"))

mcavZooxAnno$site = factor(mcavZooxAnno$site)
mcavZooxAnno$site = factor(mcavZooxAnno$site, levels = levels(mcavZooxAnno$site)[c(5, 2, 1, 3, 4)])

mcavGssSymPlot = mcavGssSym %>% left_join(mcavZooxAnno, by = "site") %>% left_join(mcavPcangsdITS)
mcavGssSymPlot$ord = as.numeric(mcavGssSymPlot$ord)

mcavZooxPlotA = ggplot(data = mcavGssSymPlot, aes(x = ord, y = count, fill = otu, order = ord)) +

geom_point(aes(x=1, y=0.5, fill = otu), shape = 22, size = 0) +
  
geom_bar(stat = "identity", position = "stack", colour = "grey25", width = 1, size = 0.2, show.legend = FALSE) +
  
xlab("Population") +
  
scale_fill_manual(values = colPalZoox, name = "Symbiodiniaceae \ngenus") +
  
geom_segment(data = (mcavGssSymPlot %>% filter(depth == "Mesophotic")), aes(x = x1, xend = x2, y = y1, yend = y2, color = site), size = 7) +

scale_color_manual(values = gomPal, guide = "none") +
  
ggnewscale::new_scale_color() +

geom_segment(data = mcavGssSymPlot, aes(x = ord-0.5, xend = ord+0.5, color = cluster), y = -.07, yend = -.07, linewidth = 4) +
  
scale_color_manual(values = mcavKColPal[c(1, 2, 7)], name = "Lineage") +

coord_cartesian(ylim = c(0, 1), xlim = c(0.5, 28.5), clip = "off") +

scale_x_discrete(expand = c(0.005, 0.005)) +
scale_y_continuous(expand = c(0.001, 0.001)) +
facet_grid(factor(depth) ~ site, drop = TRUE, scales = "free", switch = "both", space = "free") +

geom_text(data = subset(mcavGssSymPlot, subset = otu == "Symbiodinium") %>% filter(sample %in% c("MGM116", "MGM015", "MGM001", "MGM144", "MGM062")), x = 15.5, y = -.205, aes(label = site), size = 4.4, color = "#000000") +
guides(fill = guide_legend(override.aes = list(size = 4), ncol = 1, order = 1), color = guide_legend(override.aes = list(size = .1), ncol = 1)) +
  
theme_bw()

mcavZooxPlot = mcavZooxPlotA + zooxTheme

mcavZooxPlot

```

## All *M. cavernosa* plots
***

```{r, all mcav plots}
mcavPlot1 = mcavStructure + inset_element(mcavImg, -.1,.7,.3,1, align_to = "full", ignore_tag = TRUE)
mcavPlot2 = (mcavPcaPlots) + plot_layout(guides = "collect")
mcavPlot3 = (mcavLineageViolin + theme(axis.text.y = element_text(margin = ggplot2::margin(r = -15, unit = "pt")))) + mcavHetPlot + mcavNuclDivPlot
mcavPlot4 = mcavAdmixPlot + mcavZooxPlot



# theme(axis.title.y = element_text(margin = ggplot2::margin(r = -20, unit = "pt"))


mcavPlots =  mcavPlot1 + mcavPlot2 + mcavPlot3 + mcavPlot4+
  plot_layout(heights = c(2,1,1,1)) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 18),
        legend.spacing = unit(-5, "pt"),
        legend.key = element_blank(),
        legend.background = element_blank())

ggsave("../figures/figure2.png", plot = mcavPlots, height = 14, width = 12, units = "in", dpi = 300)

ggsave("../figures/figure2.svg", plot = mcavPlots, height = 14, width = 12, units = "in", dpi = 300)

```

<br><br>



# *Stephanocoenia intersepta*
***

## Dendrogram and structure
***

```{r, sint dend no clones, warning=FALSE, message=FALSE}
sintBams = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "S. intersepta") %>% dplyr::filter(!sample %in% c("SGM007", "SGM010", "SGM013", "SGM027.2", "SGM027.3", "SGM046.1", "SGM046.3", "SGM063", "SGM152.2", "SGM152.3", "SGM169", "SGM177", "SGM179", "SGM186.1", "SGM186.3", "SGM197.2", "SGM197.3", "SGM206.2", "SGM206.3"))

sintBams$sample <- gsub("\\.[1-3]*$", "", sintBams$sample)
sintBams = sintBams %>% dplyr::relocate(sample)

sintBams$bank = factor(sintBams$bank)
sintBams$bank = factor(sintBams$bank, levels = levels(sintBams$bank)[c(5, 2, 1, 3, 4)])

sintBams$depthZone = factor(sintBams$depthZone)
sintBams$depthZone = factor(sintBams$depthZone, levels = levels(sintBams$depthZone)[c(2, 1)])


sintMa = as.matrix(read.table("../data/sint/sintNoClones.ibsMat")) # reads in IBS matrix produced by ANGSD


dimnames(sintMa) = list(sintBams[,3],sintBams[,3])


## admixture K = 2
#-------------------------------------
sintK2ad = read.table("../data/sint/admix/sintK2.output") %>% dplyr::select(V6, V7) 
sintK2ad %>% summarise(sum(V6),sum(V7))
sintAdmixK2 = sintBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(sintK2ad) %>% 
  dplyr::rename("Si1" = "V6", "Si2" = "V7")
sintAdmixK2_melt = melt(sintAdmixK2, id = c("sample", "bank", "depth", "depthm"))

## admixture K = 3
#-------------------------------------
sintK3ad = read.table("../data/sint/admix/sintK3.output") %>% dplyr::select(V6, V7, V8) 
sintK3ad %>% summarise(sum(V6),sum(V7), sum(V8))
sintAdmixK3 = sintBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(sintK3ad) %>% 
  dplyr::rename("Si1" = "V6", "Si2" = "V7", "Si3" = "V8") %>%
  dplyr::select(order(colnames(.)))
sintAdmixK3_melt = melt(sintAdmixK3, id = c("sample", "bank", "depth", "depthm"))


## ngsAdmix K = 4
#-------------------------------------
sintK4ad = read.table("../data/sint/admix/sintK4.output") %>% dplyr::select(V6, V7, V8, V9) 
sintK4ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9))
sintAdmixK4 = sintBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(sintK4ad) %>% 
  dplyr::rename("Si1" = "V6", "Si2" = "V7", "Si3" = "V8", "Si4" = "V9") %>%
  dplyr::select(order(colnames(.)))
sintAdmixK4_melt = melt(sintAdmixK4, id = c("sample", "bank", "depth", "depthm"))


## admixture K = 5
#-------------------------------------
sintK5ad = read.table("../data/sint/admix/sintK5.output") %>% dplyr::select(V6, V7, V8, V9, V10)
sintK5ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9), sum(V10))
sintAdmixK5 = sintBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(sintK5ad) %>% 
  dplyr::rename("Si1" = "V6", "Si2" = "V7", "Si3" = "V8", "Si4" = "V9", "Si5" = "V10") #%>%
sintAdmixK5_melt = melt(sintAdmixK5, id = c("sample", "bank", "depth", "depthm"))


## admixture K = 6
#-------------------------------------
sintK6ad = read.table("../data/sint/admix/sintK6.output") %>% dplyr::select(V6, V7, V8, V9, V10,V11)
sintK6ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9), sum(V10), sum(V11))
sintAdmixK6 = sintBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(sintK6ad) %>% 
  dplyr::rename("Si1" = "V6", "Si2" = "V7", "Si3" = "V8", "Si4" = "V9", "Si5" = "V10", "Si6" = "V11") #%>%
sintAdmixK6_melt = melt(sintAdmixK6, id = c("sample", "bank", "depth", "depthm"))

## construct figure
#-------------------------------------
# ggtree(tr) + geom_nodelab(aes(label = node), hjust = -0.5) 

{
  sintTr = hclust(dist(sintMa),"ave") 

  sintP1 = ggtree(sintTr, layout = "rectangular", size = 0.35) 

  sintP2 = facet_plot(sintP1, panel = "location", data=sintAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = bank), color = "grey25", size = 0.1) + 
    scale_fill_manual("Site", values = gomPal, guide = guide_legend(order = 1)) + 
    new_scale_fill() 

  sintP3 = facet_plot(sintP2, panel = "depth zone", data=sintAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = depth), color = "grey25", size = 0.1) +
    scale_fill_manual("Depth zone", values = c("#A7E1BCFF", "#414083FF"), guide = guide_legend(order = 2)) +
    new_scale_fill()
  
  sintP4 = facet_plot(sintP3, panel = "depth", data=sintAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = depthm), color = "grey25", size = 0.1) + 
    scale_fill_viridis_c("Depth (m)", option = "mako", trans = "reverse", limits = c(60, 10)) +
    new_scale_fill()

  sintP5 = facet_plot(sintP4, panel="K = 2", data=sintAdmixK2_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) +
    scale_fill_manual("Lineage",values = sintKColPal[c(1,5,2,3,4,6)])
  
  sintP6 = facet_plot( sintP5, panel="K = 3", data=sintAdmixK3_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE)  

  sintP7 = facet_plot(sintP6, panel="K = 4", data=sintAdmixK4_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE)  

  sintP8 = facet_plot(sintP7, panel="K = 5", data=sintAdmixK5_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) 
    
  sintP9 = facet_plot(sintP8, panel="K = 6", data=sintAdmixK6_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) +
    theme_tree(strip.text = element_blank(),
               panel.spacing = unit(.1, "line")) 
}


sintImg = image_read("../figures/icons/sint.png") %>% image_ggplot()

sintStructure = facet_widths(sintP9, widths = c(0.5, 0.025, 0.025, 0.025, 0.2, 0.1, 0.1, 0.1, 0.1))
sintStructure

```

<br>

## Population structure
***

```{r, sint pca data}
sintData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "S. intersepta") %>% dplyr::filter(!sample %in% c("SGM007", "SGM010", "SGM013", "SGM027.2", "SGM027.3", "SGM046.1", "SGM046.3", "SGM063", "SGM152.2", "SGM152.3", "SGM169", "SGM177", "SGM179", "SGM186.1", "SGM186.3", "SGM197.2", "SGM197.3", "SGM206.2", "SGM206.3"))

sintData$depthZone = factor(sintData$depthZone)
sintData$depthZone = factor(sintData$depthZone, levels(sintData$depthZone)[c(2,1)])

sintData$bank = factor(sintData$bank)
sintData$bank = factor(sintData$bank,levels(sintData$bank)[c(5, 2, 1, 4, 3)])

sintPcadmix = read.table("../data/sint/admix/sintK2.output") 
sintPcadmix %>% summarise(sum(V6),sum(V7)) 

sintPcadmix = sintData %>% cbind(sintPcadmix) %>% rename("cluster1" = "V6", "cluster2" = "V7") %>%dplyr::select(order(colnames(.))) 

sintPcadmixClust = sintPcadmix %>% mutate(cluster = ifelse(cluster1 < 0.75 & cluster2 < 0.75, "NA", ifelse(cluster1 >=0.75, 1, ifelse(cluster2 >= 0.75, 2, 0))))

sintPcadmix = sintPcadmix %>% mutate(sintPcadmixClust)

sintPcadmix$cluster = as.factor(sintPcadmix$cluster)
levels(sintPcadmix$cluster) = c("Si_1", "Si_2", "Admixed")

sintSiteLineages = sintPcadmix %>% dplyr::select(depthZone, cluster) %>% 
group_by(depthZone) %>% count(cluster) %>% mutate(Freq = n/sum(n)) %>% apply(2, function(x) gsub("\\s+", "", x)) %>% as.data.frame()

sintSiteLineages

```

<br>

### PCA
***

```{r, sint pca}
sintCov = read.table("../data/sint/sintPcangsd.cov") %>% as.matrix()

sintPcAdmix = read.table("../data/sint/admix/sintK2.output") %>%dplyr::select(V6, V7)
sintPcAdmix %>% summarise(sum(V6),sum(V7)) 

sintPcAdmix = sintPcAdmix %>% rename("cluster1" = "V6", "cluster2" = "V7") %>% dplyr::select(order(colnames(.)))
  

sintPcaEig = eigen(sintCov)
sintPcaVar = sintPcaEig$values/sum(sintPcaEig$values)*100
head(sintPcaVar)

sintPcangsd = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "S. intersepta") %>% dplyr::filter(!sample %in% c("SGM007", "SGM010", "SGM013", "SGM027.2", "SGM027.3", "SGM046.1", "SGM046.3", "SGM063", "SGM152.2", "SGM152.3", "SGM169", "SGM177", "SGM179", "SGM186.1", "SGM186.3", "SGM197.2", "SGM197.3", "SGM206.2", "SGM206.3")) %>% dplyr::select("sample" = sample, "bank" = bank, "depth" = depthZone, "depthm" = depthM)

sintPcangsd$sitedepth = as.factor(paste(sintPcangsd$bank, sintPcangsd$depth, sep = " "))

sintPcangsd$sitedepth = factor(sintPcangsd$sitedepth, levels(sintPcangsd$sitedepth)[c(7, 6, 3, 2, 1, 4, 5)])

sintPcangsd$bank = factor(sintPcangsd$bank)
sintPcangsd$bank = factor(sintPcangsd$bank, levels(sintPcangsd$bank)[c(5, 2, 1, 3, 4)])

sintPcangsd$depth = factor(sintPcangsd$depth)
sintPcangsd$depth = factor(sintPcangsd$depth, levels(sintPcangsd$depth)[c(2, 1)])

sintPcangsd$PC1 = sintPcaEig$vectors[,1]
sintPcangsd$PC2 = sintPcaEig$vectors[,2]
sintPcangsd$PC3 = sintPcaEig$vectors[,3]
sintPcangsd$PC4 = sintPcaEig$vectors[,4]

sintPcangsdClust = sintPcAdmix %>% mutate(cluster = ifelse(cluster1 < 0.75 & cluster2 < 0.75, "NA", ifelse(cluster1 >=0.75, 1, ifelse(cluster2 >= 0.75, 2, 0))))

# pcangsdClust$clusterX = as.vector(d_clust$classification)

sintPcangsd = sintPcangsd %>% mutate(sintPcangsdClust)

sintPcangsd$cluster = as.factor(sintPcangsd$cluster)
levels(sintPcangsd$cluster) = c("Si_1", "Si_2", "Admixed")

sintBamsClusters = sintPcangsd %>% dplyr::select(sample, cluster) %>% dplyr::arrange(sample) 
sintBamssamples = read.delim("../data/sint/bamsNoClones", header = FALSE)
sintBamsClusters$sample = sintBamssamples$V1

# bamsClusters = as.data.frame(bamsClusters)

write.table(x = sintBamsClusters, file = "../data/sint/sintBamsClusters", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

sintPcangsd = merge(sintPcangsd, aggregate(cbind(mean.x = PC1, mean.y = PC2)~sitedepth, sintPcangsd, mean), by="sitedepth")


```

```{r}
sintPcangsd %>% dplyr::group_by(depth,cluster) %>% dplyr::summarize(n = n())

```

<br>

#### Plot PCA
***

```{r, sint pca plots}
sintPcaPlot12SA = ggplot() +
  geom_hline(yintercept = 0, color = "gray90", size = 0.25) +
  geom_vline(xintercept = 0, color = "gray90", size = 0.25) +
  geom_point(data = sintPcangsd, aes(x = PC1, y = PC2, fill = bank, shape = depth, color = bank), stroke = 0, size = 3, alpha = 0.5, show.legend = FALSE) +
  geom_point(data = sintPcangsd, aes(x = mean.x, y = mean.y, fill = bank, shape = depth), color = "black", size = 3.75, alpha = 1, stroke = 0.25) +
  scale_shape_manual(values = c(21, 23), name = "Depth Zone") +
  scale_fill_manual(values = gomPal, name = "Site") +
  scale_color_manual(values = gomPal, name = "Site") +
  labs(x = paste0("PC 1 (", format(round(sintPcaVar[1], 1), nsmall = 1)," %)")) +
  annotate(geom = "text", x = -0.24, y = -0.033, label = paste0("PC 2 (", format(round(sintPcaVar[2], 1), nsmall = 1), " %)"), size = 3.65, angle = 90) +
  guides(shape = guide_legend(override.aes = list(size = 4, stroke = 0.25, alpha = NA), order = 2, ncol = 1), fill = guide_legend(override.aes = list(shape = 22, size = 4, fill = gomPal[c(1,2,3,4,5)], alpha = NA), order = 1, ncol = 1)) +
  coord_cartesian(xlim = c(-.2,.25), clip = "off") +
  theme_bw()

sintPcaPlot12S = sintPcaPlot12SA +
  pcaTheme +
  theme(legend.position = c(0.87, 0.76),
        axis.title.y = element_blank())

sintPcaPlot12S


sintPcaPlot12LA = ggplot() +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_vline(xintercept = 0, color = "gray90", size = 0.5) +
  geom_point(data = sintPcangsd, aes(x = PC1, y = PC2, fill = cluster, shape = depth), color = "black", size = 3, alpha = 1, show.legend = TRUE) +
  scale_shape_manual(values = c(21, 23), name = "Depth Zone") +
  scale_fill_manual(values = sintKColPal[c(1,2,7)], name = "Lineage") +
  labs(x = paste0("PC 1 (", format(round(sintPcaVar[1], 1), nsmall = 1)," %)"), y = paste0("PC 2 (", format(round(sintPcaVar[2], 1), nsmall = 1), " %)")) +
  guides(shape = "none", fill = guide_legend(override.aes = list(shape = 22, size = 4, fill = sintKColPal[c(1,2,7)], alpha = NA), order = 1, ncol = 1))+
  coord_cartesian(xlim = c(-.2,.25), clip = "off") +
  theme_bw()

sintPcaPlot12L = sintPcaPlot12LA +
  pcaTheme +
  theme(legend.position = c(0.9,0.86))


```

Put all plots together

```{r, pca plots combined}
sintPcaPlots = (sintPcaPlot12S | sintPcaPlot12L) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 18),
        panel.background = element_rect(fill = "white"),
        legend.spacing = unit(-5, "pt"),
        legend.key = element_blank(),
        legend.background = element_blank())

sintPcaPlots

```

<br>

### Admixture
***

#### Prepare admixture outputs for plotting

```{r, sint prep admix}
sintAdmix = sintPcangsd %>%dplyr::select(-PC1, -PC2, -PC3, -PC4, -cluster, -depthm, -mean.x, -mean.y)
sintAdmix$bank = factor(sintAdmix$bank)

sintAdmix = arrange(sintAdmix, bank, depth, -cluster1, -cluster2)
sintPopCounts = sintAdmix %>% group_by(bank, depth) %>% summarise(n = n())

sintPopCounts

sintPopCountList = reshape2::melt(lapply(sintPopCounts$n,function(x){c(1:x)}))
sintAdmix$ord = sintPopCountList$value

sintAdmixMelt = melt(sintAdmix, id.vars=c("sample", "bank", "depth", "sitedepth", "ord"), variable.name="Ancestry", value.name="Fraction")

sintAdmixMelt$Ancestry = factor(sintAdmixMelt$Ancestry)
sintAdmixMelt$Ancestry = factor(sintAdmixMelt$Ancestry, levels = rev(levels(sintAdmixMelt$Ancestry)))

sintPopAnno = data.frame(x1 = c(0.5, 0.5, 0.5, 0.5, 0.5), x2 = c(30.5, 30.5, 30.5, 30.5, 30.5),
                     y1 = -0.1, y2 = -0.1, sample = NA, Ancestry = NA, depth = "Mesophotic", 
                     ord  = NA, Fraction = NA,
                     bank = c("WFGB", "EFGB", "Bright", "Geyer", "McGrail"))
sintPopAnno$bank = factor(sintPopAnno$bank)
sintPopAnno$bank = factor(sintPopAnno$bank, levels = levels(sintPopAnno$bank)[c(5, 2, 1, 3, 4)])

```

<br>

Make admixture plots

```{r, sint admix plots}
sintAdmixPlotA = ggplot(data = sintAdmixMelt, aes(x = ord, y = Fraction, fill = Ancestry, order = sample)) +
  geom_segment(data = sintPopAnno, aes(x = x1, xend = x2, y = -.1, yend = -.1, color = bank), size = 7) +
  geom_bar(stat = "identity", position = "fill", width = 1, colour = "grey25", size = 0.2) +
  facet_grid(factor(depth) ~ bank, switch = "both", scales = "free_x") +
  geom_text(data = (sintAdmixMelt %>% filter(depth == "Mesophotic", bank %in% c("WFGB", "EFGB", "Bright", "Geyer", "McGrail"), sample %in% c("SGM075", "SGM127", "SGM058", "SGM208", "SGM020"), Ancestry == "cluster1")), x = 15.5, y = -.1, aes(label = bank), size = 4, color = "black") +
  
  scale_fill_manual(values = sintKColPal) +
  scale_color_manual(values = gomPal) +
  scale_x_discrete(expand = c(0.005, 0.005)) +
  scale_y_continuous(expand = c(0.001, 0.001)) +
  coord_cartesian(ylim = c(0.0, 1.0), clip = "off") +
theme_bw()
  
sintAdmixPlot = sintAdmixPlotA + 
  theme_bw()+
  admixTheme

sintAdmixPlot

```

<br>

## Lineage demographics

### Lineage depth distribuiton

```{r, sint depth distribution}
leveneTest(lm(depthm ~ cluster, data = sintPcangsd))

sintDepthAov = welch_anova_test(depthm ~ cluster, data = subset(sintPcangsd, subset = sintPcangsd$cluster!="Admixed"))

sintDF = sintDepthAov$statistic[[1]]

sintDepthPH = games_howell_test(depthm ~ cluster, data = sintPcangsd, conf.level = 0.95) %>% as.data.frame()

sintDepthComps = paste(sintDepthPH$group1,"-",sintDepthPH$group2, sep ="")
# sintComps = c("D1-D2", "D1-D3", "D1-D4", "D1-S1", "D1-S2", "D1-A",)
sintDepthPs = sintDepthPH$p.adj

siDepthMultComp = sintDepthPs < .05
names(siDepthMultComp) = sintDepthComps
siDepthLet = multcompLetters(siDepthMultComp)
siDepthLetters = siDepthLet$Letters %>% as.data.frame()

  
sintDepthLetters = data.frame(x = factor(rownames(siDepthLetters)), y = 5,  grp = siDepthLetters$.)


sintLineageViolinA = ggplot(data = sintPcangsd, aes(x = cluster, y = depthm)) +
  annotate(geom = "rect", xmin = -Inf, xmax = Inf, ymin = 30, ymax = Inf,  fill = "black", alpha = 0.10, color = NA) +
  geom_violin(aes(fill = cluster),adjust = 1, linewidth = 0, color = "black", alpha = 0.75, width = 0.9, trim = F, scale = "area") +
    geom_beeswarm(aes(fill = cluster), shape = 21, size = 2, cex = 1.5, alpha = 0.5) +
  geom_violin(adjust = 1, linewidth = 0.4, color = "black", alpha = 1, width = 0.9, trim = F, fill = NA, scale = "area") +
  geom_boxplot(width = 0.1, color = "black", fill = "white", outlier.colour = NA, linewidth = 0.6, alpha = 0.5) +
  geom_segment(data = ticks[c(1:13),], aes(x = x, xend = xend, y = y), color = "black") +
  annotate(geom = "text", x = 2.85, y = 65, label = bquote(italic("F")~" = "~.(sintDF)*"; "~italic("p")~" = 0.131"), size = 3) +
  annotate(geom = "text", x = sintDepthLetters$x, y = (sintDepthLetters$y - 4), label = sintDepthLetters$grp) +
  scale_fill_discrete(type = sintKColPal[c(1,2,7)], name = "Lineage") +
  scale_color_discrete(type = sintKColPal[c(1,2,7)], name = "Lineage") +
  xlab("Lineage") +
  ylab("Depth (m)") +
  scale_y_reverse(breaks = seq(5, 90, 5)) +
  coord_cartesian(xlim = c(0.75, 3.25),clip = "off") +
  theme_bw()

sintLineageViolin = sintLineageViolinA + violinTheme 

sintLineageViolin

```

<br>

### Heterozygosity

```{r, sint het data}
sintData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "S. intersepta") %>% dplyr::filter(!sample %in% c("SGM007", "SGM010", "SGM013", "SGM027.2", "SGM027.3", "SGM046.1", "SGM046.3", "SGM063", "SGM152.2", "SGM152.3", "SGM169", "SGM177", "SGM179", "SGM186.1", "SGM186.3", "SGM197.2", "SGM197.3", "SGM206.2", "SGM206.3"))  

sintHetData = read.delim("../data/sint/sintHet.out", header = FALSE, sep = " ") %>% mutate(sample = sintData$sample)  %>% rename("Het" = "V2") %>% left_join(sintPcangsd) %>% select(sample, Het, cluster)
  
sintHet = sintData %>% left_join(sintHetData)

```


```{r, sint het AOV}
sintHetAOV = welch_anova_test(sintHet, Het ~ cluster)
sintHetAOV$p

sintHetPH = games_howell_test(sintHet, Het ~ cluster)

sintHetPH$p.adj

sintComps = paste(sintHetPH$group1,"-",sintHetPH$group2, sep ="")
sintPs = sintHetPH$p.adj

siMultComp = sintPs < .05
names(siMultComp) = sintComps
siLet = multcompLetters(siMultComp)
siLetters = siLet$Letters %>% as.data.frame()
  
sintHetLetters = data.frame(x = factor(rownames(siLetters)), y = 0.0025,  grp = siLetters$.)

sintFST = read.delim("../data/sint/sintKFst.out")

```

```{r, het plots}
sintHetPlot = ggplot(data = sintHet, aes(x = cluster, y = Het)) +
  geom_violin(aes(fill = cluster, group = cluster), adjust = 1, linewidth = 0.4, color = "black", alpha = 1, trim = TRUE, scale = "area", width = 1.2) +
  geom_beeswarm(aes(fill = cluster), shape = 21, cex = 0.5) +
  geom_boxplot(aes(fill = cluster, group = cluster), width = 0.2, color = "black", fill = "white", outlier.colour = NA, linewidth = 0.4, alpha = 0.5) +  xlab("Lineage") +
  geom_text(data = sintHetLetters, aes(x = x, y = y, label = grp), size = 4) +
   annotate(geom = "text", x = 2.9, y = 0, label = bquote(italic("F")~" = "~.(sintHetAOV$statistic)*"; "~italic("p")~" < 0.001"), size = 3) +
  geom_segment(data = sintFST, aes(y = 0.0022, x = pop1, xend = pop2)) +
  annotate(geom = "text", x = 1.5, y = 0.0023, label = bquote(italic("F")["ST"]~" = "~.(round(sintFST$weightedFst, 3)))) +
  scale_fill_discrete(type = sintKColPal[c(1,2,7)], name = "Lineage") +
  xlab("Lineage") +
  ylab("Heterozygosity") +
  scale_y_continuous(breaks = seq(0, 0.0025, 0.0005)) +
  coord_cartesian(expand = TRUE, xlim = c(0.85, 3.15), ylim = c(0, 0.0025)) +
  theme_bw() + 
  violinTheme +
  theme(panel.border = element_rect(linewidth = 1))

sintHetPlot

```

<br>

### Nucleotide diversity plot

```{r, sint pi}
sintNpgList = list(read_tsv("../data/sint/Si_Deep.thetas.idx.pestPG") %>% mutate(lineage = "Si_1"),
               read_tsv("../data/sint/Si_1.thetas.idx.pestPG") %>% mutate(lineage = "Si_2"))


sintPiAll = purrr::reduce(sintNpgList, rbind) %>% 
  dplyr::group_by(lineage) %>%
  filter(tP != 0) %>%
  mutate(tPps = tP/nSites) %>%
  dplyr::summarize(tPps = mean(tPps))

sintPiAll$Ne = (sintPiAll$tPps)/(4*2e-8)
sintPiAll

sintPiAll$lineage = factor(sintPiAll$lineage)


```


```{r, sint nuvDiv plot}
sintNuclDivPlot = ggplot(sintPiAll, aes(x = lineage, y = tPps, fill = lineage, group = lineage)) +
  geom_bar(position = position_dodge2(preserve = "single"), stat = "identity", color = "black") +
  scale_fill_manual(values = sintKColPal) +
  geom_text(position = position_dodge2(preserve = "single", width = 0.9), aes(y = tPps-.0015, label= scales::comma(round(Ne,0)), hjust = 0, fontface = "bold"), angle = 90, color = "black", ) +
  labs(x = "Lineage", y = bquote("Nucleotide diversity ("*pi*")")) + 
  theme_bw() +
  violinTheme +
  theme(axis.text.x = ggtext::element_markdown())
  
sintNuclDivPlot

```

<br>

## Symbiodiniaceae

```{r, sint sym}
sintData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "S. intersepta") %>% dplyr::filter(!sample %in% c("SGM007", "SGM010", "SGM013", "SGM027.2", "SGM027.3", "SGM046.1", "SGM046.3", "SGM063", "SGM152.2", "SGM152.3", "SGM169", "SGM177", "SGM179", "SGM186.1", "SGM186.3", "SGM197.2", "SGM197.3", "SGM206.2", "SGM206.3")) %>% dplyr::select(sample, "site" = bank, "depth" = depthZone)

sintZoox = read.delim("../data/sint/sintZooxReads", header = FALSE, check.names = FALSE)

head(sintZoox)

# Reconstruct read mapping output into dataframe usable for analysis
sintZoox$V2[is.na(sintZoox$V2)] <- as.character(sintZoox$V1[is.na(sintZoox$V2)])
sintZoox$V1 = gsub(pattern = "MGM*.", "chr", sintZoox$V1)
sintZoox$V2 = gsub(".trim.*", "", sintZoox$V2)
sintZoox = sintZoox %>% filter(sintZoox$V1 != "*")
sintZooxLst = split(sintZoox$V2, as.integer(gl(length(sintZoox$V2), 20, length(sintZoox$V2))))

sintZooxMaps = NULL

for(i in sintZooxLst){
  sintZooxMaps = rbind(sintZooxMaps, data.frame(t(i)))
}

# remove tech reps
sintZooxMaps = sintZooxMaps %>% dplyr::filter(!X1 %in% c("SGM007", "SGM010", "SGM013", "SGM027.2", "SGM027.3", "SGM046.1", "SGM046.3", "SGM063", "SGM152.2", "SGM152.3", "SGM169", "SGM177", "SGM179", "SGM186.1", "SGM186.3", "SGM197.2", "SGM197.3", "SGM206.2", "SGM206.3"))


colnames(sintZooxMaps) = c("sample",sintZoox$V1[c(2:20)])

# convert characters to numeric
str(sintZooxMaps)

for(i in c(2:20)){
  sintZooxMaps[,i] = as.numeric(sintZooxMaps[,i])
  }

str(sintZooxMaps)

# collapse fake chromosomes into representative genera
sintZooxMaps$Symbiodinium = rowSums(sintZooxMaps[2:6])
sintZooxMaps$Breviolum = rowSums(sintZooxMaps[7:10])
sintZooxMaps$Cladocopium = rowSums(sintZooxMaps[11:16])
sintZooxMaps$Durusdinium = rowSums(sintZooxMaps[17:20])

# keep genera totals and turn into proportions for barplot
sintZooxMaps = sintZooxMaps[,c(1, 21:24)]
sintZooxProp = sintZooxMaps
sintZooxProp$sum = apply(sintZooxProp[, c(2:length(sintZooxProp[1,]))], 1, function(x) {
sum(x, na.rm = T)
})
sintZooxProp = cbind(sintZooxProp$sample, (sintZooxProp[, c(2:(ncol(sintZooxProp)-1))]
/ sintZooxProp$sum))

colnames(sintZooxProp)[1] = "sample"

head(sintZooxProp)

# Check that all samples total to 1
apply(sintZooxProp[, c(2:(ncol(sintZooxProp)))], 1, function(x) {
sum(x, na.rm = T)
})

# add sample metadata to proportions
sintSym = sintData %>% left_join(sintZooxProp)

sintAdmixOrd = sintAdmix %>% dplyr::select(sample, ord) 
sintPcangsdITS =sintPcangsd %>% dplyr::select(sample, cluster)

sintSym = sintSym %>% left_join(sintAdmixOrd) %>% relocate(ord,.after = depth) %>% left_join(sintPcangsdITS) %>% relocate(cluster, .after = depth)

```


```{r, sint sym permanova}
sintZooxPermData = sintData %>% left_join(sintZooxMaps) %>% left_join(sintPcangsdITS) %>% relocate(cluster, .after = depth)

sintZooxPermData$depth = factor(sintZooxPermData$depth)
sintZooxPermData$depth = factor(sintZooxPermData$depth, levels = levels(sintZooxPermData$depth)[c(2, 1)])

sintZooxPermData$site = factor(sintZooxPermData$site)
sintZooxPermData$site = factor(sintZooxPermData$site, levels = levels(sintZooxPermData$site)[c(5, 2, 1, 3, 4)])

head(sintZooxPermData)


set.seed(694)


sintSymPerm = adonis2(decostand(sintZooxPermData[ c(5:((ncol(sintZooxPermData))))], "normalize") ~ depth * site, data = sintZooxPermData, permutations = 9999, method = "bray", sqrt.dist = TRUE, by = "terms")

sintSymPerm

sintSymPermPW = pairwise.adonis2(decostand(sintZooxPermData[ c(5:((ncol(sintZooxPermData))))], "normalize") ~ site, data = sintZooxPermData, nperm = 9999, method = "bray", sqrt.dist = TRUE, by = "terms")

sintSymPermPW

```


*S. intersepta* lineage  vs Symbiodiniaceae B-C distance

```{r, sint procrustes its2 snps}
sintZooxDist = sintZooxPermData[ c(5:ncol(sintZooxPermData))] %>% decostand("normalize") %>% vegdist(method = "bray")

sintZooxPCoA = cmdscale(sintZooxDist, eig = TRUE, x.ret = TRUE)

sintAdmixDist = sintAdmixK2[c(5:ncol(sintAdmixK2))] %>% vegdist(method = "euclidean")

sintAdmixPCoA = cmdscale(sintAdmixDist, eig = TRUE, x.ret = TRUE)

set.seed(981) 
sintProcrustes = protest(X = sintAdmixPCoA, Y = sintZooxPCoA , permutations = 9999)

sintProcrustes

```


```{r, sint sym 2}
sintSym$depth = factor(sintSym$depth)
sintSym$depth = factor(sintSym$depth, levels = levels(sintSym$depth)[c(2, 1)])

sintSym$site = factor(sintSym$site)
sintSym$site = factor(sintSym$site, levels = levels(sintSym$site)[c(5, 2, 1, 3, 4)])

#turn into melted dataframe with otustack() and remove "summ" rows
sintGssSym = otuStack(sintSym, count.columns = c(6:length(sintSym[1, ])),
 condition.columns = c(1:5)) %>% filter(otu != "summ") %>% droplevels()

#check that levels are correct/ordered
levels(sintGssSym$otu)
levels(sintGssSym$depth)
levels(sintGssSym$site)

```

<br>

Creating Symbiodiniaceae genera relative proportion barplot

```{r, sint sym plot}
sintZooxAnno = data.frame(x1 = c(0.5, 0.5, 0.5, 0.5, 0.5), x2 = c(30.5, 30.5, 30.5, 30.5, 30.5),
y1 = -0.22, y2 = -0.22, site = c("WFGB", "EFGB", "Bright", "Geyer", "McGrail"))

sintZooxAnno$site = factor(sintZooxAnno$site)
sintZooxAnno$site = factor(sintZooxAnno$site, levels = levels(sintZooxAnno$site)[c(5, 2, 1, 3, 4)])

sintGssSymPlot = sintGssSym %>% left_join(sintZooxAnno, by = "site") %>% left_join(sintPcangsdITS)
sintGssSymPlot$ord = as.numeric(sintGssSymPlot$ord)

sintZooxPlotA = ggplot(data = sintGssSymPlot, aes(x = ord, y = count, fill = otu, order = ord)) +

geom_point(aes(x=1, y=0.5, fill = otu), shape = 22, size = 0) +
  
geom_bar(stat = "identity", position = "stack", colour = "grey25", width = 1, size = 0.2, show.legend = FALSE) +
  
xlab("Population") +
  
scale_fill_manual(values = colPalZoox, name = "Symbiodiniaceae \ngenus") +
  
geom_segment(data = (sintGssSymPlot %>% filter(depth == "Mesophotic")), aes(x = x1, xend = x2, y = y1, yend = y2, color = site), size = 7) +

scale_color_manual(values = gomPal, guide = "none") +
  
ggnewscale::new_scale_color() +

geom_segment(data = sintGssSymPlot, aes(x = ord-0.5, xend = ord+0.5, color = cluster), y = -.07, yend = -.07, linewidth = 4) +
  
scale_color_manual(values = sintKColPal[c(1,2,7)], name = "Lineage") +

coord_cartesian(ylim = c(0, 1), xlim = c(0.5, 30.5), clip = "off") +

scale_x_discrete(expand = c(0.005, 0.005)) +
scale_y_continuous(expand = c(0.001, 0.001)) +
facet_grid(factor(depth) ~ site, drop = TRUE, scales = "free", switch = "both", space = "free") +

geom_text(data = subset(sintGssSymPlot, subset = otu == "Symbiodinium") %>% filter(sample %in% c("SGM075", "SGM127", "SGM058", "SGM208", "SGM020")), x = 15.5, y = -.205, aes(label = site), size = 4.4, color = "#000000") +
guides(fill = guide_legend(override.aes = list(size = 4), ncol = 1, order = 1), color = guide_legend(override.aes = list(size = 0.1), ncol = 1)) +
  
theme_bw()

sintZooxPlot = sintZooxPlotA + zooxTheme

sintZooxPlot

```

<br>

## All *S. intersepta* plots
***

```{r, all sint plots}
sintPlot1 = sintStructure + inset_element(sintImg, -0.1, 0.7, 0.3, 1, align_to = "full", ignore_tag = TRUE)

sintPlot2 = (sintPcaPlots) + plot_layout(guides = "collect")

sintPlot3 = (sintLineageViolin+ theme(axis.text.y = element_text(margin = ggplot2::margin(r = -15, unit = "pt")))) + sintHetPlot + sintNuclDivPlot

sintPlot4 = sintAdmixPlot + sintZooxPlot

sintPlots =  sintPlot1 + sintPlot2 + sintPlot3 + sintPlot4 +
  plot_layout(heights = c(2,1,1,1)) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 18),
        legend.spacing = unit(-5, "pt"),
        legend.key = element_blank(),
        legend.background = element_blank())

ggsave("../figures/figure3.png", plot = sintPlots, height = 14, width = 12, units = "in", dpi = 300)

ggsave("../figures/figure3.svg", plot = sintPlots, height = 14, width = 12, units = "in", dpi = 300)

```

<br><br>

# *Orbicella faveolata*
***

## Dendrogram and structure
***

```{r, ofav dend no clones, warning=FALSE, message=FALSE}
ofavBams = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "O. faveolata") %>% dplyr::filter(!sample %in% c("OGM003", "OGM043", "OGM075", "OGM076", "OGM121", "OGM126", "OGM108.2", "OGM108.3", "OGM024.3", "OGM024.1", "OGM017", "OGM066.2", "OGM066.1", "OGM071", "OGM072", "OGM016", "OGM009", "OGM077", "OGM086", "OGM013", "OGM115", "OGM118", "OGM012", "OGM124", "OGM005"))# list of bams files and their populations (tech reps removed)

ofavBams$sample <- gsub("\\.[1-3]*$", "", ofavBams$sample)

ofavBams$bank = factor(ofavBams$bank)
ofavBams$bank = factor(ofavBams$bank, levels = levels(mcavBams$bank))

ofavBams$depthZone = factor(ofavBams$depthZone)
ofavBams$depthZone = factor(ofavBams$depthZone, levels = levels(ofavBams$depthZone)[c(2, 1)])


ofavMa = as.matrix(read.table("../data/ofav/ofavNoClones.ibsMat")) # reads in IBS matrix produced by ANGSD


dimnames(ofavMa) = list(ofavBams[,3],ofavBams[,3])


## admixture K = 2
#-------------------------------------
ofavK2ad = read.table("../data/ofav/admix/ofavK2.output") %>% dplyr::select(V6, V7) 
ofavK2ad %>% summarise(sum(V6),sum(V7))
ofavAdmixK2 = ofavBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(ofavK2ad) %>% 
  dplyr::rename("Of1" = "V6", "Of2" = "V7")
ofavAdmixK2_melt = melt(ofavAdmixK2, id = c("sample", "bank", "depth", "depthm"))

## admixture K = 3
#-------------------------------------
ofavK3ad = read.table("../data/ofav/admix/ofavK3.output") %>% dplyr::select(V6, V7, V8) 
ofavK3ad %>% summarise(sum(V6),sum(V7), sum(V8))
ofavAdmixK3 = ofavBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(ofavK3ad) %>% 
  dplyr::rename("Of1" = "V6", "Of2" = "V7", "Of3" = "V8") %>%
  dplyr::select(order(colnames(.)))
ofavAdmixK3_melt = melt(ofavAdmixK3, id = c("sample", "bank", "depth", "depthm"))


## ngsAdmix K = 4
#-------------------------------------
ofavK4ad = read.table("../data/ofav/admix/ofavK4.output") %>% dplyr::select(V6, V7, V8, V9) 
ofavK4ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9))
ofavAdmixK4 = ofavBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(ofavK4ad) %>% 
  dplyr::rename("Of1" = "V6", "Of2" = "V7", "Of3" = "V8", "Of4" = "V9") %>%
  dplyr::select(order(colnames(.)))
ofavAdmixK4_melt = melt(ofavAdmixK4, id = c("sample", "bank", "depth", "depthm"))


## admixture K = 5
#-------------------------------------
ofavK5ad = read.table("../data/ofav/admix/ofavK5.output") %>% dplyr::select(V6, V7, V8, V9, V10)
ofavK5ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9), sum(V10))
ofavAdmixK5 = ofavBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(ofavK5ad) %>% 
  dplyr::rename("Of1" = "V6", "Of2" = "V7", "Of3" = "V8", "Of4" = "V9", "Of5" = "V10") #%>%
ofavAdmixK5_melt = melt(ofavAdmixK5, id = c("sample", "bank", "depth", "depthm"))


## admixture K = 6
#-------------------------------------
ofavK6ad = read.table("../data/ofav/admix/ofavK6.output") %>% dplyr::select(V6, V7, V8, V9, V10,V11)
ofavK6ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9), sum(V10), sum(V11))
ofavAdmixK6 = ofavBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(ofavK6ad) %>% 
  dplyr::rename("Of1" = "V6", "Of2" = "V7", "Of3" = "V8", "Of4" = "V9", "Of5" = "V10", "Of6" = "V11") #%>%
ofavAdmixK6_melt = melt(ofavAdmixK6, id = c("sample", "bank", "depth", "depthm"))

## construct figure
#-------------------------------------
# ggtree(tr) + geom_nodelab(aes(label = node), hjust = -0.5) 

{ 
  ofavTr = hclust(dist(ofavMa),"ave") 

  ofavP1 = ggtree(ofavTr, layout = "rectangular", size = 0.35) 

  ofavP2 = facet_plot(ofavP1, panel = "location", data=ofavAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = bank), color = "grey25", size = 0.1, show.legend = TRUE) +
    scale_fill_manual("Site", values = gomPal, guide = guide_legend(order = 1), drop = FALSE) +
    new_scale_fill()

  ofavP3 = facet_plot(ofavP2, panel = "depth zone", data=ofavAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = depth), color = "grey25", size = 0.1) +
    scale_fill_manual("Depth zone", values = c("#A7E1BCFF", "#414083FF"), guide = guide_legend(order = 2)) +
    new_scale_fill()
  
  ofavP4 = facet_plot(ofavP3, panel = "depth", data=ofavAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = depthm), color = "grey25", size = 0.1) + 
    scale_fill_viridis_c("Depth (m)", option = "mako", trans = "reverse", limits = c(60, 10)) +
    new_scale_fill()

  ofavP5 = facet_plot(ofavP4, panel="K = 2", data=ofavAdmixK2_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                stat='identity', color = "grey25", size = 0.1) +
    scale_fill_manual("Lineage", values = ofavKColPal[1:6], guide = "none")
  
  ofavP6 = facet_plot( ofavP5, panel="K = 3", data=ofavAdmixK3_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE)  

  ofavP7 = facet_plot(ofavP6, panel="K = 4", data=ofavAdmixK4_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE)  

  ofavP8 = facet_plot(ofavP7, panel="K = 5", data=ofavAdmixK5_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) 
    
  ofavP9 = facet_plot(ofavP8, panel="K = 6", data=ofavAdmixK6_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), 
                  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) +
    theme_tree(strip.text = element_blank(),
               panel.spacing = unit(.1, "line")) 
}

ofavImg = image_read("../figures/icons/ofav.png") %>% image_ggplot()

ofavStructure = facet_widths(ofavP9, widths = c(0.5, 0.025, 0.025, 0.025, 0.1, 0.1, 0.2, 0.1, 0.1)) #+ inset_element(ofavImg, 0.0, 0.9, 0.1, 1.0, align_to = "full", ignore_tag = TRUE)

ofavStructure

```

<br>

## Population structure
***

```{r}
ofavData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "O. faveolata") %>% dplyr::filter(!sample %in% c("OGM003", "OGM043", "OGM075", "OGM076", "OGM121", "OGM126", "OGM108.2", "OGM108.3", "OGM024.3", "OGM024.1", "OGM017", "OGM066.2", "OGM066.1", "OGM071", "OGM072", "OGM016", "OGM009", "OGM077", "OGM086", "OGM013", "OGM115", "OGM118", "OGM012", "OGM124", "OGM005"))

ofavData$depthZone = factor(ofavData$depthZone)
ofavData$depthZone = factor(ofavData$depthZone, levels(ofavData$depthZone)[c(2,1)])

ofavData$bank = factor(ofavData$bank)
ofavData$bank = factor(ofavData$bank,levels(ofavData$bank)[c(5, 2, 1, 4, 3)])

ofavPcadmix = read.table("../data/ofav/admix/ofavK3.output") %>%dplyr::select(V6, V7, V8)
ofavPcadmix %>% summarise(sum(V6),sum(V7), sum(V8)) 

ofavPcadmix = ofavData %>% cbind(ofavPcadmix) %>% rename("cluster1" = "V6", "cluster2" = "V7", "cluster3" = "V8") %>%dplyr::select(order(colnames(.))) 

ofavPcadmixClust = ofavPcadmix %>% mutate(cluster = ifelse(cluster1 < 0.75 & cluster2 < 0.75  & cluster3 < 0.75, "NA", ifelse(cluster1 >=0.75, 1, ifelse(cluster2 >= 0.75, 2, ifelse(cluster3 >= 0.75, 3, 0)))))

ofavPcadmix = ofavPcadmix %>% mutate(ofavPcadmixClust)

ofavPcadmix$cluster = as.factor(ofavPcadmix$cluster)
levels(ofavPcadmix$cluster) = c("Ofav1", "Ofav2", "Ofav3", "Admixed")

ofavSiteLineages = ofavPcadmix %>% dplyr::select(depthZone, cluster) %>% 
group_by(depthZone) %>% count(cluster) %>% mutate(Freq = n/sum(n)) %>% apply(2, function(x) gsub("\\s+", "", x)) %>% as.data.frame()

```

<br>

### PCA
***

```{r, ofavPcangsd}
ofavCov = read.table("../data/ofav/ofavPcangsd.cov") %>% as.matrix()

ofavPcAdmix = read.table("../data/ofav/admix/ofavK3.output") %>%dplyr::select(V6, V7, V8)
ofavPcAdmix %>% summarise(sum(V6),sum(V7), sum(V8)) 

ofavPcAdmix = ofavPcAdmix %>% rename("cluster1" = "V6", "cluster2" = "V8", "cluster3" = "V7") %>% dplyr::select(order(colnames(.)))
  

ofavPcaEig = eigen(ofavCov)
ofavPcaVar = ofavPcaEig$values/sum(ofavPcaEig$values)*100
head(ofavPcaVar)

ofavPcangsd = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "O. faveolata") %>% dplyr::filter(!sample %in% c("OGM003", "OGM043", "OGM075", "OGM076", "OGM121", "OGM126", "OGM108.2", "OGM108.3", "OGM024.3", "OGM024.1", "OGM017", "OGM066.2", "OGM066.1", "OGM071", "OGM072", "OGM016", "OGM009", "OGM077", "OGM086", "OGM013", "OGM115", "OGM118", "OGM012", "OGM124", "OGM005")) %>% dplyr::select("sample" = sample, "bank" = bank, "depth" = depthZone, "depthm" = depthM)

ofavPcangsd$sitedepth = as.factor(paste(ofavPcangsd$bank, ofavPcangsd$depth, sep = " "))

ofavPcangsd$sitedepth = factor(ofavPcangsd$sitedepth, levels(ofavPcangsd$sitedepth)[c(4, 3,  2, 1)])

ofavPcangsd$bank = factor(ofavPcangsd$bank)
ofavPcangsd$bank = factor(ofavPcangsd$bank, levels(ofavPcangsd$bank)[c(5, 2, 1, 4, 3)])

ofavPcangsd$depth = factor(ofavPcangsd$depth)
ofavPcangsd$depth = factor(ofavPcangsd$depth, levels(ofavPcangsd$depth)[c(2, 1)])

ofavPcangsd$PC1 = ofavPcaEig$vectors[,1]
ofavPcangsd$PC2 = ofavPcaEig$vectors[,2]
ofavPcangsd$PC3 = ofavPcaEig$vectors[,3]
ofavPcangsd$PC4 = ofavPcaEig$vectors[,4]

ofavPcangsdClust = ofavPcAdmix %>% mutate(cluster = ifelse(cluster1 < 0.75 & cluster2 < 0.75  & cluster3 < 0.75, "NA", ifelse(cluster1 >=0.75, 1, ifelse(cluster2 >= 0.75, 2, ifelse(cluster3 >= 0.75, 3, 0)))))

# pcangsdClust$clusterX = as.vector(d_clust$classification)

ofavPcangsd = ofavPcangsd %>% mutate(ofavPcangsdClust)

ofavPcangsd$cluster = as.factor(ofavPcangsd$cluster)
levels(ofavPcangsd$cluster) = c("Of_Deep1", "Of_Deep2", "Of_Shal", "Admixed")

bamsClusters = ofavPcangsd %>% dplyr::select(sample, cluster) %>% dplyr::arrange(sample) 
bamssamples = read.delim("../data/ofav/bamsNoClones", header = FALSE)
bamsClusters$sample = bamssamples$V1

# bamsClusters = as.data.frame(bamsClusters)

write.table(x = bamsClusters, file = "../data/ofavBamsClusters", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

ofavPcangsd = merge(ofavPcangsd, aggregate(cbind(mean.x = PC1, mean.y = PC2)~sitedepth, ofavPcangsd, mean), by="sitedepth")


```

```{r}
ofavPcangsd %>% dplyr::group_by(depth,cluster) %>% dplyr::summarize(n = n())

```

<br>

#### Plot PCA

```{r, ofav pca plots}
ofavPcaPlot12SA = ggplot() +
  geom_hline(yintercept = 0, color = "gray90", size = 0.25) +
  geom_vline(xintercept = 0, color = "gray90", size = 0.25) +
  geom_point(data = ofavPcangsd, aes(x = PC1, y = PC2, fill = bank, shape = depth, color = bank), stroke = 0, size = 3, alpha = 0.5, show.legend = FALSE) +
  geom_point(data = ofavPcangsd, aes(x = mean.x, y = mean.y, fill = bank, shape = depth), color = "black", size = 3.75, alpha = 1, stroke = 0.25) +
  scale_shape_manual(values = c(21, 23), name = "Depth Zone") +
  scale_fill_manual(values = gomPal, name = "Site") +
  scale_color_manual(values = gomPal, name = "Site") +
  labs(x = paste0("PC 1 (", format(round(ofavPcaVar[1], 1), nsmall = 1)," %)")) +
  annotate(geom = "text", x = -0.11, y = 0.15, label = paste0("PC 2 (", format(round(ofavPcaVar[2], 1), nsmall = 1), " %)"), size = 3.65, angle = 90) +
  guides(shape = guide_legend(override.aes = list(size = 4, stroke = 0.25, alpha = NA), order = 2, ncol = 1), fill = guide_legend(override.aes = list(shape = 22, size = 4, fill = gomPal[c(1,2)], alpha = NA), order = 1, ncol = 1)) +
  coord_cartesian(xlim = c(-.09,.12), clip = "off") +
  theme_bw()

ofavPcaPlot12S = ofavPcaPlot12SA +
  pcaTheme +
  theme(legend.position = c(0.87, 0.82),
        axis.title.y = element_blank())

ofavPcaPlot12S


ofavPcaPlot12LA = ggplot() +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_vline(xintercept = 0, color = "gray90", size = 0.5) +
  geom_point(data = ofavPcangsd, aes(x = PC1, y = PC2, fill = cluster, shape = depth), color = "black", size = 3, alpha = 1, show.legend = TRUE) +
  scale_shape_manual(values = c(21, 23), name = "Depth Zone") +
  scale_fill_manual(values = ofavKColPal[c(1,3,2)], name = "Lineage") +
  labs(x = paste0("PC 1 (", format(round(ofavPcaVar[1], 1), nsmall = 1)," %)"), y = paste0("PC 2 (", format(round(ofavPcaVar[2], 1), nsmall = 1), " %)")) +
  guides(shape = "none", fill = guide_legend(override.aes = list(shape = 22, size = 4, alpha = NA, fill = ofavKColPal[c(1,3,2)]), order = 1, ncol = 1))+
  coord_cartesian(xlim = c(-.09,.12), clip = "off") +
  theme_bw()

ofavPcaPlot12L = ofavPcaPlot12LA +
  pcaTheme +
  theme(legend.position = c(0.9,0.86))


```

<br>

Put all plots together

```{r, ofav pca 1 plot}
ofavPcaPlots = (ofavPcaPlot12S + ofavPcaPlot12L) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 18),
        panel.background = element_rect(fill = "white"),
        legend.spacing = unit(-5, "pt"),
        legend.key = element_blank(),
        legend.background = element_blank())

ofavPcaPlots

```

<br>

### Admixture
***

Prepare admixture outputs for plotting

```{r, ofav prep admix}
ofavAdmix = ofavPcangsd %>%dplyr::select(-PC1, -PC2, -PC3, -PC4, -cluster, -depthm, -mean.x, -mean.y)
ofavAdmix$bank = factor(ofavAdmix$bank)

ofavAdmix = arrange(ofavAdmix, bank, depth, -cluster1, -cluster2)
ofavPopCounts = ofavAdmix %>% group_by(bank, depth) %>% summarise(n = n())

ofavPopCounts

ofavPopCountList = reshape2::melt(lapply(ofavPopCounts$n,function(x){c(1:x)}))
ofavAdmix$ord = ofavPopCountList$value

ofavAdmixMelt = melt(ofavAdmix, id.vars=c("sample", "bank", "depth", "sitedepth", "ord"), variable.name="Ancestry", value.name="Fraction")

ofavAdmixMelt$Ancestry = factor(ofavAdmixMelt$Ancestry)
ofavAdmixMelt$Ancestry = factor(ofavAdmixMelt$Ancestry, levels = rev(levels(ofavAdmixMelt$Ancestry)))

ofavPopAnno = data.frame(x1 = c(0.5, 0.5), x2 = c(33.5, 33.5),
                     y1 = -0.1, y2 = -0.1, sample = NA, Ancestry = NA, depth = "Mesophotic", 
                     ord  = NA, Fraction = NA,
                     bank = c("WFGB", "EFGB"))
ofavPopAnno$bank = factor(ofavPopAnno$bank)
ofavPopAnno$bank = factor(ofavPopAnno$bank, levels = levels(ofavPopAnno$bank)[c(2, 1)])

```

<br>

Make admixture plots

```{r, ofav admix plots}
ofavAdmixPlotA = ggplot(data = ofavAdmixMelt, aes(x = ord, y = Fraction, fill = Ancestry, order = sample)) +
  geom_segment(data = ofavPopAnno, aes(x = x1, xend = x2, y = -.1, yend = -.1, color = bank), size = 7) +
  geom_bar(stat = "identity", position = "fill", width = 1, colour = "grey25", size = 0.2) +
  facet_grid(factor(depth) ~ bank, switch = "both", scales = "free_x") +
  geom_text(data = (ofavAdmixMelt %>% filter(depth == "Mesophotic", bank %in% c("WFGB", "EFGB"), sample %in% c(	
"OGM001", "OGM073"), Ancestry == "cluster1")), x = 15.5, y = -.1, aes(label = bank), size = 4, color = "black") +
  
  scale_fill_manual(values = ofavKColPal[c(1,3,2)]) +
  scale_color_manual(values = gomPal) +
  scale_x_discrete(expand = c(0.005, 0.005)) +
  scale_y_continuous(expand = c(0.001, 0.001)) +
  coord_cartesian(ylim = c(0.0, 1.0), clip = "off") +
theme_bw()
  
ofavAdmixPlot = ofavAdmixPlotA + admixTheme
  
ofavAdmixPlot

```

<br>

## Lineage demographics
***

### Depth distribution of lineages
***

```{r, ofavdepth distribution}
leveneTest(lm(depthm ~ cluster, data = ofavPcangsd))

ofavDepthAov = welch_anova_test(depthm ~ cluster, data = subset(ofavPcangsd, subset = ofavPcangsd$cluster!="Admixed"))

ofavDepthAov

ofavDF = ofavDepthAov$statistic[[1]]

ofavDepthPH = games_howell_test(depthm ~ cluster, data = subset(ofavPcangsd, subset = ofavPcangsd$cluster!="Admixed"), conf.level = 0.95) %>% as.data.frame()

ofavDepthComps = paste(ofavDepthPH$group1,"-",ofavDepthPH$group2, sep ="")
# ofavComps = c("D1-D2", "D1-D3", "D1-D4", "D1-S1", "D1-S2", "D1-A",)
ofavDepthPs = ofavDepthPH$p.adj

ofDepthMultComp = ofavDepthPs < .05
names(ofDepthMultComp) = ofavDepthComps
ofDepthLet = multcompLetters(ofDepthMultComp)
ofDepthLetters = ofDepthLet$Letters %>% as.data.frame()

  
ofavDepthLetters = data.frame(x = factor(rownames(ofDepthLetters)), y = 5,  grp = ofDepthLetters$.)


ofavLineageViolinA = ggplot(data = subset(ofavPcangsd, subset = ofavPcangsd$cluster!="Admixed"), aes(x = cluster, y = depthm)) +
  annotate(geom = "rect", xmin = -Inf, xmax = Inf, ymin = 30, ymax = Inf,  fill = "black", alpha = 0.10, color = NA) +
  geom_violin(aes(fill = cluster),adjust = 1, linewidth = 0, color = "black", alpha = 0.75, width = 0.9, trim = F, scale = "area") +
    geom_beeswarm(aes(fill = cluster), shape = 21, size = 2, cex = 1.5, alpha = 0.5) +
  geom_violin(adjust = 1, linewidth = 0.4, color = "black", alpha = 1, width = 0.9, trim = F, fill = NA, scale = "area") +
  geom_boxplot(width = 0.1, color = "black", fill = "white", outlier.colour = NA, linewidth = 0.6, alpha = 0.5) +
  geom_segment(data = ticks[c(3:13),], aes(x = x, xend = xend, y = y)) +
  annotate(geom = "text", x = 2.75, y =65, label = bquote(italic("F")~" = "~.(ofavDF)*"; "~italic("p")~" < 0.001"), size = 3) +
  annotate(geom = "text", x = ofavDepthLetters$x, y = ofavDepthLetters$y, label = ofavDepthLetters$grp) +
  scale_fill_discrete(type = ofavKColPal[c(1,3,2)], name = "Lineage") +
  scale_color_discrete(type = ofavKColPal[c(1,3,2)], name = "Lineage") +
  xlab("Lineage") +
  ylab("Depth (m)") +
  scale_y_reverse(breaks = seq(5, 90, 5)) +
  coord_cartesian(xlim = c(0.75, 3.25), clip = "off") +
  theme_bw()

ofavLineageViolin = ofavLineageViolinA + violinTheme 

ofavLineageViolin

```

<br>

### Lineage differentiation

Measuring with global weighted FST calculated from SFS

First prepare and sort FST for plotting

```{r}
of.pops = c("Of_Deep1", "Of_Deep2", "Of_Shal") 

# reads in fst 
ofFstMa1 <- read.delim("../data/ofav/ofavKFst.out") %>% dplyr::select(-fst) %>% df_to_pw_mat(., "pop1", "pop2", "weightedFst")

colnames(ofFstMa1) = of.pops
rownames(ofFstMa1) = of.pops
ofFstMa = ofFstMa1

upperTriangle(ofFstMa, byrow = TRUE) <- lowerTriangle(ofFstMa)
ofFstMa <- ofFstMa[,of.pops] %>%
  .[of.pops,]
ofFstMa[lower.tri(ofFstMa)] <- NA
ofFstMa <- as.data.frame(ofFstMa)

# rearrange and reformat matrix
ofFstMa$Pop = factor(row.names(ofFstMa), levels = unique(of.pops))



# melt matrix data (turn from matrix into long dataframe)
ofFst = melt(ofFstMa, id.vars = "Pop", value.name = "Fst", variable.name = "Pop2", na.rm = FALSE)

ofFst$Fst = round(ofFst$Fst, 3)

ofFst$site = ofFst$Pop
ofFst$site = factor(ofFst$site)

ofFst$site2 = ofFst$Pop2

```

Construct a heatmap of FST values

```{r, ofav fstPlot}
ofFstHeatmapA = ggplot(data = ofFst %>% filter(Fst != 0), aes(Pop, Pop2, fill = as.numeric(as.character(Fst)))) +
  geom_tile(color = "white") +
  geom_segment(data = ofFst, aes(x = 0.475, xend = 0.25, y = Pop2, yend = Pop2, color = site2), size = 21.25) + #colored boxes for Y-axis labels
  geom_segment(data = ofFst, aes(x = Pop, xend = Pop, y = 0.2, yend = 0.475, color = site), size = 41) + #colored boxes for X-axis labels
  scale_color_manual(values = ofavKColPal[c(1,3,2)], guide = NULL) +
  scale_fill_gradient(low = "white", high = "#EA526F", limit = c(0, 0.24), space = "Lab", name = expression(paste(italic("F")[ST])), na.value = NA,  guide = "colourbar") +

  geom_text(data = ofFst %>% filter(Fst != 0), aes(Pop, Pop2, label = Fst), color = "black", size = 3.5, fontface = "bold") +
  guides(fill = guide_colorbar(barwidth = 7.5, barheight = 0.75, title.position = "top", title.hjust = 0.5, direction = "horizontal", ticks.colour = "black", frame.colour = "black")) +
  scale_y_discrete(position = "left", limits = rev(levels(ofFst$Pop2))) +
  scale_x_discrete(limits = levels(ofFst$Pop2)[c(1:3)]) +
  coord_cartesian(xlim = c(1, 3), ylim = c(1, 3), clip = "off") +
  theme_minimal()

ofFstHeatmap = ofFstHeatmapA + theme(
  axis.text.x = element_text(vjust = 3.5, size = 10, hjust = 0.5, color = "black"),
  axis.text.y = element_text(size = 10, color = "black", angle = 90, hjust = 0.5, vjust = -1.5),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  legend.title = element_text(size = 8, color = "black"),
  legend.text = element_text(size = 8, color = "black"),
  legend.position = c(0.6, 0.9),
  plot.background = element_blank(),
  panel.background = element_blank(),
)

ofFstHeatmap

```

<br>

### Heterozygosity

```{r, ofav het data}
ofavData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "O. faveolata") %>% dplyr::filter(!sample %in% c("OGM003", "OGM043", "OGM075", "OGM076", "OGM121", "OGM126", "OGM108.2", "OGM108.3", "OGM024.3", "OGM024.1", "OGM017", "OGM066.2", "OGM066.1", "OGM071", "OGM072", "OGM016", "OGM009", "OGM077", "OGM086", "OGM013", "OGM115", "OGM118", "OGM012", "OGM124", "OGM005"))

ofavHetData = read.delim("../data/ofav/ofavHet.out", header = FALSE, sep = " ") %>% mutate(sample = ofavData$sample)  %>% rename("Het" = "V2") %>% left_join(ofavPcangsd) %>% select(sample, Het, cluster)
  
ofavHet = ofavData %>% left_join(ofavHetData)

```


```{r, ofav het AOV}
ofavHetAOV = welch_anova_test(ofavHet, Het ~ cluster)

ofavHetPH = games_howell_test(ofavHet, Het ~ cluster)

ofavHetPH$p.adj

ofavComps = paste(ofavHetPH$group1,"-",ofavHetPH$group2, sep ="")
ofavPs = ofavHetPH$p.adj

ofMultComp = ofavPs < .05
names(ofMultComp) = ofavComps
ofLet = multcompLetters(ofMultComp)
ofLetters = ofLet$Letters %>% as.data.frame()
  
ofavHetLetters = data.frame(x = factor(rownames(ofLetters)), y = 0.0024,  grp = ofLetters$.)

ofavFST = read.delim("../data/ofav/ofavKFst.out")
ofavFST$y = c(0.0025, 0.0031, 0.0028)
```



```{r, ofav het plots}
ofavHetPlot = ggplot(data = ofavHet, aes(x = cluster, y = Het)) +
  geom_violin(aes(fill = cluster, group = cluster), adjust = 1, linewidth = 0.4, color = "black", alpha = 1, trim = FALSE, scale = "area", width = 1) +
  geom_beeswarm(aes(fill = cluster), shape = 21, cex = 0.5) +
  geom_boxplot(aes(fill = cluster, group = cluster), width = 0.2, color = "black", fill = "white", outlier.colour = NA, linewidth = 0.4, alpha = 0.5) +  xlab("Lineage") +
  geom_text(data = ofavHetLetters, aes(x = x, y = y, label = grp), size = 4) +
   annotate(geom = "text", x = 2.85, y = 0, label = bquote(italic("F")~" = "~.(ofavHetAOV$statistic)*"; "~italic("p")~" =  "~.(ofavHetAOV$p)), size = 3) +
  geom_segment(data = ofavFST, aes(y = y, x = pop1, xend = pop2)) +
  annotate(geom = "text", x = 1.5, y = 0.0026, label = bquote(italic("F")["ST"]~" = "~.(round(ofavFST$weightedFst[1], 3)))) +
  annotate(geom = "text", x = 2, y = 0.0032, label = bquote(italic("F")["ST"]~" = "~.(round(ofavFST$weightedFst[3], 3)))) +
  annotate(geom = "text", x = 2.5, y = 0.0029, label = bquote(italic("F")["ST"]~" = "~.(round(ofavFST$weightedFst[2], 3)))) +
  scale_fill_discrete(type = ofavKColPal[c(1,3,2)], name = "Lineage") +
  xlab("Lineage") +
  ylab("Heterozygosity") +
  scale_y_continuous(breaks = seq(0, 0.0035, 0.0005)) +
  coord_cartesian(expand = TRUE, xlim = c(0.85, 3.15), ylim = c(0, 0.0032)) +
  theme_bw() + 
  violinTheme +
  theme(panel.border = element_rect(linewidth = 1),
          axis.ticks.y = element_line(color = "black"))

ofavHetPlot

```

<br>

### Nucleotide diversity plot

```{r, of pi}
ofavNpgList = list(read_tsv("../data/ofav/Ofav_Deep1.thetas.idx.pestPG") %>% mutate(lineage = "Of_Deep1"),
               read_tsv("../data/ofav/Ofav_Deep2.thetas.idx.pestPG") %>% mutate(lineage = "Of_Deep2"),
               read_tsv("../data/ofav/Ofav_Shallow.thetas.idx.pestPG") %>% mutate(lineage = "Of_Shallow"))


ofavPiAll = purrr::reduce(ofavNpgList, rbind) %>% 
  dplyr::group_by(lineage) %>%
  filter(tP != 0) %>%
  mutate(tPps = tP/nSites) %>%
  dplyr::summarize(tPps = mean(tPps))

ofavPiAll$Ne = (ofavPiAll$tPps)/(4*2e-8)
ofavPiAll

ofavPiAll$lineage = factor(ofavPiAll$lineage)

```


```{r, nuvDiv plot}
ofavNuclDivPlot = ggplot(ofavPiAll, aes(x = lineage, y = tPps, fill = lineage, group = lineage)) +
  geom_bar(position = position_dodge2(preserve = "single"), stat = "identity", color = "black") +
  scale_fill_manual(values = ofavKColPal[c(1,3,2)]) +
  geom_text(position = position_dodge2(preserve = "single", width = 0.9), aes(y = tPps-.0015, label= scales::comma(round(Ne,0)), hjust = 0, fontface = "bold"), angle = 90, color = "black", ) +
  labs(x = "Lineage", y = bquote("Nucleotide diversity ("*pi*")")) + 
  theme_bw() +
  violinTheme +
  theme(axis.text.x = ggtext::element_markdown())
  
ofavNuclDivPlot

```

<br>

## Symbiodiniaceae
***

```{r, ofav sym}
ofavData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "O. faveolata") %>% dplyr::filter(!sample %in% c("OGM003", "OGM043", "OGM075", "OGM076", "OGM121", "OGM126", "OGM108.2", "OGM108.3", "OGM024.3", "OGM024.1", "OGM017", "OGM066.2", "OGM066.1", "OGM071", "OGM072", "OGM016", "OGM009", "OGM077", "OGM086", "OGM013", "OGM115", "OGM118", "OGM012", "OGM124", "OGM005")) %>% dplyr::select("sample" = sample, "site" = bank, "depth" = depthZone)

ofavZoox = read.delim("../data/ofav/ofavZooxReads", header = FALSE, check.names = FALSE)

head(ofavZoox)

# Reconstruct read mapping output into dataframe usable for analysis
ofavZoox$V2[is.na(ofavZoox$V2)] <- as.character(ofavZoox$V1[is.na(ofavZoox$V2)])
ofavZoox$V1 = gsub(pattern = "MGM*.", "chr", ofavZoox$V1)
ofavZoox$V2 = gsub(".trim.*", "", ofavZoox$V2)
ofavZoox = ofavZoox %>% filter(ofavZoox$V1 != "*")
ofavZooxLst = split(ofavZoox$V2, as.integer(gl(length(ofavZoox$V2), 20, length(ofavZoox$V2))))

ofavZooxMaps = NULL

for(i in ofavZooxLst){
  ofavZooxMaps = rbind(ofavZooxMaps, data.frame(t(i)))
}

# remove tech reps
ofavZooxMaps = ofavZooxMaps %>% dplyr::filter(!X1 %in% c("OGM003", "OGM043", "OGM075", "OGM076", "OGM121", "OGM126", "OGM108.2", "OGM108.3", "OGM024.3", "OGM024.1", "OGM017", "OGM066.2", "OGM066.1", "OGM071", "OGM072", "OGM016", "OGM009", "OGM077", "OGM086", "OGM013", "OGM115", "OGM118", "OGM012", "OGM124", "OGM005"))


colnames(ofavZooxMaps) = c("sample",ofavZoox$V1[c(2:20)])

# convert characters to numeric
str(ofavZooxMaps)

for(i in c(2:20)){
  ofavZooxMaps[,i] = as.numeric(ofavZooxMaps[,i])
  }

str(ofavZooxMaps)

# collapse fake chromosomes into representative genera
ofavZooxMaps$Symbiodinium = rowSums(ofavZooxMaps[2:6])
ofavZooxMaps$Breviolum = rowSums(ofavZooxMaps[7:10])
ofavZooxMaps$Cladocopium = rowSums(ofavZooxMaps[11:16])
ofavZooxMaps$Durusdinium = rowSums(ofavZooxMaps[17:20])

# keep genera totals and turn into proportions for barplot
ofavZooxMaps = ofavZooxMaps[,c(1, 21:24)]
ofavZooxProp = ofavZooxMaps
ofavZooxProp$sum = apply(ofavZooxProp[, c(2:length(ofavZooxProp[1,]))], 1, function(x) {
sum(x, na.rm = T)
})
ofavZooxProp = cbind(ofavZooxProp$sample, (ofavZooxProp[, c(2:(ncol(ofavZooxProp)-1))]
/ ofavZooxProp$sum))

colnames(ofavZooxProp)[1] = "sample"

head(ofavZooxProp)

# Check that all samples total to 1
apply(ofavZooxProp[, c(2:(ncol(ofavZooxProp)))], 1, function(x) {
sum(x, na.rm = T)
})

# add sample metadata to proportions
ofavSym = ofavData %>% left_join(ofavZooxProp)

ofavAdmixOrd = ofavAdmix %>% dplyr::select(sample, ord) 
ofavPcangsdITS =ofavPcangsd %>% dplyr::select(sample, cluster)

ofavSym = ofavSym %>% left_join(ofavAdmixOrd) %>% relocate(ord,.after = depth) %>% left_join(ofavPcangsdITS) %>% relocate(cluster, .after = depth)
```



```{r, ofav sym permanova}
ofavZooxPermData = ofavData %>% left_join(ofavZooxMaps) %>% left_join(ofavPcangsdITS) %>% relocate(cluster, .after = depth)

ofavZooxPermData$depth = factor(ofavZooxPermData$depth)
ofavZooxPermData$depth = factor(ofavZooxPermData$depth, levels = levels(ofavZooxPermData$depth)[c(2, 1)])

ofavZooxPermData$site = factor(ofavZooxPermData$site)
ofavZooxPermData$site = factor(ofavZooxPermData$site, levels = levels(ofavZooxPermData$site)[c(2, 1)])

head(ofavZooxPermData)


set.seed(694)


ofavSymPerm = adonis2(decostand(ofavZooxPermData[ c(5:((ncol(ofavZooxPermData))))], "normalize") ~ depth * site, data = ofavZooxPermData, permutations = 9999, method = "bray", sqrt.dist = TRUE, by = "terms")

ofavSymPerm

```

<br>

*O. faveolata* lineage  vs Symbiodiniaceae B-C distance

```{r, ofav procrustes its2 snps}
ofavZooxDist = ofavZooxPermData[ c(5:ncol(ofavZooxPermData))] %>% decostand("normalize") %>% vegdist(method = "bray")

ofavZooxPCoA = cmdscale(ofavZooxDist, eig = TRUE, x.ret = TRUE)

ofavAdmixDist = ofavAdmixK2[c(5:ncol(ofavAdmixK2))] %>% vegdist(method = "euclidean")

ofavAdmixPCoA = cmdscale(ofavAdmixDist, eig = TRUE, x.ret = TRUE)

set.seed(981) 
ofavProcrustes = protest(X = ofavAdmixPCoA, Y = ofavZooxPCoA , permutations = 9999)

ofavProcrustes

```

<br>

```{r, ofav sym 2}
ofavSym$depth = factor(ofavSym$depth)
ofavSym$depth = factor(ofavSym$depth, levels = levels(ofavSym$depth)[c(2, 1)])

ofavSym$site = factor(ofavSym$site)
ofavSym$site = factor(ofavSym$site, levels = levels(ofavSym$site)[c(5, 2, 1, 3, 4)])

#turn into melted dataframe with otustack() and remove "summ" rows
ofavGssSym = otuStack(ofavSym, count.columns = c(6:length(ofavSym[1, ])),
 condition.columns = c(1:5)) %>% filter(otu != "summ") %>% droplevels()

#check that levels are correct/ordered
levels(ofavGssSym$otu)
levels(ofavGssSym$depth)
levels(ofavGssSym$site)

```

<br>

### Symbiodiniaceae genera barplot
***

```{r, ofav sym plot}
ofavZooxAnno = data.frame(x1 = c(0.5, 0.5), x2 = c(33.5, 33.5),
y1 = -0.22, y2 = -0.22, site = c("WFGB", "EFGB"))

ofavZooxAnno$site = factor(ofavZooxAnno$site)
ofavZooxAnno$site = factor(ofavZooxAnno$site, levels = levels(ofavZooxAnno$site)[c(5, 2, 1, 3, 4)])

ofavGssSymPlot = ofavGssSym %>% left_join(ofavZooxAnno, by = "site") %>% left_join(ofavPcangsdITS)
ofavGssSymPlot$ord = as.numeric(ofavGssSymPlot$ord)

ofavZooxPlotA = ggplot(data = ofavGssSymPlot, aes(x = ord, y = count, fill = otu, order = ord)) +

geom_point(aes(x=1, y=0.5, fill = otu), shape = 22, size = 0) +
  
geom_bar(stat = "identity", position = "stack", colour = "grey25", width = 1, size = 0.2, show.legend = FALSE) +
  
xlab("Population") +
  
scale_fill_manual(values = colPalZoox, name = "Symbiodiniaceae \ngenus") +
  
geom_segment(data = (ofavGssSymPlot %>% filter(depth == "Mesophotic")), aes(x = x1, xend = x2, y = y1, yend = y2, color = site), size = 7) +

scale_color_manual(values = gomPal, guide = "none") +
  
ggnewscale::new_scale_color() +

geom_segment(data = ofavGssSymPlot, aes(x = ord-0.5, xend = ord+0.5, color = cluster), y = -.07, yend = -.07, linewidth = 4) +
  
scale_color_manual(values = ofavKColPal[c(1,3,2,7)], name = "Lineage") +

coord_cartesian(ylim = c(0, 1), xlim = c(0.5, 33.5), clip = "off") +

scale_x_discrete(expand = c(0.005, 0.005)) +
scale_y_continuous(expand = c(0.001, 0.001)) +
facet_grid(factor(depth) ~ site, drop = TRUE, scales = "free", switch = "both", space = "free") +

geom_text(data = subset(ofavGssSymPlot, subset = otu == "Symbiodinium") %>% filter(sample %in% c("OGM001", "OGM073")), x = 15.5, y = -.205, aes(label = site), size = 4.4, color = "#000000") +
guides(fill = guide_legend(override.aes = list(size = 4), ncol = 1, order = 1), color = guide_legend(override.aes = list(size = 0.1), ncol = 1)) +
  
theme_bw()

ofavZooxPlot = ofavZooxPlotA + zooxTheme

ofavZooxPlot

```

<br>

## All *O. faveolata* plots
***

```{r, all ofav plots}
ofavPlot1 = ofavStructure + inset_element(ofavImg, -0.1, 0.7, 0.3, 1, align_to = "full", ignore_tag = TRUE)
ofavPlot2 = (ofavPcaPlots) + plot_layout(guides = "collect")
ofavPlot3 = (ofavLineageViolin + theme(axis.text.y = element_text(margin = ggplot2::margin(r = -15, unit = "pt")))) + ofavHetPlot + ofavNuclDivPlot 
ofavPlot4 = ofavAdmixPlot + ofavZooxPlot

ofavPlots =  ofavPlot1 + ofavPlot2 + ofavPlot3 + ofavPlot4 +
  plot_layout(heights = c(2,1,1,1)) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 18),
        legend.spacing = unit(-5, "pt"),
        legend.key = element_blank(),
        legend.background = element_blank())

ggsave("../figures/figure4.png", plot = ofavPlots, height = 14, width = 12, units = "in", dpi = 300)

ggsave("../figures/figure4.svg", plot = ofavPlots, height = 14, width = 12, units = "in", dpi = 300)

```

<br><br>


# *Xestospongia muta*
***

## Dendrogram and structure
***

```{r, xesto dend no clones, warning=FALSE, message=FALSE}
xestoBams = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "X. muta") %>% dplyr::filter(!sample %in% c("XGM193.2", "XGM193.3", "XGM203.1", "XGM203.3", "XGM158.1", "XGM158.2", "XGM149", "XGM118", "XGM132", "XGM117", "XGM099", "XGM110", "XGM179.2", "XGM179.3", "XGM072.1", "XGM072.3", "XGM034.1", "XGM034.3", "XGM013", "XGM067", "XGM129"))# list of bams files and their populations (tech reps removed)

xestoBams$sample <- gsub("\\.[1-3]*$", "", xestoBams$sample)

xestoBams$bank = factor(xestoBams$bank)
xestoBams$bank = factor(xestoBams$bank, levels = levels(xestoBams$bank)[c(5, 2, 1, 3, 4)])

xestoBams$depthZone = factor(xestoBams$depthZone)
xestoBams$depthZone = factor(xestoBams$depthZone, levels = levels(xestoBams$depthZone)[c(2, 1)])


xestoMa = as.matrix(read.table("../data/xesto/xestoNoClones.ibsMat")) # reads in IBS matrix produced by ANGSD


dimnames(xestoMa) = list(xestoBams[,3],xestoBams[,3])


## admixture K = 2
#-------------------------------------
xestoK2ad = read.table("../data/xesto/admix/xestoK2.output") %>% dplyr::select(V6, V7) 
xestoK2ad %>% summarise(sum(V6),sum(V7))
xestoAdmixK2 = xestoBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(xestoK2ad) %>% 
  dplyr::rename("Xm1" = "V6", "Xm2" = "V7")
xestoAdmixK2_melt = melt(xestoAdmixK2, id = c("sample", "bank", "depth", "depthm"))

## admixture K = 3
#-------------------------------------
xestoK3ad = read.table("../data/xesto/admix/xestoK3.output") %>% dplyr::select(V6, V7, V8) 
xestoK3ad %>% summarise(sum(V6),sum(V7), sum(V8))
xestoAdmixK3 = xestoBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(xestoK3ad) %>% 
  dplyr::rename("Xm1" = "V6", "Xm2" = "V7", "Xm3" = "V8") %>%
  dplyr::select(order(colnames(.)))
xestoAdmixK3_melt = melt(xestoAdmixK3, id = c("sample", "bank", "depth", "depthm"))


## admixture K = 4
#-------------------------------------
xestoK4ad = read.table("../data/xesto/admix/xestoK4.output") %>% dplyr::select(V6, V7, V8, V9) 
xestoK4ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9))
xestoAdmixK4 = xestoBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(xestoK4ad) %>% 
  dplyr::rename("Xm1" = "V6", "Xm2" = "V9", "Xm3" = "V7", "Xm4" = "V8") %>%
  dplyr::select(order(colnames(.)))
xestoAdmixK4_melt = melt(xestoAdmixK4, id = c("sample", "bank", "depth", "depthm"))


## admixture K = 5
#-------------------------------------
xestoK5ad = read.table("../data/xesto/admix/xestoK5.output") %>% dplyr::select(V6, V7, V8, V9, V10)
xestoK5ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9), sum(V10))
xestoAdmixK5 = xestoBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(xestoK5ad) %>% 
  dplyr::rename("Xm1" = "V6", "Xm2" = "V9", "Xm3" = "V10", "Xm4" = "V7", "Xm5" = "V8") #%>%
xestoAdmixK5_melt = melt(xestoAdmixK5, id = c("sample", "bank", "depth", "depthm"))


## admixture K = 6
#-------------------------------------
xestoK6ad = read.table("../data/xesto/admix/xestoK6.output") %>% dplyr::select(V6, V7, V8, V9, V10,V11)
xestoK6ad %>% summarise(sum(V6),sum(V7), sum(V8), sum(V9), sum(V10), sum(V11))
xestoAdmixK6 = xestoBams %>%  
  dplyr::select(sample, bank, "depth" = depthZone, "depthm" = depthM) %>% 
  cbind(xestoK6ad) %>% 
  dplyr::rename("Xm1" = "V6", "Xm2" = "V9", "Xm3" = "V10", "Xm4" = "V11", "Xm5" = "V7", "Xm6" = "V8") #%>%
xestoAdmixK6_melt = melt(xestoAdmixK6, id = c("sample", "bank", "depth", "depthm"))

## construct figure
#-------------------------------------
# ggtree(tr) + geom_nodelab(aes(label = node), hjust = -0.5) 

{
  xestoTr = hclust(dist(xestoMa),"ave") 

  xestoP1 = ggtree(xestoTr, layout = "rectangular", size = 0.35) 

  xestoP2 = facet_plot(xestoP1, panel = "location", data=xestoAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = bank), color = "grey25", size = 0.1) + 
    scale_fill_manual("Site", values = gomPal, guide = guide_legend(order = 1)) + 
    new_scale_fill() 

  xestoP3 = facet_plot(xestoP2, panel = "depth zone", data=xestoAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = depth), color = "grey25", size = 0.1) +
    scale_fill_manual("Depth zone", values = c("#A7E1BCFF", "#414083FF"), guide = guide_legend(order = 2)) +
    new_scale_fill()
  
  xestoP4 = facet_plot(xestoP3, panel = "depth", data=xestoAdmixK2, geom = geom_tile, mapping = aes(x = 1, fill = depthm), color = "grey25", size = 0.1) + 
    scale_fill_viridis_c("Depth (m)", option = "mako", trans = "reverse", limits = c(60, 10)) +
    new_scale_fill()

  xestoP5 = facet_plot(xestoP4, panel="K = 2", data=xestoAdmixK2_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) +
    scale_fill_manual("Lineage",values = xestoKColPal[1:6])

  xestoP6 = facet_plot( xestoP5, panel="K = 3", data=xestoAdmixK3_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable),  stat='identity', color = "grey25", size = 0.1, show.legend = FALSE)  

  xestoP7 = facet_plot(xestoP6, panel="K = 4", data=xestoAdmixK4_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), stat='identity', color = "grey25", size = 0.1, show.legend = FALSE)  

  xestoP8 = facet_plot(xestoP7, panel="K = 5", data=xestoAdmixK5_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) 
    
  xestoP9 = facet_plot(xestoP8, panel="K = 6", data=xestoAdmixK6_melt, geom=geom_barh, width = 1.0, mapping = aes(x = value, fill = variable), stat='identity', color = "grey25", size = 0.1, show.legend = FALSE) +
    theme_tree(strip.text = element_blank(),
               panel.spacing = unit(.1, "line")) 
}

xestoImg = image_read("../figures/icons/xesto.png") %>% image_ggplot()

xestoStructure = facet_widths(xestoP9, widths = c(0.5, 0.025, 0.025, 0.025, 0.1, 0.1, 0.1, 0.1, 0.2))


xestoStructure

```

<br>

## Population structure
***

```{r}
xestoData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "X. muta") %>% dplyr::filter(!sample %in% c("XGM193.2", "XGM193.3", "XGM203.1", "XGM203.3", "XGM158.1", "XGM158.2", "XGM149", "XGM118", "XGM132", "XGM117", "XGM099", "XGM110", "XGM179.2", "XGM179.3", "XGM072.1", "XGM072.3", "XGM034.1", "XGM034.3", "XGM013", "XGM067", "XGM129"))

xestoData$depthZone = factor(xestoData$depthZone)
xestoData$depthZone = factor(xestoData$depthZone, levels(xestoData$depthZone)[c(2,1)])

xestoData$bank = factor(xestoData$bank)
xestoData$bank = factor(xestoData$bank,levels(xestoData$bank)[c(5, 2, 1, 4, 3)])

xestoPcadmix = read.table("../data/xesto/admix/xestoK6.output") %>%dplyr::select(V6, V7, V8, V9, V10, V11)
xestoPcadmix %>% summarise(sum(V6), sum(V7), sum(V8), sum(V9), sum(V10), sum(V11))


xestoPcadmix = xestoData %>% cbind(xestoPcadmix) %>% rename("cluster1" = "V6", "cluster2" = "V9", "cluster3" = "V10", "cluster4" = "V11", "cluster5" = "V7", "cluster6" = "V8") %>% dplyr::select(order(colnames(.)))

xestoPcadmixClust = xestoPcadmix %>% mutate(cluster = ifelse(cluster1 < 0.75 & cluster2 < 0.75 & cluster3 < 0.75 & cluster4 < 0.75 & cluster5 < 0.75 & cluster6 < 0.75, "NA", ifelse(cluster1 >=0.75, 1, ifelse(cluster2 >= 0.75, 2, ifelse(cluster3 >=0.75, 3, ifelse(cluster4 >= 0.75, 4, ifelse(cluster5 >=0.75, 5, ifelse(cluster6 >= 0.75, 6, 0))))))))   

xestoPcadmix = xestoPcadmix %>% mutate(xestoPcadmixClust)

xestoPcadmix$cluster = as.factor(xestoPcadmix$cluster)
levels(xestoPcadmix$cluster) = c("Xm_Deep1", "Xm_Deep2",  "Xm_Deep3", "Xm_Deep4", "Xm_Shal1", "Xm_Shal2", "Admixed")

xestoSiteLineages = xestoPcadmix %>% dplyr::select(depthZone, cluster) %>% 
group_by(depthZone) %>% count(cluster) %>% mutate(Freq = n/sum(n)) %>% apply(2, function(x) gsub("\\s+", "", x)) %>% as.data.frame()

xestoSiteLineages

```

### PCA
***

```{r, xesto pcangsd}
xestoCov = read.table("../data/xesto/xestoPcangsd.cov") %>% as.matrix()

xestoPcAdmix = read.table("../data/xesto/admix/xestoK6.output") %>%dplyr::select(V6, V7, V8, V9, V10, V11)

xestoPcAdmix %>% summarise(sum(V6), sum(V7), sum(V8), sum(V9), sum(V10), sum(V11))

# xestoPcAdmix = xestoPcAdmix %>% rename("cluster1" = "V1", "cluster2" = "V2", "cluster3" = "V3", "cluster4" = "V4", "cluster5" = "V5", "cluster6" = "V6", "cluster7" = "V7") %>% dplyr::select(order(colnames(.)))

xestoPcAdmix = xestoPcAdmix %>% rename("Xm_Deep1" = "V6", "Xm_Deep2" = "V9", "Xm_Deep3" = "V10", "Xm_Deep4" = "V11", "Xm_Shal1" = "V7", "Xm_Shal2" = "V8") %>% dplyr::select(order(colnames(.)))
  

xestoPcaEig = eigen(xestoCov)
xestoPcaVar = xestoPcaEig$values/sum(xestoPcaEig$values)*100
head(xestoPcaVar)

xestoPcangsd = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "X. muta") %>% dplyr::filter(!sample %in% c("XGM193.2", "XGM193.3", "XGM203.1", "XGM203.3", "XGM158.1", "XGM158.2", "XGM149", "XGM118", "XGM132", "XGM117", "XGM099", "XGM110", "XGM179.2", "XGM179.3", "XGM072.1", "XGM072.3", "XGM034.1", "XGM034.3", "XGM013", "XGM067", "XGM129")) %>% dplyr::select("sample" = sample, "bank" = bank, "depth" = depthZone, "depthm" = depthM)

xestoPcangsd$sitedepth = as.factor(paste(xestoPcangsd$bank, xestoPcangsd$depth, sep = " "))

xestoPcangsd$sitedepth = factor(xestoPcangsd$sitedepth, levels(xestoPcangsd$sitedepth)[c(7, 6, 3, 2, 1, 4, 5)])

xestoPcangsd$bank = factor(xestoPcangsd$bank)
xestoPcangsd$bank = factor(xestoPcangsd$bank, levels(xestoPcangsd$bank)[c(5, 2, 1, 3, 4)])

xestoPcangsd$depth = factor(xestoPcangsd$depth)
xestoPcangsd$depth = factor(xestoPcangsd$depth, levels(xestoPcangsd$depth)[c(2, 1)])

xestoPcangsd$PC1 = xestoPcaEig$vectors[,1]
xestoPcangsd$PC2 = xestoPcaEig$vectors[,2]
xestoPcangsd$PC3 = xestoPcaEig$vectors[,3]
xestoPcangsd$PC4 = xestoPcaEig$vectors[,4]

xestoPcangsdClust = xestoPcAdmix %>%  mutate(cluster = ifelse(Xm_Deep1 < 0.75 & Xm_Deep2 < 0.75 & Xm_Deep3 < 0.75 & Xm_Deep4 < 0.75 & Xm_Shal1 < 0.75 & Xm_Shal2 < 0.75, "Admixed", ifelse(Xm_Deep1 >=0.75, "Xm_Deep1", ifelse(Xm_Deep2 >= 0.75, "Xm_Deep2", ifelse(Xm_Deep3 >=0.75, "Xm_Deep3", ifelse(Xm_Deep4 >= 0.75, "Xm_Deep4", ifelse(Xm_Shal1 >=0.75, "Xm_Shal1", ifelse(Xm_Shal2 >= 0.75, "Xm_Shal2", 0))))))))

# pcangsdClust$clusterX = as.vector(d_clust$classification)

xestoPcangsd = xestoPcangsd %>% mutate(xestoPcangsdClust)

xestoPcangsd$cluster = as.factor(xestoPcangsd$cluster)
xestoPcangsd$cluster = factor(xestoPcangsd$cluster, levels = levels(xestoPcangsd$cluster)[c(2:7,1)])


xestoBamsClusters = xestoPcangsd %>% dplyr::select(sample, cluster) %>% dplyr::arrange(sample) 

xestoBamssamples = read.delim("../data/xesto/bamsNoClones", header = FALSE)
xestoBamsClusters$sample = xestoBamssamples$V1

# bamsClusters = as.data.frame(bamsClusters)

write.table(x = xestoBamsClusters, file = "../data/xesto/xestoBamsClusters", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

xestoPcangsd = merge(xestoPcangsd, aggregate(cbind(mean.x = PC1, mean.y = PC2)~sitedepth, xestoPcangsd, mean), by="sitedepth")


```

```{r}
xestoPcangsd %>% dplyr::group_by(depth,cluster) %>% dplyr::summarize(n = n())

```

<br>

#### Plot PCA

```{r, xesto pca plots}
xestoPcaPlot12SA = ggplot() +
  geom_hline(yintercept = 0, color = "gray90", size = 0.25) +
  geom_vline(xintercept = 0, color = "gray90", size = 0.25) +
  geom_point(data = xestoPcangsd, aes(x = PC1, y = PC2, fill = bank, shape = depth, color = bank), stroke = 0, size = 3, alpha = 0.5, show.legend = FALSE) +
  geom_point(data = xestoPcangsd, aes(x = mean.x, y = mean.y, fill = bank, shape = depth), color = "black", size = 3.75, alpha = 1, stroke = 0.25) +
  scale_shape_manual(values = c(21, 23), name = "Depth Zone") +
  scale_fill_manual(values = gomPal, name = "Site") +
  scale_color_manual(values = gomPal, name = "Site") +
  labs(x = paste0("PC 1 (", format(round(xestoPcaVar[1], 1), nsmall = 1)," %)"), y = NULL) +
  guides(shape = guide_legend(override.aes = list(size = 4, stroke = 0.25, alpha = NA), order = 2, ncol = 1), fill = guide_legend(override.aes = list(shape = 22, size = 4, fill = gomPal[c(1,2,3,4,5)], alpha = NA), order = 1, ncol = 1)) +
  annotate(geom = "text", x = -0.1205, y = -0.017, label = paste0("PC 2 (", format(round(xestoPcaVar[2], 1), nsmall = 1), " %)"), angle = 90, size = 3.65) +
  coord_cartesian(xlim = c(-0.1, 0.12), clip = "off") +
  theme_bw()

xestoPcaPlot12S = xestoPcaPlot12SA +
  pcaTheme +
  theme(legend.position = c(0.1, 0.24))

xestoPcaPlot12S


xestoPcaPlot12LA = ggplot() +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_vline(xintercept = 0, color = "gray90", size = 0.5) +
  geom_point(data = xestoPcangsd, aes(x = PC1, y = PC2, fill = cluster, shape = depth), color = "black", size = 3, alpha = 1, show.legend = TRUE) +
  scale_shape_manual(values = c(21, 23), name = "Depth Zone") +
  scale_fill_manual(values = xestoKColPal, name = "Lineage") +
  labs(x = paste0("PC 1 (", format(round(xestoPcaVar[1], 1), nsmall = 1)," %)"), y = paste0("PC 2 (", format(round(xestoPcaVar[2], 1), nsmall = 1), " %)")) +
  guides(shape = "none", fill = guide_legend(override.aes = list(shape = 22, size = 4, alpha = NA), order = 1, ncol = 1))+
  coord_cartesian(xlim = c(-0.1, 0.12), clip = "off") +
  theme_bw()

xestoPcaPlot12L = xestoPcaPlot12LA +
  pcaTheme +
  theme(legend.position = c(0.9, 0.2))


```

Put all plots together

```{r, xesto pca 1 plot}
xestoPcaPlots = ((xestoPcaPlot12S + theme(axis.title.y = element_text(margin = ggplot2::margin(r = -20, unit = "pt")))) | xestoPcaPlot12L) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 18),
        panel.background = element_rect(fill = "white"),
        legend.spacing = unit(-5, "pt"),
        legend.key = element_blank(),
        legend.background = element_blank())

xestoPcaPlots

```

<br>

### Admixture
***

Prepare admixture outputs for plotting

```{r, xesto prep admix}
xestoAdmix = xestoPcangsd %>%dplyr::select(-PC1, -PC2, -PC3, -PC4, -cluster, -depthm, -mean.x, -mean.y)
xestoAdmix$bank = factor(xestoAdmix$bank)

xestoAdmix = arrange(xestoAdmix, bank, depth, -Xm_Deep1, -Xm_Deep4, -Xm_Shal2)
xestoPopCounts = xestoAdmix %>% group_by(bank, depth) %>% summarise(n = n())

xestoPopCounts

xestoPopCountList = reshape2::melt(lapply(xestoPopCounts$n,function(x){c(1:x)}))
xestoAdmix$ord = xestoPopCountList$value

xestoAdmixMelt = melt(xestoAdmix, id.vars=c("sample", "bank", "depth", "sitedepth", "ord"), variable.name="Ancestry", value.name="Fraction")

xestoAdmixMelt$Ancestry = factor(xestoAdmixMelt$Ancestry)
xestoAdmixMelt$Ancestry = factor(xestoAdmixMelt$Ancestry, levels = rev(levels(xestoAdmixMelt$Ancestry)))

xestoPopAnno = data.frame(x1 = c(0.5, 0.5, 0.5, 0.5, 0.5), x2 = c(30.5, 30.5, 30.5, 30.5, 30.5),
                     y1 = -0.1, y2 = -0.1, sample = NA, Ancestry = NA, depth = "Mesophotic", 
                     ord  = NA, Fraction = NA,
                     bank = c("WFGB", "EFGB", "Bright", "Geyer", "McGrail"))
xestoPopAnno$bank = factor(xestoPopAnno$bank)
xestoPopAnno$bank = factor(xestoPopAnno$bank, levels = levels(xestoPopAnno$bank)[c(5, 2, 1, 3, 4)])

```

<br>

Make admixture plots

```{r, xesto admix plots}
xestoAdmixPlotA = ggplot(data = xestoAdmixMelt, aes(x = ord, y = Fraction, fill = Ancestry, order = sample)) +
  geom_segment(data = xestoPopAnno, aes(x = x1, xend = x2, y = -.1, yend = -.1, color = bank), size = 7) +
  geom_bar(stat = "identity", position = "fill", width = 1, colour = "grey25", size = 0.2) +
  facet_grid(factor(depth) ~ bank, switch = "both", scales = "free_x") +
  geom_text(data = (xestoAdmixMelt %>% filter(depth == "Mesophotic", bank %in% c("WFGB", "EFGB", "Bright", "Geyer", "McGrail"), sample %in% c("XGM101", "XGM135", "XGM078", "XGM177", "XGM058"), Ancestry == "Xm_Deep1")), x = 15.5, y = -.1, aes(label = bank), size = 4, color = "black") +
  
  scale_fill_manual(values = xestoKColPal[c(1:6)]) +
  scale_color_manual(values = gomPal) +
  scale_x_discrete(expand = c(0.005, 0.005)) +
  scale_y_continuous(expand = c(0.001, 0.001)) +
  coord_cartesian(ylim = c(0.0, 1.0), clip = "off") +
theme_bw()
  
xestoAdmixPlot = xestoAdmixPlotA + 
  theme_bw()+
  admixTheme

xestoAdmixPlot

```

<br>

## Lineage demographics
***

### Lineage depth distribution
***

```{r, xm depth distribution}
leveneTest(lm(depthm ~ cluster, data = xestoPcangsd))

xestoDepthAov = welch_anova_test(depthm ~ cluster, data = subset(xestoPcangsd, subset = xestoPcangsd$cluster!="Admixed"))

xestoDF = xestoDepthAov$statistic[[1]]

xestoDepthPH = games_howell_test(depthm ~ cluster, data = xestoPcangsd, conf.level = 0.95) %>% as.data.frame()

xestoDepthComps = paste(xestoDepthPH$group1,"-",xestoDepthPH$group2, sep ="")
# xestoComps = c("D1-D2", "D1-D3", "D1-D4", "D1-S1", "D1-S2", "D1-A",)
xestoDepthPs = xestoDepthPH$p.adj

xmDepthMultComp = xestoDepthPs < .05
names(xmDepthMultComp) = xestoDepthComps
xmDepthLet = multcompLetters(xmDepthMultComp)
xmDepthLetters = xmDepthLet$Letters %>% as.data.frame()

  
xestoDepthLetters = data.frame(x = factor(rownames(xmDepthLetters)), y = 5,  grp = xmDepthLetters$.)

xestoLineageViolinA = ggplot(data = xestoPcangsd, aes(x = cluster, y = depthm)) +
  annotate(geom = "rect", xmin = -Inf, xmax = Inf, ymin = 30, ymax = Inf,  fill = "black", alpha = 0.10, color = NA) +
  geom_violin(aes(fill = cluster),adjust = 1, linewidth = 0, color = "black", alpha = 0.75, width = 0.7, trim = F, scale = "width") +
    geom_beeswarm(aes(fill = cluster), shape = 21, size = 2, cex = 1.5, alpha = 0.5) +
  geom_violin(adjust = 1, linewidth = 0.4, color = "black", alpha = 1, width = 0.7, trim = F, fill = NA, scale = "width") +
  geom_boxplot(width = 0.2, color = "black", fill = "white", outlier.colour = NA, linewidth = 0.6, alpha = 0.5) +
  geom_segment(data = ticks[c(1:13),], aes(x = x, xend = xend, y = y)) +
  annotate(geom = "text", x = 2.75, y =65, label = bquote(italic("F")~" = "~.(xestoDF)*"; "~italic("p")~" < 0.001"), size = 3) +
  annotate(geom = "text", x = xestoDepthLetters$x, y = xestoDepthLetters$y, label = xestoDepthLetters$grp) +
  scale_fill_discrete(type = xestoKColPal, name = "Lineage") +
  scale_color_discrete(type = xestoKColPal, name = "Lineage") +
  xlab("Lineage") +
  ylab("Depth (m)") +
  scale_y_reverse(breaks = seq(5, 90, 5)) +
  coord_cartesian(xlim = c(0.75, 7.25), clip = "off") +
  theme_bw()

xestoLineageViolin = xestoLineageViolinA + violinTheme + theme(axis.text.x = element_text(color = "black", size = 10, angle = 45, hjust = 0.9))

xestoLineageViolin

```

<br>

### Lineage differentiation

Measuring with global weighted FST calculated from SFS

First prepare and sort FST for plotting

```{r}
xm.pops = c("Xm_Deep1", "Xm_Deep2", "Xm_Deep3", "Xm_Deep4", "Xm_Shal1", "Xm_Shal2") 

# reads in fst 
xmFstMa1 <- read.delim("../data/xesto/xestoKFst.out") %>% dplyr::select(-fst) %>% df_to_pw_mat(., "pop1", "pop2", "weightedFst")

colnames(xmFstMa1) = xm.pops
rownames(xmFstMa1) = xm.pops
xmFstMa = xmFstMa1

upperTriangle(xmFstMa, byrow = TRUE) <- lowerTriangle(xmFstMa)
xmFstMa <- xmFstMa[,xm.pops] %>%
  .[xm.pops,]
xmFstMa[lower.tri(xmFstMa)] <- NA
xmFstMa <- as.data.frame(xmFstMa)

# rearrange and reformat matrix
xmFstMa$Pop = factor(row.names(xmFstMa), levels = unique(xm.pops))



# melt matrix data (turn from matrix into long dataframe)
xmFst = melt(xmFstMa, id.vars = "Pop", value.name = "Fst", variable.name = "Pop2", na.rm = FALSE)

xmFst$Fst = round(xmFst$Fst, 3)

xmFst$site = xmFst$Pop
xmFst$site = factor(xmFst$site)

xmFst$site2 = xmFst$Pop2

```

Construct a heatmap of FST values

```{r, xmuta fstPlot}
xmFstHeatmapA = ggplot(data = xmFst %>% filter(Fst != 0), aes(Pop, Pop2, fill = as.numeric(as.character(Fst)))) +
  geom_tile(color = "white") +
  geom_segment(data = xmFst, aes(x = 0.475, xend = 0.25, y = Pop2, yend = Pop2, color = site2), size = 11.5) + #colored boxes for Y-axis labels
  geom_segment(data = xmFst, aes(x = Pop, xend = Pop, y = 0, yend = 0.475, color = site), size = 22.5) + #colored boxes for X-axis labels
  scale_color_manual(values = xestoKColPal, guide = NULL) +
  scale_fill_gradient(low = "white", high = "#EA526F", limit = c(0, 0.165), space = "Lab", name = expression(paste(italic("F")[ST])), na.value = NA,  guide = "colourbar") +
  geom_text(data = xmFst %>% filter(Fst != 0), aes(Pop, Pop2, label = Fst), color = "black", size = 3.5, fontface = "bold") +
  guides(fill = guide_colorbar(barwidth = 7.5, barheight = 0.75, title.position = "top", title.hjust = 0.5, direction = "horizontal", ticks.colour = "black", frame.colour = "black")) +
  scale_y_discrete(position = "left", limits = rev(levels(xmFst$Pop2))) +
  scale_x_discrete(limits = levels(xmFst$Pop2)[c(1:6)]) +
  coord_cartesian(xlim = c(1, 6), ylim = c(1, 6), clip = "off") +
  theme_minimal()

xmFstHeatmap = xmFstHeatmapA + theme(
  axis.text.x = element_text(vjust = 13, size = 8, hjust = 0.5, color = "black"),
  axis.text.y = element_text(size = 5, color = "black", angle = 90, hjust = 0.5, vjust = -5),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  legend.title = element_text(size = 8, color = "black"),
  legend.text = element_text(size = 8, color = "black"),
  legend.position = c(0.6, 0.9),
  plot.background = element_blank(),
  panel.background = element_blank(),
)

xmFstHeatmap

```

<br>

### Heterozygosity

```{r, xesto het data}
xestoData = read.csv("../data/nwgomMetaData.csv") %>% dplyr::filter(species == "X. muta") %>% dplyr::filter(!sample %in% c("XGM193.2", "XGM193.3", "XGM203.1", "XGM203.3", "XGM158.1", "XGM158.2", "XGM149", "XGM118", "XGM132", "XGM117", "XGM099", "XGM110", "XGM179.2", "XGM179.3", "XGM072.1", "XGM072.3", "XGM034.1", "XGM034.3", "XGM013", "XGM067", "XGM129"))

xestoHetData = read.delim("../data/xesto/xestoHet.out", header = FALSE, sep = " ") %>% mutate(sample = xestoData$sample)  %>% rename("Het" = "V2") %>% left_join(xestoPcangsd) %>% select(sample, Het, cluster)
  
xestoHet = xestoData %>% left_join(xestoHetData)

```


```{r, xesto het AOV}
xestoHetAOV = welch_anova_test(xestoHet, Het ~ cluster)

xestoHetAOV$p

xestoHetPH = games_howell_test(xestoHet, Het ~ cluster)

xestoHetPH

xestoHetPH$p.adj

xestoComps = paste(xestoHetPH$group1,"-",xestoHetPH$group2, sep ="")
# xestoComps = c("D1-D2", "D1-D3", "D1-D4", "D1-S1", "D1-S2", "D1-A",)
xestoPs = xestoHetPH$p.adj

xmMultComp = xestoPs < .05
names(xmMultComp) = xestoComps
xmLet = multcompLetters(xmMultComp)
xmLetters = xmLet$Letters %>% as.data.frame()
  
xestoHetLetters = data.frame(x = factor(rownames(xmLetters)), y = 0.001,  grp = xmLetters$.)

```


```{r, xesto het plots}
xestoHetPlot = ggplot(data = xestoHet, aes(x = cluster, y = Het)) +
  geom_violin(aes(fill = cluster, group = cluster), adjust = 1, linewidth = 0.4, color = "black", alpha = 1, trim = FALSE, scale = "area", width = 1) +
  geom_beeswarm(aes(fill = cluster), shape = 21, cex = 0.5) +
  geom_boxplot(aes(fill = cluster, group = cluster), width = 0.2, color = "black", fill = "white", outlier.colour = NA, linewidth = 0.4, alpha = 0.5) +  xlab("Lineage") +
  geom_text(data = xestoHetLetters, aes(x = x, y = y, label = grp), size = 4) +
   annotate(geom = "text", x = 2.2, y = 0, label = bquote(italic("F")~" = "~.(xestoHetAOV$statistic)*"; "~italic("p")~" < 0.001"), size = 3) +
  scale_fill_discrete(type = xestoKColPal, name = "Lineage") +
  xlab("Lineage") +
  ylab("Heterozygosity") +
  scale_y_continuous(breaks = seq(0, 0.001, 0.0002)) +
  coord_cartesian(expand = TRUE, xlim = c(0.85, 7.15), ylim = c(0, 0.001)) +
  theme_bw() + 
  violinTheme +
  theme(axis.text.x = element_text(color = "black", size = 10, angle = 45, hjust = 0.9), axis.ticks.y = element_line(color = "black"))

xestoHetPlot

```

<br>

### Nucleotide diversity plot

```{r, xesto pi}
xestoNpgList = list(read_tsv("../data/xesto/Xm_Deep1.thetas.idx.pestPG") %>% mutate(lineage = "Xm_Deep1"),
                  read_tsv("../data/xesto/Xm_Deep2.thetas.idx.pestPG") %>% mutate(lineage = "Xm_Deep2"),
                  read_tsv("../data/xesto/Xm_Deep3.thetas.idx.pestPG") %>% mutate(lineage = "Xm_Deep3"),
                  read_tsv("../data/xesto/Xm_Deep4.thetas.idx.pestPG") %>% mutate(lineage = "Xm_Deep4"),
                  read_tsv("../data/xesto/Xm_Shal1.thetas.idx.pestPG") %>% mutate(lineage = "Xm_Shal1"),
                  read_tsv("../data/xesto/Xm_Shal2.thetas.idx.pestPG") %>% mutate(lineage = "Xm_Shal2")
               )


xestoPiAll = purrr::reduce(xestoNpgList, rbind) %>% 
  dplyr::group_by(lineage) %>%
  filter(tP != 0) %>%
  mutate(tPps = tP/nSites) %>%
  dplyr::summarize(tPps = mean(tPps))

xestoPiAll$Ne = (xestoPiAll$tPps)/(4*2e-8)
xestoPiAll

xestoPiAll$lineage = factor(xestoPiAll$lineage)

```

<br>

```{r, xesto nuvDiv plot}
xestoNuclDivPlot = ggplot(xestoPiAll, aes(x = lineage, y = tPps, fill = lineage, group = lineage)) +
  geom_bar(position = position_dodge2(preserve = "single"), stat = "identity", color = "black") +
  scale_fill_manual(values = xestoKColPal) +
  geom_text(position = position_dodge2(preserve = "single", width = 0.9), aes(y = tPps-.0015, label= scales::comma(round(Ne,0)), hjust = 0, fontface = "bold"), angle = 90, color = "black", ) +
  labs(x = "Lineage", y = bquote("Nucleotide diversity ("*pi*")")) + 
  theme_bw() +
  violinTheme +
  theme(axis.text.x = element_text(color = "black", size = 10, angle = 45, hjust = 0.9))#,
        #axis.text.x = ggtext::element_markdown())
  
xestoNuclDivPlot

```

<br>

## All *X. muta* plots
***

```{r, all xesto plots}
xestoPlot1 = xestoStructure + inset_element(xestoImg, -0.1, 0.7, 0.3, 1, align_to = "full", ignore_tag = TRUE)
xestoPlot2 = xestoPcaPlots + plot_layout(guides = "collect")
xestoPlot3 = (xestoLineageViolin + theme(axis.text.y = element_text(margin = ggplot2::margin(r = -15, unit = "pt")))) + xestoHetPlot + xestoNuclDivPlot
xestoPlot4 = xestoAdmixPlot + xmFstHeatmap

xestoPlots =  xestoPlot1 + xestoPlot2 + xestoPlot3 + xestoPlot4 +
  plot_layout(heights = c(2,1,1,1)) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 18),
        legend.spacing = unit(-5, "pt"),
        legend.key = element_blank(),
        legend.background = element_blank())

ggsave("../figures/figure5.png", plot = xestoPlots, height = 14, width = 11, units = "in", dpi = 300)

ggsave("../figures/figure5.svg", plot = xestoPlots, height = 14, width = 11, units = "in", dpi = 300)

```
 
<br><br>

# Genetic connectivity

## *Montastraea cavernosa*

Checking deviance among model runs from BayesAss we ran on HPC

```{r, BA3 mcav traces}
# fileList = substr(list.files("../data/snps/bayesAss/", "BA3trace.*.txt$"),1,10)
fileList = substr(list.files("../data/mcav/BA3/", "BA3trace.*.txt$"),1,15)

bayesian_deviance <- function(trace, burnin = 0, sampling.interval = 0){
  if(burnin == 0) stop('No burnin specified')
  if(sampling.interval == 0) stop('No sampling interval specified')
  range <- (trace$State > burnin & trace$State %% sampling.interval == 0)
  D <- -2*mean(trace$LogProb[range])
  return(D)
}

for(i in 1:length(fileList)){
  assign(fileList[i], read.delim(paste("../data/mcav/BA3/", fileList[i], ".txt", sep = ""))) %>% dplyr::select(-last_col())
  print(paste(fileList[i], bayesian_deviance(get(fileList[i]), burnin = 10000000, sampling.interval = 1000)))
}

# [1] "mcavBA3trace.01 669318.0188"
# [1] "mcavBA3trace.02 668830.0212"
# [1] "mcavBA3trace.03 669141.0585"
# [1] "mcavBA3trace.04 668427.0425"
# [1] "mcavBA3trace.05 668596.3534"
# [1] "mcavBA3trace.06 668835.2774"
# [1] "mcavBA3trace.07 668961.4509"
# [1] "mcavBA3trace.08 668469.9586"
# [1] "mcavBA3trace.09 668742.6304"
# [1] "mcavBA3trace.10 668837.7336"

```

All traces have similar deviance (this is good!).
Using the trace with the lowest deviance (BA3trace.02.txt, in this case)

```{r, BA3}
mcavBayesAss = read.delim("../data/mcav/BA3/mcavBA3trace.04.txt") %>% filter(State > 10000000) %>% dplyr::select(-State, -LogProb, -X)

mcavBaMean = mcavBayesAss %>% summarise(across(everything(), list(mean))) %>% t() %>% as_tibble() %>% rename(., mean=V1) %>% mutate(pops = colnames(mcavBayesAss))

mcavBaSumm = mcavBayesAss %>% summarise(across(everything(), list(median))) %>% t() %>% as.data.frame() %>% rename(., median=V1) %>% mutate(pops = mcavBaMean$pops, mean = round(mcavBaMean$mean, 3)) %>% relocate(median, .after = mean)

mcavBaSumm$median = round(mcavBaSumm$median, 3)

mcavBaHpd =as.data.frame(t(sapply(mcavBayesAss, emp.hpd)))
colnames(mcavBaHpd) = c("hpdLow", "hpdHigh")
mcavBaHpd$pops = rownames(mcavBaHpd)

mcavESS = as.data.frame(sapply(mcavBayesAss, ESS))
colnames(mcavESS) = "ESS"

mcavBaSumm = mcavBaSumm %>% left_join(mcavBaHpd)
mcavBaSumm$hpdLow = round(mcavBaSumm$hpdLow, 3)
mcavBaSumm$hpdHigh = round(mcavBaSumm$hpdHigh, 3)
mcavBaSumm$ESS = mcavESS$ESS

### FROM BAYESASS: ###
## Population Index -> Population Label:
## 0->Geyer_Mesophotic 1->EFGB_Mesophotic 
## 2->EFGB_Shallow 3->WFGB_Mesophotic 
## 4->WFGB_Shallow 5->Bright_Mesophotic
## 6->McGrail_Mesophotic

popi = rep(c("Geyer\nMesophotic", "EFGB\nMesophotic", "EFGB\nShallow", "WFGB\nMesophotic", "WFGB\nShallow", "Bright\nMesophotic", "McGrail\nMesophotic"), each = 7)

popj = rep(c("Geyer\nMesophotic", "EFGB\nMesophotic", "EFGB\nShallow", "WFGB\nMesophotic", "WFGB\nShallow", "Bright\nMesophotic", "McGrail\nMesophotic"), times = 7)

mcavBaSumm = mcavBaSumm %>% mutate(pop.i = popi, pop.j = popj) %>% relocate(c(pop.i, pop.j), .after = pops) %>% dplyr::select(-pops)

mcavBaSumm$pop.i = factor(mcavBaSumm$pop.i)
mcavBaSumm$pop.i = factor(mcavBaSumm$pop.i, levels = levels(mcavBaSumm$pop.i)[c(7, 3, 6, 2, 1, 4, 5)])

mcavBaSumm$pop.j = factor(mcavBaSumm$pop.j)
mcavBaSumm$pop.j = factor(mcavBaSumm$pop.j, levels = levels(mcavBaSumm$pop.j)[c(7, 3, 6, 2, 1, 4, 5)])

mcavBaSumm$site.i = word(mcavBaSumm$pop.i, 1, sep = "\n")
mcavBaSumm$site.i = factor(mcavBaSumm$site.i)
mcavBaSumm$site.i = factor(mcavBaSumm$site.i, levels = levels(mcavBaSumm$site.i)[c(5, 2, 1, 3, 4)])

mcavBaSumm$site.j = word(mcavBaSumm$pop.j, 1, sep = "\n")
mcavBaSumm$site.j = factor(mcavBaSumm$site.j)
mcavBaSumm$site.j = factor(mcavBaSumm$site.j, levels = levels(mcavBaSumm$site.j)[c(5, 2, 1, 3, 4)])

mcavBaSumm$depth.i = word(mcavBaSumm$pop.i, 2, sep = "\n")
mcavBaSumm$depth.i = factor(mcavBaSumm$depth.i)
mcavBaSumm$depth.i = factor(mcavBaSumm$depth.i, levels = levels(mcavBaSumm$depth.i)[c(2, 1)])

mcavBaSumm$depth.j = word(mcavBaSumm$pop.j, 2, sep = "\n")
mcavBaSumm$depth.j = factor(mcavBaSumm$depth.j)
mcavBaSumm$depth.j = factor(mcavBaSumm$depth.j, levels = levels(mcavBaSumm$depth.j)[c(2, 1)])

```


```{r, bayesAss means}
#All sites (excluding self retention)
mcavBaMeans = mcavBaSumm %>% filter(pop.i != pop.j) %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Global")

#mesophotic sources
mcavBaMeans = mcavBaSumm %>% filter(pop.i != pop.j, depth.j == "Mesophotic") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Mesophotic Source") %>% bind_rows(mcavBaMeans, .)

#shallow sources
mcavBaMeans = mcavBaSumm %>% filter(pop.i != pop.j, depth.j == "Shallow") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Shallow Source") %>% bind_rows(mcavBaMeans, .)

#mesophotic sinks
mcavBaMeans = mcavBaSumm %>% filter(pop.i != pop.j, depth.i == "Mesophotic") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.)))  %>% mutate(dataset = "Mesophotic Sink") %>% bind_rows(mcavBaMeans, .)

#shallow sinks
mcavBaMeans = mcavBaSumm %>% filter(pop.i != pop.j, depth.i == "Shallow") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.)))  %>% mutate(dataset = "Shallow Sink") %>% bind_rows(mcavBaMeans, .)

#mesophotic -> shallow
mcavBaMeans = mcavBaSumm %>% filter(pop.i != pop.j, depth.j == "Mesophotic", depth.i == "Shallow") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Mesophotic -> Shallow") %>% bind_rows(mcavBaMeans, .)

#mesophotic -> mesophotic
mcavBaMeans = mcavBaSumm %>% filter(pop.i != pop.j, depth.j == "Mesophotic", depth.i == "Mesophotic") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Mesophotic -> Mesophotic") %>% bind_rows(mcavBaMeans, .)

#shallow -> mesophotic
mcavBaMeans = mcavBaSumm %>% filter(pop.i != pop.j, depth.j == "Shallow", depth.i == "Mesophotic") %>% summarise(mean = mean(.$mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Shallow -> Mesophotic") %>% bind_rows(mcavBaMeans, .)

#shallow -> shallow
mcavBaMeans = mcavBaSumm %>% filter(pop.i != pop.j, depth.j == "Shallow", depth.i == "Shallow") %>% summarise(mean = round(mean(.$mean), 5), sd = round(sd(.$mean), 5), se = round(sd(.$mean)/sqrt(nrow(.)), 3)) %>% mutate(dataset = paste("Shallow -> Shallow")) %>% bind_rows(mcavBaMeans, .) %>% relocate(dataset, .before = mean) %>% as.data.frame()

mcavBaMeans[,c(2:4)] = mcavBaMeans[,c(2:4)] %>% round(4)

mcavBaMeans

mcavBaMeansPub = mcavBaMeans %>%
  flextable() %>%
  flextable::compose(part = "header", j = "dataset", value = as_paragraph("Dataset")) %>%
  flextable::compose(part = "header", j = "mean", value = as_paragraph(as_i("m"))) %>%
  flextable::compose(part = "header", j = "sd", value = as_paragraph("SD")) %>%
  flextable::compose(part = "header", j = "se", value = as_paragraph("SEM")) %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  flextable::fontsize(size = 10, part = "all") %>%
  flextable::bold(part = "header") %>%
  flextable::align(align = "left", part = "all") %>%
  flextable::autofit()

table = read_docx()
table = body_add_flextable(table, value = mcavBaMeansPub)
print(table, target = "../tables/table1.docx")

mcavBaMeansPub

```

```{r, migrationPlots}
mcavBaSumm$mean = sprintf('%.3f', mcavBaSumm$mean)
mcavBaSumm$mean2 = mcavBaSumm$mean
mcavBaSumm$hpdLow = sprintf('%.3f', mcavBaSumm$hpdLow)
mcavBaSumm$hpdHigh = sprintf('%.3f', mcavBaSumm$hpdHigh)

mcavBaLabs = tibble(pop.i = unique(mcavBaSumm$pop.i), pop.j = unique(mcavBaSumm$pop.j))

mcavMigrateA = ggplot(data = mcavBaSumm, aes(pop.i, pop.j))+
  geom_tile(data = subset(mcavBaSumm, subset = mcavBaSumm$mean2>0.65), fill = "gray35", color = "white") +
  geom_segment(data = mcavBaSumm, aes(x = 0.4755, xend = -0.42, y = pop.j, yend = pop.j, color = pop.j), size = 14) +
  geom_segment(data = mcavBaSumm, aes(x = pop.i, xend = pop.i, y = 0.45, yend = -0.425, color = pop.i), size = 32) +
  scale_color_manual(values = gomPal[c(1:2, 1:5)], guide = NULL) +

  guides(fill = guide_colorbar(ticks.colour = "black", barwidth = 1, barheight = 10, frame.colour = "black")) +
  # new_scale("fill") +
  geom_tile(data = subset(mcavBaSumm, subset = mcavBaSumm$mean<0.65), aes(fill = as.numeric(as.character(mean))), color = "white") +
  scale_fill_gradientn(colours = paletteer_c("viridis::mako", n = 10, direction = -1)[c(1:7)], limit = c(0,0.23), space = "Lab", name = expression(paste(italic("m"))), na.value = "transparent",  guide = "colourbar", values = c(0, 0.05, 0.1, 0.15, 0.2,0.5,0.75,1)) +
  # scale_fill_gradientn(colours = paletteer_d("khroma::smoothrainbow"), limit = c(0,0.27), space = "Lab", name = expression(paste(italic("m"))), na.value = "transparent",  guide = "colourbar", values = c(0, 0.05, 0.1, 0.15, 0.2,0.5,0.75,1)) +
  geom_text(data = mcavBaSumm, aes(x = pop.i, y = pop.j, label = paste(mean, "\n", sep = "")), color = ifelse(mcavBaSumm$mean > 0.6, "white", "gray5"), fontface = ifelse(as.numeric(mcavBaSumm$hpdLow)>0, "bold", "plain"), size = ifelse(as.numeric(mcavBaSumm$hpdLow)>0, 4.75, 4)) +
  geom_text(data = mcavBaSumm, aes(x = pop.i, y = pop.j, label = paste("\n(",hpdLow,"",hpdHigh, ")", sep = "")), color = ifelse(mcavBaSumm$mean > 0.6, "white", "gray5"), size = 3.25) +
  
  geom_text(data = (mcavBaLabs %>% filter(pop.j %in% c("Tortugas Bank\nMesophotic", "Tortugas Bank\nShallow", "Riley's Hump\nMesophotic", "Riley's Hump\nShallow"))), x = -.02, aes(y = pop.j, label = pop.j), size = 3.75, color = "#FFFFFF", family = "sans") +
  geom_text(data = (mcavBaLabs %>% filter(!pop.j %in% c("Tortugas Bank\nMesophotic", "Tortugas Bank\nShallow", "Riley's Hump\nMesophotic", "Riley's Hump\nShallow"))), x = -.02, aes(y = pop.j, label = pop.j), size = 3.75, color = "#000000", family = "sans") +
  geom_text(data = (mcavBaLabs %>% filter(pop.i %in% c("Tortugas Bank\nMesophotic", "Tortugas Bank\nShallow", "Riley's Hump\nMesophotic", "Riley's Hump\nShallow"))), y = -.01, aes(x = pop.i, label = pop.i), size = 3.75, color = "#FFFFFF", family = "sans") +
  geom_text(data = (mcavBaLabs %>% filter(!pop.i %in% c("Tortugas Bank\nMesophotic", "Tortugas Bank\nShallow", "Riley's Hump\nMesophotic", "Riley's Hump\nShallow"))), y = -.01, aes(x = pop.i, label = pop.i), size = 3.75, color = "#000000", family = "sans") +
  
  labs(x = "Sink", y = "Source") +
  scale_y_discrete(limits = rev(levels(mcavBaSumm$pop.i))[c(1:8)], position = "left") +
  coord_cartesian(xlim = c(1, 8), ylim = c(1, 8), clip = "off") +
  theme_minimal()

mcavMigrate = mcavMigrateA + theme(
  axis.text.x = element_text(vjust = 1, size = 12, hjust = 0.5, color = NA),
  axis.text.y = element_text(size = 10, color = NA),
   axis.title.x = element_text(size = 14, hjust = 0.425),
  axis.title.y = element_text(size = 14, hjust = 0.425),
  panel.grid.major = element_blank(),
  axis.ticks = element_blank(),
  # legend.position = c(1.055, 0.5),
  legend.direction = "vertical",
  legend.title = element_text(size = 12, face = "bold")
)

mcavMigrate


```

## *Stephanocoenia intersepta*

Checking deviance among model runs from BayesAss we ran on HPC

```{r, BA3 sint traces}
# fileList = substr(list.files("../data/snps/bayesAss/", "BA3trace.*.txt$"),1,10)
fileList = substr(list.files("../data/sint/BA3/", "BA3trace.*.txt$"),1,15)

bayesian_deviance <- function(trace, burnin = 0, sampling.interval = 0){
  if(burnin == 0) stop('No burnin specified')
  if(sampling.interval == 0) stop('No sampling interval specified')
  range <- (trace$State > burnin & trace$State %% sampling.interval == 0)
  D <- -2*mean(trace$LogProb[range])
  return(D)
}

for(i in 1:length(fileList)){
  assign(fileList[i], read.delim(paste("../data/sint/BA3/", fileList[i], ".txt", sep = ""))) %>% dplyr::select(-last_col())
  print(paste(fileList[i], bayesian_deviance(get(fileList[i]), burnin = 5000000, sampling.interval = 1000)))
}

# [1] "sintBA3trace.01 4095847.462"
# [1] "sintBA3trace.02 4094678.242"
# [1] "sintBA3trace.03 4095399.988"
# [1] "sintBA3trace.04 4095630.518"
# [1] "sintBA3trace.05 4095470.046"
# [1] "sintBA3trace.06 4094276.6"
# [1] "sintBA3trace.07 4095140.89"
# [1] "sintBA3trace.08 4095926.754"
# [1] "sintBA3trace.09 4095283.68"
# [1] "sintBA3trace.10 4095434.772"

```

All traces have similar deviance (this is good!).
Using the trace with the lowest deviance (BA3trace.02.txt, in this case)

```{r, sint BA3}
sintBayesAss = read.delim("../data/sint/BA3/sintBA3trace.06.txt") %>% filter(State > 5000000) %>% dplyr::select(-State, -LogProb, -X)

sintBaMean = sintBayesAss %>% summarise(across(everything(), list(mean))) %>% t() %>% as_tibble() %>% rename(., mean=V1) %>% mutate(pops = colnames(sintBayesAss))

sintBaSumm = sintBayesAss %>% summarise(across(everything(), list(median))) %>% t() %>% as.data.frame() %>% rename(., median=V1) %>% mutate(pops = sintBaMean$pops, mean = round(sintBaMean$mean, 3)) %>% relocate(median, .after = mean)

sintBaSumm$median = round(sintBaSumm$median, 3)

sintBaHpd =as.data.frame(t(sapply(sintBayesAss, emp.hpd)))
colnames(sintBaHpd) = c("hpdLow", "hpdHigh")
sintBaHpd$pops = rownames(sintBaHpd)

sintESS = as.data.frame(sapply(sintBayesAss, ESS))
colnames(sintESS) = "ESS"

sintBaSumm = sintBaSumm %>% left_join(sintBaHpd)
sintBaSumm$hpdLow = round(sintBaSumm$hpdLow, 3)
sintBaSumm$hpdHigh = round(sintBaSumm$hpdHigh, 3)
sintBaSumm$ESS = sintESS$ESS

### FROM BAYESASS: ###
## Population Index -> Population Label:
## 0->Geyer_Mesophotic 1->EFGB_Mesophotic 
## 2->EFGB_Shallow 3->WFGB_Mesophotic 
## 4->WFGB_Shallow 5->Bright_Mesophotic
## 6->McGrail_Mesophotic

popi = rep(c("Geyer\nMesophotic", "EFGB\nMesophotic", "EFGB\nShallow", "WFGB\nMesophotic", "WFGB\nShallow", "Bright\nMesophotic", "McGrail\nMesophotic"), each = 7)

popj = rep(c("Geyer\nMesophotic", "EFGB\nMesophotic", "EFGB\nShallow", "WFGB\nMesophotic", "WFGB\nShallow", "Bright\nMesophotic", "McGrail\nMesophotic"), times = 7)

sintBaSumm = sintBaSumm %>% mutate(pop.i = popi, pop.j = popj) %>% relocate(c(pop.i, pop.j), .after = pops) %>% dplyr::select(-pops)

sintBaSumm$pop.i = factor(sintBaSumm$pop.i)
sintBaSumm$pop.i = factor(sintBaSumm$pop.i, levels = levels(sintBaSumm$pop.i)[c(7, 3, 6, 2, 1, 4, 5)])

sintBaSumm$pop.j = factor(sintBaSumm$pop.j)
sintBaSumm$pop.j = factor(sintBaSumm$pop.j, levels = levels(sintBaSumm$pop.j)[c(7, 3, 6, 2, 1, 4, 5)])

sintBaSumm$site.i = word(sintBaSumm$pop.i, 1, sep = "\n")
sintBaSumm$site.i = factor(sintBaSumm$site.i)
sintBaSumm$site.i = factor(sintBaSumm$site.i, levels = levels(sintBaSumm$site.i)[c(5, 2, 1, 3, 4)])

sintBaSumm$site.j = word(sintBaSumm$pop.j, 1, sep = "\n")
sintBaSumm$site.j = factor(sintBaSumm$site.j)
sintBaSumm$site.j = factor(sintBaSumm$site.j, levels = levels(sintBaSumm$site.j)[c(5, 2, 1, 3, 4)])

sintBaSumm$depth.i = word(sintBaSumm$pop.i, 2, sep = "\n")
sintBaSumm$depth.i = factor(sintBaSumm$depth.i)
sintBaSumm$depth.i = factor(sintBaSumm$depth.i, levels = levels(sintBaSumm$depth.i)[c(2, 1)])

sintBaSumm$depth.j = word(sintBaSumm$pop.j, 2, sep = "\n")
sintBaSumm$depth.j = factor(sintBaSumm$depth.j)
sintBaSumm$depth.j = factor(sintBaSumm$depth.j, levels = levels(sintBaSumm$depth.j)[c(2, 1)])

```


```{r, sint bayesAss means}
#All sites (excluding self retention)
sintBaMeans = sintBaSumm %>% filter(pop.i != pop.j) %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Global")

#mesophotic sources
sintBaMeans = sintBaSumm %>% filter(pop.i != pop.j, depth.j == "Mesophotic") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Mesophotic Source") %>% bind_rows(sintBaMeans, .)

#shallow sources
sintBaMeans = sintBaSumm %>% filter(pop.i != pop.j, depth.j == "Shallow") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Shallow Source") %>% bind_rows(sintBaMeans, .)

#mesophotic sinks
sintBaMeans = sintBaSumm %>% filter(pop.i != pop.j, depth.i == "Mesophotic") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.)))  %>% mutate(dataset = "Mesophotic Sink") %>% bind_rows(sintBaMeans, .)

#shallow sinks
sintBaMeans = sintBaSumm %>% filter(pop.i != pop.j, depth.i == "Shallow") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.)))  %>% mutate(dataset = "Shallow Sink") %>% bind_rows(sintBaMeans, .)

#mesophotic -> shallow
sintBaMeans = sintBaSumm %>% filter(pop.i != pop.j, depth.j == "Mesophotic", depth.i == "Shallow") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Mesophotic -> Shallow") %>% bind_rows(sintBaMeans, .)

#mesophotic -> mesophotic
sintBaMeans = sintBaSumm %>% filter(pop.i != pop.j, depth.j == "Mesophotic", depth.i == "Mesophotic") %>% summarise(mean = mean(mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Mesophotic -> Mesophotic") %>% bind_rows(sintBaMeans, .)

#shallow -> mesophotic
sintBaMeans = sintBaSumm %>% filter(pop.i != pop.j, depth.j == "Shallow", depth.i == "Mesophotic") %>% summarise(mean = mean(.$mean), sd = sd(.$mean), se = sd(.$mean)/sqrt(nrow(.))) %>% mutate(dataset = "Shallow -> Mesophotic") %>% bind_rows(sintBaMeans, .)

#shallow -> shallow
sintBaMeans = sintBaSumm %>% filter(pop.i != pop.j, depth.j == "Shallow", depth.i == "Shallow") %>% summarise(mean = round(mean(.$mean), 5), sd = round(sd(.$mean), 5), se = round(sd(.$mean)/sqrt(nrow(.)), 3)) %>% mutate(dataset = paste("Shallow -> Shallow")) %>% bind_rows(sintBaMeans, .) %>% relocate(dataset, .before = mean) %>% as.data.frame()

sintBaMeans[,c(2:4)] = sintBaMeans[,c(2:4)] %>% round(4)

sintBaMeans

sintBaMeansPub = sintBaMeans %>%
  flextable() %>%
  flextable::compose(part = "header", j = "dataset", value = as_paragraph("Dataset")) %>%
  flextable::compose(part = "header", j = "mean", value = as_paragraph(as_i("m"))) %>%
  flextable::compose(part = "header", j = "sd", value = as_paragraph("SD")) %>%
  flextable::compose(part = "header", j = "se", value = as_paragraph("SEM")) %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  flextable::fontsize(size = 10, part = "all") %>%
  flextable::bold(part = "header") %>%
  flextable::align(align = "left", part = "all") %>%
  flextable::autofit()

table = read_docx()
table = body_add_flextable(table, value = sintBaMeansPub)
print(table, target = "../tables/table2.docx")

sintBaMeansPub

```

```{r, sint migrationPlots}
sintBaSumm$mean = sprintf('%.3f', sintBaSumm$mean)
sintBaSumm$mean2 = sintBaSumm$mean
sintBaSumm$hpdLow = sprintf('%.3f', sintBaSumm$hpdLow)
sintBaSumm$hpdHigh = sprintf('%.3f', sintBaSumm$hpdHigh)

sintBaLabs = tibble(pop.i = unique(sintBaSumm$pop.i), pop.j = unique(sintBaSumm$pop.j))

sintMigrateA = ggplot(data = sintBaSumm, aes(pop.i, pop.j))+
  geom_tile(data = subset(sintBaSumm, subset = sintBaSumm$mean2>0.65), fill = "gray35", color = "white") +
  geom_segment(data = sintBaSumm, aes(x = 0.4755, xend = -0.42, y = pop.j, yend = pop.j, color = pop.j), size = 14) +
  geom_segment(data = sintBaSumm, aes(x = pop.i, xend = pop.i, y = 0.45, yend = -0.425, color = pop.i), size = 32) +
  scale_color_manual(values = gomPal[c(1:2, 1:5)], guide = NULL) +

  guides(fill = guide_colorbar(ticks.colour = "black", barwidth = 1, barheight = 10, frame.colour = "black")) +
  # new_scale("fill") +
  geom_tile(data = subset(sintBaSumm, subset = sintBaSumm$mean<0.65), aes(fill = as.numeric(as.character(mean))), color = "white") +
  scale_fill_gradientn(colours = paletteer_c("viridis::mako", n = 10, direction = -1)[c(1:7)], limit = c(0,0.23), space = "Lab", name = expression(paste(italic("m"))), na.value = "transparent",  guide = "colourbar", values = c(0, 0.05, 0.1, 0.15, 0.2,0.5,0.75,1)) +
  # scale_fill_gradientn(colours = paletteer_d("khroma::smoothrainbow"), limit = c(0,0.27), space = "Lab", name = expression(paste(italic("m"))), na.value = "transparent",  guide = "colourbar", values = c(0, 0.05, 0.1, 0.15, 0.2,0.5,0.75,1)) +
  geom_text(data = sintBaSumm, aes(x = pop.i, y = pop.j, label = paste(mean, "\n", sep = "")), color = ifelse(sintBaSumm$mean > 0.6, "white", "gray5"), fontface = ifelse(as.numeric(sintBaSumm$hpdLow)>0, "bold", "plain"), size = ifelse(as.numeric(sintBaSumm$hpdLow)>0, 4.75, 4)) +
  geom_text(data = sintBaSumm, aes(x = pop.i, y = pop.j, label = paste("\n(",hpdLow,"",hpdHigh, ")", sep = "")), color = ifelse(sintBaSumm$mean > 0.6, "white", "gray5"), size = 3.25) +
  
  geom_text(data = (sintBaLabs %>% filter(pop.j %in% c("Tortugas Bank\nMesophotic", "Tortugas Bank\nShallow", "Riley's Hump\nMesophotic", "Riley's Hump\nShallow"))), x = -.02, aes(y = pop.j, label = pop.j), size = 3.75, color = "#FFFFFF", family = "sans") +
  geom_text(data = (sintBaLabs %>% filter(!pop.j %in% c("Tortugas Bank\nMesophotic", "Tortugas Bank\nShallow", "Riley's Hump\nMesophotic", "Riley's Hump\nShallow"))), x = -.02, aes(y = pop.j, label = pop.j), size = 3.75, color = "#000000", family = "sans") +
  geom_text(data = (sintBaLabs %>% filter(pop.i %in% c("Tortugas Bank\nMesophotic", "Tortugas Bank\nShallow", "Riley's Hump\nMesophotic", "Riley's Hump\nShallow"))), y = -.01, aes(x = pop.i, label = pop.i), size = 3.75, color = "#FFFFFF", family = "sans") +
  geom_text(data = (sintBaLabs %>% filter(!pop.i %in% c("Tortugas Bank\nMesophotic", "Tortugas Bank\nShallow", "Riley's Hump\nMesophotic", "Riley's Hump\nShallow"))), y = -.01, aes(x = pop.i, label = pop.i), size = 3.75, color = "#000000", family = "sans") +
  
  labs(x = "Sink", y = "Source") +
  scale_y_discrete(limits = rev(levels(sintBaSumm$pop.i))[c(1:8)], position = "left") +
  coord_cartesian(xlim = c(1, 8), ylim = c(1, 8), clip = "off") +
  theme_minimal()

sintMigrate = sintMigrateA + theme(
  axis.text.x = element_text(vjust = 1, size = 12, hjust = 0.5, color = NA),
  axis.text.y = element_text(size = 10, color = NA),
  axis.title.x = element_text(size = 14, hjust = 0.425),
  axis.title.y = element_text(size = 14, hjust = 0.425),
  panel.grid.major = element_blank(),
  axis.ticks = element_blank(),
  # legend.position = c(1.055, 0.5),
  legend.direction = "vertical",
  legend.title = element_text(size = 12, face = "bold")
)

sintMigrate

```

## Put plots together

```{r, migrate plots}
migratePlots = mcavMigrate + inset_element(mcavImg, -0.02, 0.85, 0.17, 1.0, align_to = "full", ignore_tag = TRUE) + 
  sintMigrate +  inset_element(sintImg, -0.02, 0.85, 0.17, 1.0, align_to = "full", ignore_tag = TRUE) +
    plot_annotation(tag_levels = c("A")) +
    plot_layout(ncol = 1, guides = "collect") &
    theme(plot.tag = element_text(size = 18),
          legend.box.margin = unit(units = "cm",c(0,0,0,-2)),
          legend.key = element_blank(),
          legend.background = element_blank())


ggsave("../figures/figure6.png", plot = migratePlots, height = 9.5, width = 9.5, units = "in", dpi = 300)

```


# Other plots

```{r, trees}
gomTrees = mcavStructure + inset_element(mcavImg, 0.02, 0.85, 0.17, 1.0, align_to = "full", ignore_tag = TRUE) + 
  ofavStructure +  inset_element(ofavImg, 0.02, 0.85, 0.17, 1.0, align_to = "full", ignore_tag = TRUE) +
  sintStructure + inset_element(sintImg, 0.02, 0.85, 0.17, 1.0, align_to = "full", ignore_tag = TRUE) + 
  xestoStructure + inset_element(xestoImg, 0.02, 0.85, 0.17, 1.0, align_to = "full", ignore_tag = TRUE) +
    plot_annotation(tag_levels = c("A")) +
    theme(plot.tag = element_text(size = 18),
          legend.spacing = unit(-5, "pt"),
          legend.key = element_blank(),
          legend.background = element_blank())

gomTrees

ggsave("../figures/extras/nwgom_trees.png", plot = gomTrees, dpi = 300, height = 10, width = 12, units = "in")

```


#### Saving R data
```{r, save data}
save.image("nwgom.RData")

```